### 2011-01-11 doc/man/rt1d_manual.tex
Added bandpass averaging discretization technique in RadiationSource.

%%% rt1d

\documentclass[letterpaper,titlepage,12pt]{article}

%%% Preamble

\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\marginparsep}{0in}
\setlength{\marginparwidth}{0in}

\usepackage{amsmath, amsthm, amssymb}
\numberwithin{equation}{section}
\usepackage{graphicx}
\usepackage{pslatex}

%%% Beginning of Document

\begin{document}

\author{Jordan Mirocha}

\title{\Large {\bf rt1d: User's Manual}}
\date{Last Updated: \today}
\maketitle

\setcounter{tocdepth}{2}
\tableofcontents
\newpage

%%% 
%% Methods
%%%
\section{Methods}

%%% 
%% Classes
%%%
\section{Classes}
In order of appearance in \texttt{rt1d.py}.

\subsection{\texttt{InitializeGrid}}
\subsection{\texttt{InitializeIntegralTables}}
\subsection{\texttt{RadiationSource}}
\subsection{\texttt{Radiate}}

%%% 
%% Start to Finish
%%%
\section{Start to Finish}

%%% 
%% Parameters
%%%
\section{Parameters}

\subsection{Initial Conditions}
\begin{description}

\item [\texttt{DensityProfile}]

\item [TemperatureProfile]

\end{description}

\subsection{Control Parameters}

\subsection{Source Parameters}

\end{document}

### 2011-01-11 anl/LocateIfront.py

@@ -29,7 +29,7 @@ sigma_r = LengthUnits / GridDims / cm_per_kpc
 # Analytic solution
 T = f["Data"]["Temperature"].value[0]
 n_H = f["Data"]["HIDensity"].value[-1]
-Ndot = f["ParameterFile"]["SourcePhotonLuminosity"].value
+Ndot = f["ParameterFile"]["SpectrumPhotonLuminosity"].value
 alpha_HII = 2.6e-13 * (T / 1.e4)**-0.85
 trec = 1. / alpha_HII / n_H / s_per_myr
 rs = (3. * Ndot / 4. / np.pi / alpha_HII / n_H**2)**(1. / 3.) / cm_per_kpc

### 2011-01-11 src/RadiationSource.py

@@ -14,7 +14,7 @@ Notes:
 
 import numpy as np
 from scipy.integrate import quad
-import numpy as np
+from scipy.optimize import fsolve
 
 h = 6.626068 * 10**-27 			                            # Planck's constant - [h] = erg*s
 k_B = 1.3806503 * 10**-16			                        # Boltzmann's constant - [k_B] = erg/K
@@ -48,15 +48,7 @@ class RadiationSource:
         self.DiscreteSpectrumMinEnergy = pf["DiscreteSpectrumMinEnergy"]
         self.DiscreteSpectrumMaxEnergy = pf["DiscreteSpectrumMaxEnergy"]
         self.DiscreteSpectrumNumberOfBins = pf["DiscreteSpectrumNumberOfBins"]
-        
-        if self.DiscreteSpectrumMethod == 1:
-            self.DiscreteSpectrumSED = np.array(pf["DiscreteSpectrumSED"])
-        elif self.DiscreteSpectrumMethod == 2:
-            self.DiscreteSpectrumSED = np.linspace(self.DiscreteSpectrumMinEnergy, self.DiscreteSpectrumMaxEnergy, self.DiscreteSpectrumNumberOfBins)
-        elif self.DiscreteSpectrumMethod == 3:
-            self.DiscreteSpectrumSED = np.logspace(np.log10(self.DiscreteSpectrumMinEnergy), np.log10(self.DiscreteSpectrumMaxEnergy), self.DiscreteSpectrumNumberOfBins)
-        else:
-            pass
+        self.DiscreteSpectrumBinEdges = pf["DiscreteSpectrumBinEdges"]
         
         # Set source-specific parameters
         if self.SourceType < 0:
@@ -91,7 +83,36 @@ class RadiationSource:
             self.EmaxNorm = pf["SpectrumMaxNormEnergy"]
             self.alpha = -pf["SpectrumPowerLawIndex"] 
             self.epsilon = pf["SourceRadiativeEfficiency"] 
-        
+                
+        # Source discretization
+        if self.DiscreteSpectrumMethod == 1:
+            self.DiscreteSpectrumSED = np.array(pf["DiscreteSpectrumSED"])
+        elif self.DiscreteSpectrumMethod == 2:
[…]

### 2011-01-11 src/SetDefaultParameterValues.py

@@ -86,6 +86,8 @@ def SetDefaultParameterValues():
           "DiscreteSpectrumMinEnergy": 100., \
           "DiscreteSpectrumMaxEnergy": 1e4, \
           "DiscreteSpectrumNumberOfBins": 20, \
+          "DiscreteSpectrumSmoothingWidth": 10, \
+          "DiscreteSpectrumBinEdges": [13.6, 20, 30, 40, 50], \
           "DiscreteSpectrumSED": [13.61], \
           
           # Cosmological Parameters (WMAP 7)

### 2011-01-22 doc/man/rt1d_manual.tex
Needed to add rtan.py to the repository, previous name was rtanl.py.

@@
 
+\subsection{Problem Types}
+These were added for convenience in January, 2011.
+If you want to make changes to these test problems you can, just make sure \texttt{ProblemType} is the first (uncommented) line in your parameter file.
+Additional parameters afterwards will be interpreted as usual.
+
+\begin{description}
+
+\item [\texttt{ProblemType = 1}] This will set up a standard test with known analytic solution: the propagation of an ionization front in an isothermal, hydrogen-only, static universe.
+The spectrum used is monochromatic at $E = 13.6 \ \text{eV}$, with photon luminosity $L_{\nu} = 5\times 10^{48} \ s^{-1}$.
+The initial density, temperature, and box size are the same as the values used in Test 1 of Wise \& Abel 2010.
+
+\item [\texttt{ProblemType = 2}] This problem is almost exactly the same as \texttt{ProblemType = 1}.
+However, the temperature is allowed to evolve, and instead of using a monochromatic spectrum, we sample a $10^5$ K blackbody with four energy groups.
+Again, the initial density, temperature, box size, and energy groups are the same as the values used in Test 1 of Wise \& Abel 2010.
+
+\item [\texttt{ProblemType = 2.1}] Same as \texttt{ProblemType = 2} for everything, except we use a continuous blackbody spectrum rather than a discrete one.
+
+\end{description}
+
 \subsection{Initial Conditions}

### 2011-01-22 src/InitializeParameterSpace.py

@@ -46,7 +46,6 @@ class InitializeParameterSpace:
             if parname.strip() == 'ProblemType' and float(parval) > 0:
                 pf_new = self.ProblemType(float(parval))
                 for param in pf_new: pf_dict[param] = pf_new[param]
-                break
                 
             # Else, actually read in the parameter                                     
             try: parval = float(parval)
@@ -176,7 +175,7 @@ class InitializeParameterSpace:
                   "SavePrefix": 'rt', "SolveTemperatureEvolution": 1, "MultiSpecies": 0, "SecondaryElectronMethod": 0, "CosmologicalExpansion": 0, \
                   "DensityProfile": 0, "InitialDensity": 1e-3, "TemperatureProfile": 0, "InitialTemperature": 1e2, \
                   "IonizationProfile": 1, "InitialHIIFraction": 1.2e-3, "SourceType": 1, "SourceLifetime": 500.0, \
-                  "SourceTemperature": 1e5, "DiscreteSpectrumMethod": 0
+                  "SourceTemperature": 1e5, "DiscreteSpectrumMethod": 0, "SpectrumMinEnergy": 10., "SpectrumMaxEnergy": 100,
                  }               
                 
             

### 2011-08-26 doc/paper/ms.tex
Added paper to repo.

%
%  Template for writing APJ style papers
%
%  Created by Jordan Mirocha on 2009-09-17.
%  Copyright (c) 2009 . All rights reserved. (really? ha)
%

%---------------------------------------- Declarations -----------------------------------------%

%\documentclass{emulateapj}
%\documentclass[12pt,preprint]{aastex}
%\usepackage{emulateapj5}
%\usepackage{apjfonts}
%
%% Define General Packages
%\usepackage{graphicx}
%\usepackage{pslatex}
%\usepackage{amsmath, amsthm, amssymb}
%\usepackage{natbib}
%\usepackage{epsfig}
%
%% Use utf-8 encoding for foreign characters
%\usepackage[utf8]{inputenc}

%%% Declarations
\documentclass[preprint2]{aastex}              % double-column, single-spaced document
%\documentclass[preprint2,longabstract]{aastex}% if abstract too long
%\documentclass[manuscript]{aastex}            % one-column, double-spaced document:

\usepackage{graphicx}
\usepackage{pslatex}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{natbib}
\usepackage{epsfig}

% Short names
\shorttitle{Optimized Multi-Frequency Spectra for Computational Radiative Transfer}
\shortauthors{Mirocha & Darling}

%-------------------------------------- PAPER STARTS HERE --------------------------------------%

\begin{document}

%%% START custom commands
\newcommand{\HI}{\text{HI}}
\newcommand{\HII}{\text{HII}}
\newcommand{\HeI}{\text{HeI}}
\newcommand{\HeII}{\text{HeII}}
\newcommand{\HeIII}{\text{HeIII}}
\newcommand{\nHI}{n_{\text{HI}}} 
\newcommand{\nHII}{n_{\text{HII}}}  
\newcommand{\nHeI}{n_{\text{HeI}}}
\newcommand{\nHeII}{n_{\text{HeII}}}
\newcommand{\nHeIII}{n_{\text{HeIII}}}  
\newcommand{\nel}{n_{\text{e}}}  
\newcommand{\nb}{n_{\text{b}}}
\newcommand{\nnu}{$n_{\nu}$ }
\newcommand{\ncol}{n_i^{\text{col}}}
\newcommand{\gHI}{\Gamma_{\text{HI}}}  
\newcommand{\gHeI}{\Gamma_{\text{HeI}}}
\newcommand{\gHeII}{\Gamma_{\text{HeII}}}
\newcommand{\aHII}{\alpha_{\text{HII}}}  
\newcommand{\aHeII}{\alpha_{\text{HeII}}}  
\newcommand{\aHeIII}{\alpha_{\text{HeIII}}} 
\newcommand{\bHeI}{\beta_{\text{HeI}}}  
\newcommand{\bHeII}{\beta_{\text{HeII}}}  
\newcommand{\xHeII}{\xi_{\text{HeII}}}
\newcommand{\kB}{k_{\text{B}}}
\newcommand{\fheat}{f_{\text{heat}}}
\newcommand{\Lbol}{\mathcal{L}_{\text{bol}}}
\newcommand{\spec}{\mathcal{N}}
\newcommand{\Heat}{\mathcal{H}}
%%% END custom commands

\title {Optimized Multi-Frequency Spectra for Numerical Radiative Transfer}
\author{Jordan Mirocha, Stephen Skory, Jack Burns}
\affil{Center for Astrophysics and Space Astronomy, University of Colorado, Campus Box 389, Boulder, CO 80309}
\affil{NASA Lunar Science Institute, NASA Ames Research Center, Moffett Field, CA 94043, USA}

%%%
%% Abstract
%%%
\begin{abstract}
The recent emergence and implementation of complex algorithms for radiative transfer in numerous cosmological hydrodynamics codes has led to a dramatic improvement in simulations of radiative feedback in various astrophysical environments.  However, due to computational expense, the spectra of radiation sources are often discretized into a small number of frequency bins.  Using simple 1D radiative transfer calculations around model stars and accreting black holes, we find that even in the idealized case of a static density field the reduction of continuous spectra to only a few discrete emission frequencies has significant consequences for the resulting temperature and ionization profiles.  We quantify the effects of different discretization schemes for objects of different luminosities and spectral models, and propose a new technique to reduce errors while minimizing computational expense.
\end{abstract} 
\subjectheadings {computational methods -- radiative transfer -- reionization -- radiative feedback}

%%%
%% Introduction
%%%
\section{Introduction}
The recent development of advanced numerical methods for radiative transfer has led to a substantial increase in the number of astrophysical problems that are computationally accessible.   Several fully 3D cosmological hydrodynamics codes have implemented adaptive ray-tracing methods (\cite{Abel2002}) for propagating radiation around point sources, including \cite{Wise2011} and others.  Flux-limited diffusion schemes (\cite{Reynolds2009} and others) have begun to surface as well, both in cosmological problems and accretion disk problems.  More recently, hybrid-schemes (VET, Petkova \& Springel, etc.) have begun to make an appearance.  For a review on the state of convergence among these varying numerical methods, see \cite{Iliev2006} and \cite{Iliev2009}.

Unfortunately, the vast majority of radiative transfer simulations run to date have only considered radiation at a single frequency.  While this monochromatic approach is the natural first step computationally-speaking, sampling radiation fields at multiple frequencies is essential for accurately calculating the properties of the medium through which photons propagate.  The mere presence of ionizing photons ensures that gas surrounding a radiation source will become ionized and heated, but the degree to which gas is ionized and heated depends on the total number of ionizing photons and their spectral energy distribution (SED).  As a result, we should expect the properties of gas surrounding a blackbody source of ultraviolet (UV) photons to be quite different from those of a power-law X-ray source such as an accreting black hole.  Even relatively small changes in SED can produce noticeable differences in surrounding gas properties, like changing the power-law index of a black-hole accretion spectrum (\cite{Kuhlen2005}; \cite{Thomas2008}).  But, these expectations are built on the assumption that each source is represented by a continuous SED, not one we have constructed with \nnu discrete frequency bins.  As we'll see in \S2, for a given source, the properties of the surrounding medium can be altered considerably solely due to the manner in which we discretize its spectrum.

The consequences of poor frequency resolution have been explored recently by \cite{Wise2011}, who find that for the expansion of an HII region around a $10^5$ K blackbody source in a hydrogen-only medium, the density, temperature, velocity, and ionization profiles are well converged for $n_{\nu} \ge 4$.  Use of a monochromatic spectrum for this problem introduces significant errors since all photons are absorbed at a characteristic column density, whereas multi-frequency treatments achieve some column density dependent behavior, and can thus mimic the behavior of a truly continuous spectrum.  The convergence for this test problem is reassuring, though the prospects for convergence are less clear if one were interested in the absorption processes of multiple chemical species, ionization and heating due to X-rays and their energetic secondary photo-electrons (\cite{Shull1985}; \cite{Furlanetto2010}), or inhomogeneous media.  Such effects are likely important in studies of radiative feedback from stars and active galactic nuclei (AGN), and most certainly in efforts to simulate cosmological reionization.  As a result, an effort must be made to ensure the SEDs used in numerical simulations to represent these sources produce

Though we will find that multi-frequency radiative transfer is necessary in most cases to accurately calculate the ionization and temperature state of the gas, it can also add significant computational expense.  Because the efficiency with which photons ionize and heat a gas depends on their energy, adding more energy groups will alter the amount of ionization and heating taking place in a given computational grid element.  But, these properties of the gas also dictate the computational time-step.  Generally, it is recommended that radiation time-steps are limited by a maximum fractional change in neutral species abundances so that ionization fronts are propagated at the correct speed (\cite{Shapiro2004}).  As a result, the addition of photons that ionize gas more efficiently will lead to a more expensive calculation since the time-step necessary to restrict changes in the neutral fraction will shrink.  Similarly, radiation-hydrodynamic simulations may be slowed by the inclusion of photons that heat gas efficiently, since higher temperature cells have greater sound speeds, and thus require a smaller time-step.  Expense due to additional energy groups is certainly method-dependent, and may even be implementation-dependent.  For example, including high energy photons with long mean free paths will be very costly for a ray-tracing algorithm, but add no expense to a problem solved by flux-limited diffusion.  In addition, some adaptive ray-tracing implementations require each additional energy group in the source spectrum to be traced independently, in which case the time required by the radiative transfer solver scales roughly as \nnu.  Other implementations are now avoiding this problem by tracing all energy groups at once, though keeping them separate can be advantageous since energy groups with small mean free paths will be completely attenuated at small radii (\cite{Wise2011}).

With all of these factors in mind, we want to determine the following.  How many energy groups are required to adequately sample the true SED of a given radiation source?  Where should they be positioned in frequency?  What fraction of the bolometric luminosity should be allocated to each energy group?  What are the consequences of spectral discretization both physically and computationally?

The contents of this paper fall into three general areas.  In \S2 we will quantitatively assess the accuracy with which multi-frequency calculations reproduce the ionization and heating profiles of continuous spectral energy distributions.  \S3 is devoted to introducing a technique for optimally selecting discrete SED templates so as to minimize the errors between discrete and continuous spectra, and \S4 will present the results obtained with this method.  In \S5, we conclude.  Details concerning the code developed for this work, as well as a discussion regarding the factors that influence the solution of radiative transfer calculations, whether it be the time-step, grid resolution, frequency resolution, or the interplay of all three, can be found in the Appendix.

%Reduce guesswork!  Choose frequency resolution in the same way that one chooses spatial/mass resolution.

%%% 
%% Methods & Motivation
%%%
\section{Methods \& Motivation}
The key to our subsequent analysis was inspired by this same convenient feature: if $\Gamma_i$ and $\Heat_i$ can be tabulated for a given source \textit{a-priori}, it means that \textit{precision determination of the ionization state and temperature of the gas surrounding that source relies solely on accurate numerical computation of the photoionization and heating rate integrals in Equations \ref{PhotoionizationRate} and \ref{HeatingRate}}.  So, for a given source, we can compute the values of these integrals numerically for the continuous spectrum, and then as discrete sums given $n_{\nu}$ frequency bins with 
\begin{equation}
    I(E) = \sum_{j=1}^{n_{\nu}} I(E_j) .
\end{equation}

% Code Verification
%\subsection{Code Verification}
%Our code, \texttt{rt1d}, solves Equations \ref{HIIRateEquation}, \ref{HeIIRateEquation}, \ref{HeIIIRateEquation}, and \ref{TemperatureEvolution} using the implicit Euler method for integration and the Newton-Raphson technique for root finding.  In order to track the propagation of ionization fronts accurately, we limit the time-step based on a maximum neutral fraction change as introduced in \cite{Shapiro2004},
%\begin{equation}
%    \Delta t = \epsilon_{\mathrm{ion}} \frac{\nHI}{|d\nHI/dt|} ,
%\end{equation}
%and additionally require that the time-step increase by a factor of two at most, as in \cite{Wise2011}.  
%
%To demonstrate the functionality of the code, we repeat tests 1 and 2 from the Radiative Transfer Comparison Project (\cite{Iliev2006}; hereafter referred to as RT06).  Test 1 is the expansion of an HII region in a hydrogen-only, isothermal medium surrounding a monochromatic source of 13.6 eV photons.  Test 2 is the same, except for the temperature is allowed to evolve according to Equation \ref{TemperatureEvolution}, and the monochromatic radiation source is replaced by a $10^5$ K blackbody spectrum.  Results for Test 1 are shown in Figures \ref{fig:RT_Test1_IfrontEvolution} and \ref{fig:RT_Test1_RadialProfiles}, and for Test 2 in Figure \ref{fig:RT_Test2_RadialProfiles}.
%\begin{figure*}
%\begin{center}
%\leavevmode
%\psfig{file=figures/RT_Test1_IfrontEvolution.ps,width=6in}
%\vspace{-20pt}
%\caption[]{\textit{Top:} Comparison of the calculated (dashed) and analytic (solid) positions of an expanding ionization front in a hydrogen-only, isothermal medium. \textit{Bottom:} Ratio of the calculated and analytic solutions.}
%\label{fig:RT_Test1_IfrontEvolution}
%\end{center}
%\end{figure*}
%
%\begin{figure*}
%\begin{center}
%\leavevmode
%\psfig{file=figures/RT_Test1_RadialProfiles.ps,width=6in}
%\vspace{-20pt}
%\caption[]{Radial profiles of the neutral (solid) and ionized (dashed) fractions at $t = 10$, $30$, $100$, and $500$ Myr.}
%\label{fig:RT_Test1_RadialProfiles}
%\end{center}
%\end{figure*}
%
%\begin{figure*}
%\begin{center}
%\leavevmode
%\psfig{file=figures/RT_Test2_RadialProfiles.ps,width=6in}
%\vspace{-20pt}
%\caption[]{\textit{Top:} Radial profiles of the neutral (solid) and ionized (dashed) fractions at $t = 10$, $100$, and $500$ Myr. \textit{Bottom:} Radial profiles of the kinetic temperature at $t = 10$, $100$, and $500$ Myr (solid, dashed, and dotted lines, respectively).}
%\label{fig:RT_Test2_RadialProfiles}
%\end{center}
%\end{figure*}

% Discretization Artifacts
\subsection{Discretization Artifacts}

% Optimization Strategy
\subsection{Optimization Strategy}
Our goal is to find the discrete spectrum, $I(E_j)$, that optimally reproduces the values of $\Gamma_i$ and $\Heat_i$ as a function of column density.  Denoting the photoionization and heating rates for a discrete spectrum as $\Gamma_i^d$ and $\Heat_i^d$, we want to minimize the difference between continuous and discrete solutions, i.e. $|\Gamma_i - \Gamma_i^d| = |\Heat_i - \Heat_i^d| = 0$.  As we will see in subsequent sections, these functions span several orders of magnitude in the ordinate over the column density regime of interest.  In order to place equal emphasis on the accuracy of solutions for high column densities (for which $\Gamma_i$ and $\Heat_i$ are small), it is more practical to seek solutions to
\begin{equation}
    \mathrm{log_{10}}\left | \frac{\Gamma_i}{\Gamma_i^d}\right | = \mathrm{log_{10}}\left | \frac{\Heat_i}{\Heat_i^d}\right | = 0 .
\end{equation}    
To do this, we employ a Monte-Carlo method called Simulated Annealing, in which we run $K$ Monte-Carlo trials, or random walks, each of which consists of $L$ steps.  In order to steer each random walk towards the global minimum, we will evaluate the quantity  
\begin{equation}
    P = \mathrm{exp}\left[-(f_{k,l} - f_{k,l-1}) / T \right]  \label{Probability}
\end{equation}
where $l=0,1,2,...L$ represents the current step in the current random walk, $k$, where $k = 0,1,2,...K$.  At each step in a given random walk, we generate a random number, $q$, that will determine whether we keep our current guess, $I(E_{k,l})$, or revert to our previous guess, $I(E_{k,l-1})$.  The condition for keeping our current guess is
\begin{equation}
    P \ge q .
\end{equation}
The key aspect of this analysis is how we vary the control parameter $T$, which is called the temperature in analogy with Boltzmann's equation.  Equation \ref{Probability} tells us that regardless of the value of $T$, if $f_l < f_{l-1}$ (i.e. our guess is better than the last), then
\begin{equation}
    P \ge 1
\end{equation}
and we have a 100\% chance of keeping our current guess.  It is clear now that how we control the temperature only effects how we deal with bad guesses (decreasing the temperature means we become less tolerant of bad guesses).  There are many ways of doing this, but for simplicity we adopt the following technique.  Every $m$ steps, we take
\begin{equation}
    T \rightarrow \gamma T ,
\end{equation}
where $\gamma$ is an experimentally determined quantity of order unity.

\section{Results}

\section{Why not tabulate? (discussion)}

% Could FLD methods tabulate? OTVET? RSPH?  MC?

\section{Conclusions}

\bibliography{references}
\bibliographystyle{apj}

%%%
%% APPENDIX
%%%
\appendix

% RT Framework
\section{A. Radiative Transfer Framework}

Our radiative transfer scheme is modeled closely after the work of \cite{Fukugita1994} and \cite{Thomas2008}, though for completeness we will reiterate the aspects of these methods most pertinent to the problem at hand.  In general, propagating a radiation field relies on solving differential equations governing the rate of change in the number densities of all ions and the temperature of the gas.  Assuming a medium consisting of hydrogen and helium only, we first solve for the relative abundances of each ion via
\begin{align}
    \frac{d \nHII}{dt} & = \gHI \nHI - \aHII \nel \nHII   \label{HIIRateEquation} \\
    \frac{d \nHeII}{dt} & = \gHeI \nHeI - \bHeI \nel \nHeI + \bHeII \nel \nHeII \label{HeIIRateEquation} \\ \nonumber 
    & - \aHeII \nel \nHeII + \aHeIII \nel \nHeIII - \xHeII \nel \nHeIII  \\
    \frac{d \nHeIII}{dt} & = \gHeII \nHeII - \bHeII \nel \nHeII + \aHeIII \nel \nHeIII . \label{HeIIIRateEquation}
\end{align}
Each of these equations represents the balance of ionizations and recombinations for ion $i = \HII, \HeII, \HeIII$, where $\Gamma_i$ denotes the photoionization rate coefficient, $\alpha_i$ ($\xi_i$) the case-A (dielectric) recombination rate coefficients, and $\beta_i$ the collisional ionization rate coefficients.  In this work, we use the formulae in Appendix B of \cite{Fukugita1994} to compute the values of $\alpha$, $\beta$, and $\xi$.  The photoionization rate coefficients, $\Gamma_i$, are given by
\begin{equation}
    \Gamma_i = \int_{E_i}^{\infty} \sigma_i(E) \spec(E) \frac{dE}{E}    \label{PhotoionizationRate}
\end{equation}
where $E_i$ and $\sigma_i$ are the ionization threshold and bound-free absorption cross section for species $i$, respectively.  $\spec$ is the spectral intensity, which we write as
\begin{equation}
    \spec(E, r) = \frac{\Lbol}{4\pi r^2} I(E) e^{-\tau(E, r)} .
\end{equation}
Here, $r$ is the distance from the source, $I(E)$ is the normalized spectral energy distribution (SED), satisfying
\begin{equation}
    \int_E I(E) dE = 1 ,
\end{equation}
and the optical depth $\tau(E, r)$ is given as a sum over absorbing species,
\begin{align}
    \tau(E, r) & = \sum_i \int_r \sigma_i(E) n_i(r) dr \nonumber \\
               & = \sum_i \sigma_i(E) \ncol(r)
\end{align}
where $\ncol(r)$ is the column density of species $i$ at distance $r$ from the source.
We calculate the bound-free absorption cross-sections, $\sigma_i$, using the fits of \cite{Verner1996}.  At each time-step, we also solve for the temperature evolution, which is given by
\begin{align}
    \frac{3}{2}\frac{d}{dt}\left(\frac{\kB T \nb}{\mu}\right) & = \fheat \sum_i n_i \Heat_i \nonumber \\
    & - \sum_i \zeta_i n_e n_i - \sum_i \eta_i n_e n_i - \sum_i \psi_i n_e n_i \label{TemperatureEvolution} 
\end{align}
where $\Heat_i$ is the photo-electric heating rate, given by
\begin{equation}
    \Heat_i = \int_{E_i}^{\infty} \sigma_i(E) \spec(E, r)(E - E_i) \frac{dE}{E} \label{HeatingRate} ,
\end{equation}    
and $\zeta_i$, $\eta_i$, and $\psi_i$ are the collisional ionization, recombination, and collisional excitation cooling coefficients, respectively.  Additional parameters include $\nb$, the total number density of baryons (check on this), the mean molecular weight $\mu$, Boltzmann's constant $\kB$, the kinetic temperature of the gas $T$, and $\fheat$, the fraction of a photon's original energy deposited as heat (see \cite{Shull1985}, \cite{Furlanetto2010}).  For simplicity, we neglect Hubble cooling due to cosmological expansion, as well as Compton heating/cooling, and cooling by free-free emission.

It is important to note that we can pull radial and time dependencies outside of the integrals in Equations \ref{PhotoionizationRate} and \ref{HeatingRate}, leaving
\begin{equation}
    \Gamma_i(\ncol) = A \int_{E_i}^{\infty} \sigma_i(E) I(E) e^{-\tau(E, \ncol)} \frac{dE}{E} \
\end{equation}
and
\begin{equation}
    \Heat_i(\ncol) = A \int_{E_i}^{\infty} \sigma_i(E) I(E) (E - E_i) e^{-\tau(E, \ncol)} \frac{dE}{E}
\end{equation}
where $A \equiv \Lbol / 4 \pi r^2$.

It is now apparent that the values of $\Gamma_i$ and $\Heat_i$ can be tabulated as a function of column density for a given source.  This greatly improves the speed of 1D radiative transfer calculations, like those of \cite{Thomas2008}, since the integrals in these expressions (Equations \ref{PhotoionizationRate} and \ref{HeatingRate}) no longer need to be computed numerically on-the-fly.  However, the lookup tables will only be 1-dimensional for the case of a hydrogen-only universe.

% Code Verification
\section{B. Code Verification}
Our code, \texttt{rt1d}, solves Equations \ref{HIIRateEquation}, \ref{HeIIRateEquation}, \ref{HeIIIRateEquation}, and \ref{TemperatureEvolution} using the implicit Euler method for integration and the Newton-Raphson technique for root finding.  In order to track the propagation of ionization fronts accurately, we limit the time-step based on a maximum neutral fraction change as introduced in \cite{Shapiro2004},
\begin{equation}
    \Delta t = \epsilon_{\mathrm{ion}} \frac{\nHI}{|d\nHI/dt|} ,
\end{equation}
and additionally require that the time-step increase by a factor of two at most, as in \cite{Wise2011}.

At this juncture it is worth mentioning the numerical artifacts that can result due to poor temporal or spatial resolution.

To demonstrate the functionality of the code, we repeat tests 1 and 2 from the Radiative Transfer Comparison Project (\cite{Iliev2006}; hereafter referred to as RT06).  Test 1 is the expansion of an HII region in a hydrogen-only, isothermal medium surrounding a monochromatic source of 13.6 eV photons.  Test 2 is the same, except for the temperature is allowed to evolve according to Equation \ref{TemperatureEvolution}, and the monochromatic radiation source is replaced by a $10^5$ K blackbody spectrum.  Results for Test 1 are shown in Figures \ref{fig:RT_Test1_IfrontEvolution} and \ref{fig:RT_Test1_RadialProfiles}, and for Test 2 in Figure \ref{fig:RT_Test2_RadialProfiles}.
\begin{figure*}
\begin{center}
\leavevmode
\psfig{file=figures/RT_Test1_IfrontEvolution.ps,width=6in}
\vspace{-20pt}
\caption[]{\textit{Top:} Comparison of the calculated (dashed) and analytic (solid) positions of an expanding ionization front in a hydrogen-only, isothermal medium. \textit{Bottom:} Ratio of the calculated and analytic solutions.}
\label{fig:RT_Test1_IfrontEvolution}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\leavevmode
\psfig{file=figures/RT_Test1_RadialProfiles.ps,width=6in}
\vspace{-20pt}
\caption[]{Radial profiles of the neutral (solid) and ionized (dashed) fractions at $t = 10$, $30$, $100$, and $500$ Myr.}
\label{fig:RT_Test1_RadialProfiles}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\leavevmode
\psfig{file=figures/RT_Test2_RadialProfiles.ps,width=6in}
\vspace{-20pt}
\caption[]{\textit{Top:} Radial profiles of the neutral (solid) and ionized (dashed) fractions at $t = 10$, $100$, and $500$ Myr. \textit{Bottom:} Radial profiles of the kinetic temperature at $t = 10$, $100$, and $500$ Myr (solid, dashed, and dotted lines, respectively).}
\label{fig:RT_Test2_RadialProfiles}
\end{center}
\end{figure*}

\end{document}

### 2011-08-29 doc/paper/ms.tex
c != inf seems to be working correctly, at least for a constant timestep.

@@
 The recent development of advanced numerical methods for radiative transfer has led to a substantial increase in the number of astrophysical problems that are computationally accessible.
-Several fully 3D cosmological hydrodynamics codes have implemented adaptive ray-tracing methods (\cite{Abel2002}) for propagating radiation around point sources, including \cite{Wise2011} and others.
-Flux-limited diffusion schemes (\cite{Reynolds2009} and others) have begun to surface as well, both in cosmological problems and accretion disk problems.
-More recently, hybrid-schemes (VET, Petkova \& Springel, etc.) have begun to make an appearance.
-For a review on the state of convergence among these varying numerical methods, see \cite{Iliev2006} and \cite{Iliev2009}.
+Several fully 3D cosmological hydrodynamics codes have implemented adaptive ray-tracing methods (\cite{Abel2002}) for propagating radiation around point sources,  or flux-limited diffusion schemes (\cite{Reynolds2009}) for diffuse radiation fields.
+More recently, hybrid-schemes have been implemented in smoothed-particle hydrodynamic codes (e.g.
+\cite{Petkova2011}).
+For a review on the state of convergence among several widely used codes, see \cite{Iliev2006} and \cite{Iliev2009}.
 
@@
 Such effects are likely important in studies of radiative feedback from stars and active galactic nuclei (AGN), and most certainly in efforts to simulate cosmological reionization.
-As a result, an effort must be made to ensure the SEDs used in numerical simulations to represent these sources produce
+As a result, an effort must be made to ensure that the SEDs used to represent stars in black holes in numerical simulations actually reflect the behavior expected for those sources.
+Additional motivation for this work comes from the fact that while spatial, temporal, and mass resolution for a simulation are simple to calculate and choose to optimize for a given problem, frequency resolution is not.
+We aim to reduce the guesswork involved in discretizing SEDs, and provide a method for optimization on a problem-by-problem basis, just as one does for the other forms of resolution.
 
@@
 %%% 
-%% Methods & Motivation
+%% Multi-Frequency vs. Continuous
 %%%
-\section{Methods \& Motivation}
+\section{Consequences of Discrete Radiation Fields}
+Before we compute the optimal multi-frequency spectra for SEDs commonly used in numerical simulations, we will first demonstrate the difference in results brought about by varying frequency resolution.
+To do this, we use \texttt{rt1d}, a 1D radiative transfer code developed for this work that is described in more detail in the appendices.
+We choose to use 1D
+
+\section{Methods}
 The key to our subsequent analysis was inspired by this same convenient feature: if $\Gamma_i$ and $\Heat_i$ can be tabulated for a given source \textit{a-priori}, it means that \textit{precision determination of the ionization state and temperature of the gas surrounding that source relies solely on accurate numerical computation of the photoionization and heating rate integrals in Equations \ref{PhotoionizationRate} and \ref{HeatingRate}}.
@@
 \end{align}
-Each of these equations represents the balance of ionizations and recombinations for ion $i = \HII, \HeII, \HeIII$, where $\Gamma_i$ denotes the photoionization rate coefficient, $\alpha_i$ ($\xi_i$) the case-A (dielectric) recombination rate coefficients, and $\beta_i$ the collisional ionization rate coefficients.
+Each of these equations represents the balance of ionizations and recombinations for ion $i = \HII, \HeII, \HeIII$, where $\Gamma_i$ denotes the ionization rate coefficient, $\alpha_i$ ($\xi_i$) the case-A (dielectric) recombination rate coefficients, and $\beta_i$ the collisional ionization rate coefficients.
 In this work, we use the formulae in Appendix B of \cite{Fukugita1994} to compute the values of $\alpha$, $\beta$, and $\xi$.
-The photoionization rate coefficients, $\Gamma_i$, are given by
+The ionization rate coefficients are composed of terms taking into account photo-ionization, collisional ionization, and secondary ionizations due to energetic photo-electrons.
+The photoionization rate coefficients, $\Gamma_i^P$, are given by
 \begin{equation}

### 2011-08-29 rt1d/mods/Radiate.py

@@ -87,6 +87,7 @@ class Radiate:
         self.grid = np.arange(self.GridDimensions)
         self.r = self.LengthUnits * self.grid / self.GridDimensions  
         self.dx = self.LengthUnits / self.GridDimensions
+        self.CellCrossingTime = self.dx / c
         self.HIColumn = n_col[0]
         self.HeIColumn = n_col[1]
         self.HeIIColumn = n_col[2]
@@ -239,6 +240,8 @@ class Radiate:
         # Deal with c < infinity
         if not self.InfiniteSpeedOfLight: 
             
+            # PhotonPackage guide: pack = [EmissionTime, EmissionTimeInterval, ncolHI, ncolHeI, ncolHeII, Energy]
+            
             # Photon packages going from oldest to youngest - will have to create it on first timestep
             try: packs = list(data['PhotonPackages']) 
             except KeyError: packs = []            
@@ -252,8 +255,8 @@ class Radiate:
                     packs[-1][-1] += Lbol * dt
                 else:
                     # Launch new photon package - [t_birth, ncol_HI_0, ncol_HeI_0, ncol_HeII_0]                
-                    packs.append(np.array([t, 0., 0., 0.]))
-            else: packs.append(np.array([t, 0., 0., 0., Lbol * dt]))
+                    packs.append(np.array([t, dt, 0., 0., 0., Lbol * dt]))
+            else: packs.append(np.array([t, dt, 0., 0., 0., Lbol * dt]))
         
         if self.InfiniteSpeedOfLight:
 
@@ -339,7 +342,7 @@ class Radiate:
                 newdata["Temperature"][cell] = newT    
                 
                 if self.HIIRestrictedTimestep: 
-                    dtphot[cell] = self.ComputePhotonTimestep(newdata, cell_pack, r, Lbol, n_H, n_He)                            
+                    dtphot[cell] = self.ComputePhotonTimestep(newdata, cell, r, Lbol, n_H, n_He)                            
                 
                 ######################################
                 ################ DONE ################
@@ -347,42 +350,47 @@ class Radiate:
                 
         # If the speed of light = c, things are trickier    
         else:                
-            
+                        
             # Loop over photon packages, updating values in cells: data -> newdata
             for j, pack in enumerate(packs):
                 t_birth = pack[0]
-                r_pack = (t - t_birth) * c                      # Position of package before evolving photons
-                r_max = r_pack + (t + dt - t_birth) * c         # Furthest this package will get this timestep
-                                                
[…]

### 2011-09-01 doc/paper/ms.tex
For HII restricted step, now only consider cells at tau < 0.5.

@@
 \usepackage{epsfig}
+\usepackage{appendix}
 
@@
 \shorttitle{Optimized Multi-Frequency Spectra for Computational Radiative Transfer}
-\shortauthors{Mirocha & Darling}
+\shortauthors{Mirocha et al.}
 
@@
 \affil{NASA Lunar Science Institute, NASA Ames Research Center, Moffett Field, CA 94043, USA}
+\author{John Wise}
+\affil{Georgia Institute of Technology}
+\author{Steven Furlanetto}
+\affil{University of California at Los Angeles}
 
@@
 %%%
-\section{Introduction}
+\section{INTRODUCTION}
 The recent development of advanced numerical methods for radiative transfer has led to a substantial increase in the number of astrophysical problems that are computationally accessible.
@@
 The mere presence of ionizing photons ensures that gas surrounding a radiation source will become ionized and heated, but the degree to which gas is ionized and heated depends on the total number of ionizing photons and their spectral energy distribution (SED).
-As a result, we should expect the properties of gas surrounding a blackbody source of ultraviolet (UV) photons to be quite different from those of a power-law X-ray source such as an accreting black hole.
+As a result, we should expect the properties of gas surrounding a blackbody source of ultraviolet (UV) photons to be quite different from those of a power-law X-ray source such as an accreting black hole, for example.
 Even relatively small changes in SED can produce noticeable differences in surrounding gas properties, like changing the power-law index of a black-hole accretion spectrum (\cite{Kuhlen2005}; \cite{Thomas2008}).
@@
 Such effects are likely important in studies of radiative feedback from stars and active galactic nuclei (AGN), and most certainly in efforts to simulate cosmological reionization.
-As a result, an effort must be made to ensure that the SEDs used to represent stars in black holes in numerical simulations actually reflect the behavior expected for those sources.
-Additional motivation for this work comes from the fact that while spatial, temporal, and mass resolution for a simulation are simple to calculate and choose to optimize for a given problem, frequency resolution is not.
+As a result, an effort must be made to ensure that the SEDs used to represent stars and black holes in numerical simulations actually reflect the behavior expected for those sources.
+Additional motivation for this work comes from the fact that while spatial, temporal, and mass resolution for simulations are simple to calculate and straightforward to optimize, frequency resolution is not.
 We aim to reduce the guesswork involved in discretizing SEDs, and provide a method for optimization on a problem-by-problem basis, just as one does for the other forms of resolution.
@@
 In \S2 we will quantitatively assess the accuracy with which multi-frequency calculations reproduce the ionization and heating profiles of continuous spectral energy distributions.
-\S3 is devoted to introducing a technique for optimally selecting discrete SED templates so as to minimize the errors between discrete and continuous spectra, and \S4 will present the results obtained with this method.
+\S3 is devoted to introducing a technique for optimally selecting discrete SED templates so as to minimize the errors between discrete and continuous solutions, and \S4 will present the results obtained with this method.
 In \S5, we conclude.
-Details concerning the code developed for this work, as well as a discussion regarding the factors that influence the solution of radiative transfer calculations, whether it be the time-step, grid resolution, frequency resolution, or the interplay of all three, can be found in the Appendix.
+Details concerning the code used for this work, as well as a discussion regarding the factors that influence the solution of radiative transfer calculations, whether it be the temporal resolution, spatial resolution, frequency resolution, or the interplay of all three, can be found in the Appendix.
 
@@
 %%%
-\section{Consequences of Discrete Radiation Fields}
-Before we compute the optimal multi-frequency spectra for SEDs commonly used in numerical simulations, we will first demonstrate the difference in results brought about by varying frequency resolution.
-To do this, we use \texttt{rt1d}, a 1D radiative transfer code developed for this work that is described in more detail in the appendices.
-We choose to use 1D
+\section{ASSESSING THE CONSEQUENCES OF DISCRETE RADIATION FIELDS}
+Before computing the optimal multi-frequency spectra for SEDs most commonly used in numerical simulations (\S3), we will first demonstrate the difference in results brought about by varying degrees of frequency resolution.
+To do this, we use \texttt{rt1d}, a 1D radiative transfer code developed for this work that is described and validated in the appendices.
 
-\section{Methods}
+1D calculations were the natural choice for the problem at hand for a variety of reasons.
+Since our sole focus is on the resolution dependence of radiative effects, it would be unnecessary to perform calculations in a more complex setting with additional unrelated physics.
+In fact, doing so would likely make our results more opaque (ha).
+Additionally, the algorithm of choice allows one to pre-tabulate the relevant ionization and heating quantities for a given radiation source, effectively allowing us to simulate sources with continuous SEDs.
+Given our motives, this is clearly a necessary requirement.
+Pre-tabulating key quantities also grants a significant speed-up for calculations that are already cheap, but most importantly, provides the basis for our optimization strategy that will be introduced in the next section.
+
+In the remainder of this section, we will focus on two simple test problems.
+First, the standard case of ionization and heating surrounding a $10^5$ K blackbody (test problem 2 in RT06), and second, the identical scenario for a power-law emitting X-ray source.
+Initially, we will consider each case in a hydrogen-only medium, and include the effects of helium shortly thereafter.
+
+\subsection{$10^5$ K Blackbody}
+
+\subsection{Power-Law X-ray Source}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[scale=0.4]{figures/RT_Test1_RadialProfiles.ps}
+\caption{this is a test.}
+\label{fig:test}
+\end{center}
+\end{figure}
+
+\section{METHODS}
+Upon careful inspection of the coupled rate equations which govern the balance of ionizations/recombinations and heating/cooling, one realizes that the integral values, i.e.
+equations \ref{HIIRateEquation}, \ref{HeIIRateEquation}, \ref{HeIIIRateEquation}, and \ref{TemperatureEvolution} can be tabulated as a function of column density (i.e.
+optical depth), allowing for a significant speed-up in run time.
 The key to our subsequent analysis was inspired by this same convenient feature: if $\Gamma_i$ and $\Heat_i$ can be tabulated for a given source \textit{a-priori}, it means that \textit{precision determination of the ionization state and temperature of the gas surrounding that source relies solely on accurate numerical computation of the photoionization and heating rate integrals in Equations \ref{PhotoionizationRate} and \ref{HeatingRate}}.
@@
 \end{equation}
-
-% Code Verification
-%\subsection{Code Verification}
-%Our code, \texttt{rt1d}, solves Equations \ref{HIIRateEquation}, \ref{HeIIRateEquation}, \ref{HeIIIRateEquation}, and \ref{TemperatureEvolution} using the implicit Euler method for integration and the Newton-Raphson technique for root finding.
-In order to track the propagation of ionization fronts accurately, we limit the time-step based on a maximum neutral fraction change as introduced in \cite{Shapiro2004},
-%\begin{equation}
-%    \Delta t = \epsilon_{\mathrm{ion}} \frac{\nHI}{|d\nHI/dt|} ,
-%\end{equation}
-%and additionally require that the time-step increase by a factor of two at most, as in \cite{Wise2011}.  
-%
-%To demonstrate the functionality of the code, we repeat tests 1 and 2 from the Radiative Transfer Comparison Project (\cite{Iliev2006}; hereafter referred to as RT06).
-Test 1 is the expansion of an HII region in a hydrogen-only, isothermal medium surrounding a monochromatic source of 13.6 eV photons.
-Test 2 is the same, except for the temperature is allowed to evolve according to Equation \ref{TemperatureEvolution}, and the monochromatic radiation source is replaced by a $10^5$ K blackbody spectrum.
-Results for Test 1 are shown in Figures \ref{fig:RT_Test1_IfrontEvolution} and \ref{fig:RT_Test1_RadialProfiles}, and for Test 2 in Figure \ref{fig:RT_Test2_RadialProfiles}.
-%\begin{figure*}
-%\begin{center}
-%\leavevmode
-%\psfig{file=figures/RT_Test1_IfrontEvolution.ps,width=6in}
-%\vspace{-20pt}
-%\caption[]{\textit{Top:} Comparison of the calculated (dashed) and analytic (solid) positions of an expanding ionization front in a hydrogen-only, isothermal medium.
-\textit{Bottom:} Ratio of the calculated and analytic solutions.}
-%\label{fig:RT_Test1_IfrontEvolution}
-%\end{center}
-%\end{figure*}
-%
-%\begin{figure*}
-%\begin{center}
-%\leavevmode
-%\psfig{file=figures/RT_Test1_RadialProfiles.ps,width=6in}
-%\vspace{-20pt}
-%\caption[]{Radial profiles of the neutral (solid) and ionized (dashed) fractions at $t = 10$, $30$, $100$, and $500$ Myr.}
-%\label{fig:RT_Test1_RadialProfiles}
-%\end{center}
-%\end{figure*}
-%
-%\begin{figure*}
-%\begin{center}
-%\leavevmode
-%\psfig{file=figures/RT_Test2_RadialProfiles.ps,width=6in}
-%\vspace{-20pt}
-%\caption[]{\textit{Top:} Radial profiles of the neutral (solid) and ionized (dashed) fractions at $t = 10$, $100$, and $500$ Myr.
-\textit{Bottom:} Radial profiles of the kinetic temperature at $t = 10$, $100$, and $500$ Myr (solid, dashed, and dotted lines, respectively).}
-%\label{fig:RT_Test2_RadialProfiles}
-%\end{center}
-%\end{figure*}
 
@@
 
-\section{Results}
+\section{RESULTS}
 
-\section{Why not tabulate? (discussion)}
+\section{DISCUSSION}
 
@@
 
-\section{Conclusions}
+\section{CONCLUSIONS}
 
@@
 
+%%% BEGIN APPENDIX
+\appendix
+\twocolumn
+\appendixpage
+
 %%% 
-%% APPENDIX
+%% RT Framework
 %%%
-\appendix
-
-% RT Framework
-\section{A. Radiative Transfer Framework}
+\section{RADIATIVE TRANSFER FRAMEWORK}
 
@@
 
-% Code Verification
-\section{B. Code Verification}
+%%% 
+%% Code Verification
+%%%
+\section{CODE VERIFICATION}
 Our code, \texttt{rt1d}, solves Equations \ref{HIIRateEquation}, \ref{HeIIRateEquation}, \ref{HeIIIRateEquation}, and \ref{TemperatureEvolution} using the implicit Euler method for integration and the Newton-Raphson technique for root finding.
@@
 Test 1 is the expansion of an HII region in a hydrogen-only, isothermal medium surrounding a monochromatic source of 13.6 eV photons.
-Test 2 is the same, except for the temperature is allowed to evolve according to Equation \ref{TemperatureEvolution}, and the monochromatic radiation source is replaced by a $10^5$ K blackbody spectrum.
-Results for Test 1 are shown in Figures \ref{fig:RT_Test1_IfrontEvolution} and \ref{fig:RT_Test1_RadialProfiles}, and for Test 2 in Figure \ref{fig:RT_Test2_RadialProfiles}.
-\begin{figure*}
+
+\begin{figure}[htbp]
 \begin{center}
-\leavevmode
-\psfig{file=figures/RT_Test1_IfrontEvolution.ps,width=6in}
+\includegraphics[scale=0.4]{figures/RT_Test1_IfrontEvolution.ps}
 \vspace{-20pt}
-\caption[]{\textit{Top:} Comparison of the calculated (dashed) and analytic (solid) positions of an expanding ionization front in a hydrogen-only, isothermal medium.
+\caption{\textit{Top:} Comparison of the calculated (dashed) and analytic (solid) positions of an expanding ionization front in a hydrogen-only, isothermal medium.
 \textit{Bottom:} Ratio of the calculated and analytic solutions.}
@@
 \end{center}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
+\begin{figure}[htbp]
 \begin{center}
-\leavevmode
-\psfig{file=figures/RT_Test1_RadialProfiles.ps,width=6in}
+\includegraphics[scale=0.4]{figures/RT_Test1_RadialProfiles.ps}
 \vspace{-20pt}
@@
 \end{center}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
+Test 2 is the same, except for the temperature is allowed to evolve according to Equation \ref{TemperatureEvolution}, and the monochromatic radiation source is replaced by a $10^5$ K blackbody spectrum.
+Results for Test 1 are shown in Figures \ref{fig:RT_Test1_IfrontEvolution} and \ref{fig:RT_Test1_RadialProfiles}, and for Test 2 in Figure \ref{fig:RT_Test2_RadialProfiles}.
+
+\begin{figure}[htbp]
 \begin{center}
-\leavevmode
-\psfig{file=figures/RT_Test2_RadialProfiles.ps,width=6in}
+\includegraphics[scale=0.4]{figures/RT_Test2_RadialProfiles.ps}
 \vspace{-20pt}
-\caption[]{\textit{Top:} Radial profiles of the neutral (solid) and ionized (dashed) fractions at $t = 10$, $100$, and $500$ Myr.
+\caption{\textit{Top:} Radial profiles of the neutral (solid) and ionized (dashed) fractions at $t = 10$, $100$, and $500$ Myr.
 \textit{Bottom:} Radial profiles of the kinetic temperature at $t = 10$, $100$, and $500$ Myr (solid, dashed, and dotted lines, respectively).}
@@
 \end{center}
-\end{figure*}
+\end{figure}
 

### 2011-09-01 rt1d/mods/WriteData.py

@@ -41,7 +41,9 @@ class WriteData:
             if par == "CurrentTime": pf_grp.create_dataset(par, data = t / self.TimeUnits)
             elif par == "CurrentTimestep": pf_grp.create_dataset(par, data = dt / self.TimeUnits)
             else: pf_grp.create_dataset(par, data = self.pf[par])
-        for field in data: data_grp.create_dataset(field, data = data[field])
+        for field in data: 
+            if len(data[field]) == 0: continue
+            data_grp.create_dataset(field, data = data[field])
         
         f.close()
         

### 2011-09-01 scripts/rt1d.py

@@ -164,6 +164,6 @@ for i, pf in enumerate(all_pfs):
         fdt.close()
                 
     elapsed = time.time() - start    
-    print "Calculation {0} complete (output to {1}).  Elapsed time = {2} seconds.".format(i + 1, pf["OutputDirectory"], int(elapsed))
+    print "Calculation {0} complete (output to {1}).  Elapsed time = {2} hours.".format(i + 1, pf["OutputDirectory"], round(elapsed / 3600., 2))
 
 print "Successful run. Exiting."

### 2011-09-03 doc/paper/ms.tex
Everything is looking good.  Time to run big resolution suite.

@@
 In the remainder of this section, we will focus on two simple test problems.
-First, the standard case of ionization and heating surrounding a $10^5$ K blackbody (test problem 2 in RT06), and second, the identical scenario for a power-law emitting X-ray source.
-Initially, we will consider each case in a hydrogen-only medium, and include the effects of helium shortly thereafter.
+First, the standard case of ionization and heating surrounding a $10^5$ K blackbody (problem \#2 in RT06), and second, the identical scenario for a power-law emitting X-ray source.
+Initially, we will consider each case in a hydrogen-only medium, and later include the effects of helium.
 
@@
 \begin{figure}[htbp]
-\begin{center}
+\centering
 \includegraphics[scale=0.4]{figures/RT_Test1_IfrontEvolution.ps}
@@
 \label{fig:RT_Test1_IfrontEvolution}
-\end{center}
+%\end{center}
 \end{figure}

### 2011-09-03 rt1d/analysis/Dataset.py

@@ -73,71 +73,5 @@ class Dataset:
             f.close()
             
         return ds
-            
-    #def get_energy_deposition(self, dataset, pf, bin_size = 100, bin_spacing = 'lin'):
-    #    """
-    #    Return a dictionary where each entry is a 2D array representing the energy deposition rate (erg / s / cm^3)
-    #    deposition by photons in each bin of the spectrum in a given channel (heat, ionization, etc.).
-    #    
-    #        bin_size (eV): size of energy bins for the spectrum
-    #        dataset: only single time dump, i.e. sim[50]
-    #        
-    #        currently only doing photoionization
-    #        dataset must contain column density information
-    #    """
-    #    
-    #    t = dataset.t
-    #    r = dataset.r
-    #    rs = rm.RadiationSource(pf)
-    #    
-    #    n_bins = (pf["SpectrumMaxEnergy"] - pf["SpectrumMinEnergy"]) / bin_size + 1
-    #    if bin_spacing == 'lin':
-    #        bin_edges = np.linspace(pf["SpectrumMinEnergy"], pf["SpectrumMaxEnergy"], n_bins)
-    #    elif bin_spacing == 'log':
-    #        bin_edges = np.logspace(pf["SpectrumMinEnergy"], pf["SpectrumMaxEnergy"], n_bins)
-    #    else:
-    #        print 'Bin spacing not recognized!'
-    #        return 0
-    #     
-    #    result = np.zeros([pf["GridDimensions"] - 1, n_bins], float) 
-    #        
-    #    n = zip(*[dataset.ncol_HI, dataset.ncol_HeI, dataset.ncol_HeII])   
-    #    
-    #    for i in np.arange(int(pf["GridDimensions"]) - 1):    
-    #        for j, bin in enumerate(bin_edges):
-    #            if bin == bin_edges[-1]: continue
-    #            integrand = lambda E: PhotoIonizationCrossSection(E, 0) * rs.Spectrum(E) * \
-    #                np.exp(-self.tau(E, n[i])) / (E * erg_per_ev)
-    #            result[i][j] = rs.BolometricLuminosity(t) * integrate(integrand, bin, bin + bin_size)[0] \
-    #                / 4. / np.pi / r[i]**2
-    #                            
-    #    return {'PhotoIonization': result}
-    #                                
-    #def tau(self, E, n):
-    #    """
-    #    Returns the optical depth at energy E due to column densities of HI, HeI, and HeII, which
[…]

### 2011-09-03 rt1d/analysis/__init__.py

@@ -1,3 +1,3 @@
 from rt1d.analysis.Dataset import Dataset 
-from rt1d.analysis.Analysis import Analysis
+from rt1d.analysis.Analysis import Analyze
  

### 2011-09-03 rt1d/mods/Radiate.py

@@ -79,7 +79,7 @@ class Radiate:
         self.AdaptiveTimestep = pf["ODEAdaptiveStep"]
         self.CosmologicalExpansion = pf["CosmologicalExpansion"]
         self.InitialHIIFraction = pf["InitialHIIFraction"]
-        self.GridDimensions = pf["GridDimensions"]
+        self.GridDimensions = int(pf["GridDimensions"])
         self.InitialRedshift = pf["InitialRedshift"]
         self.LengthUnits = pf["LengthUnits"]
         self.TimeUnits = pf["TimeUnits"]
@@ -97,7 +97,6 @@ class Radiate:
         self.HeIIColumn = n_col[2]
         
         self.OnePhotonPackagePerCell = pf["OnePhotonPackagePerCell"]
-        self.RadiativeTransferCourantCondition = pf["RadiativeTransferCourantCondition"]        
         self.LightCrossingTimeRestrictedStep = pf["LightCrossingTimeRestrictedStep"]
         self.AdaptiveStep = pf["ODEAdaptiveStep"]
         self.MaxStep = pf["ODEMaxStep"] * self.TimeUnits
@@ -263,7 +262,7 @@ class Radiate:
             else: packs.append(np.array([t, dt, 0., 0., 0., Lbol * dt]))
         
         if self.HIIRestrictedTimestep: dtphot = 1e50 * np.ones_like(self.grid)
-        
+                
         if self.InfiniteSpeedOfLight:
             
             # Loop over cells radially, solve rate equations, update values in data -> newdata
@@ -318,14 +317,20 @@ class Radiate:
             
                 # Unpack results of coupled equations - remember, these are lists and we only need the last entry 
                 newHII, newHeII, newHeIII, newE = qnew
-                                        
+                
+                # Ensure monotonicity in ionized species
+                #if cell > self.StartCell:
+                #    newHII[-1] = self.MonotonicityFudge(newdata['HIIDensity'][cell - 1], newHII[-1])  
+                #    newHeII[-1] = self.MonotonicityFudge(newdata['HeIIDensity'][cell - 1], newHeII[-1])
+                #    newHeIII[-1] = self.MonotonicityFudge(newdata['HeIIIDensity'][cell - 1], newHeIII[-1])
+                
                 # Convert from internal energy back to temperature
                 if not self.Isothermal: newT = newE[-1] * 2. * mu / 3. / k_B / n_B
                 else: newT = newdata['Temperature'][cell]
                 
[…]

### 2011-09-03 scripts/AnalyzeTest1.py

@@ -46,7 +46,12 @@ grid = np.arange(GridDims)
 r = []
 t = []
 mean_xH = []
-for dd in ds.data.keys()[1:]:
+for dd in ds.data.keys():
+    
+    if ds.data[dd].t == 0:
+        r.append(0)
+        t.append(0)
+    
     x_H = ds.data[dd].x_HI
     
     mean_xH.append(np.mean(x_H))
@@ -69,7 +74,7 @@ r = np.array(r)
 t = np.array(t)
 
 func = lambda t: rs * (1. - np.exp(-t / trec))**(1. / 3.) + StartRadius
-t_anl = np.linspace(0, max(t), 500)
+t_anl = np.linspace(0, 500, 500)
 r_anl = map(func, t_anl)
 r_anl_bin = map(func, t)
         
@@ -77,15 +82,19 @@ mp = multiplot(dims = (2, 1), panel_size = (0.5, 1))
 
 mp.axes[0].plot(t / trec, r, color = 'k', ls = '--')
 mp.axes[0].plot(t_anl / trec, r_anl, linestyle = '-', color = 'k')
-mp.axes[0].set_xlim(0, 1 * max(t/trec))
+mp.axes[0].set_xlim(0, max(t / trec))
 mp.axes[0].set_ylim(0, 1.1 * max(max(r), max(r_anl)))
 mp.axes[0].set_ylabel(r'$r \ (\mathrm{kpc})$')  
 
 mp.axes[1].plot(t / trec, r / r_anl_bin, ls = '-', color = 'k')
-mp.axes[1].set_xlim(0, 1 * max(t/trec))
+mp.axes[1].set_xlim(0, max(t / trec))
 mp.axes[1].set_ylim(0.9, 1.1)
 mp.axes[1].set_xlabel(r'$t / t_{\mathrm{rec}}$')
 mp.axes[1].set_ylabel(r'$r/r_{\mathrm{anl}}$') 
+
+mp.axes[0].xaxis.set_ticks(np.linspace(0, 4, 5))
+mp.axes[1].xaxis.set_ticks(np.linspace(0, 4, 5))
+
 mp.fix_ticks()
 pl.savefig('{0}/RT_Test1_IfrontEvolution.ps'.format(OutputDirectory))
 pl.savefig('{0}/RT_Test1_IfrontEvolution.png'.format(OutputDirectory))
@@ -93,7 +102,7 @@ pl.clf()
 
 # Write out data
[…]

### 2011-09-07 doc/paper/ms.tex
Added super high-res parameter files for RT06 #1.
Free of scipy

@@
 %
-
-%---------------------------------------- Declarations -----------------------------------------%
-
-%\documentclass{emulateapj}
-%\documentclass[12pt,preprint]{aastex}
-%\usepackage{emulateapj5}
-%\usepackage{apjfonts}
-%
-%% Define General Packages
-%\usepackage{graphicx}
-%\usepackage{pslatex}
-%\usepackage{amsmath, amsthm, amssymb}
-%\usepackage{natbib}
-%\usepackage{epsfig}
-%
-%% Use utf-8 encoding for foreign characters
-%\usepackage[utf8]{inputenc}
 
@@
 \title {Optimized Multi-Frequency Spectra for Numerical Radiative Transfer}
-\author{Jordan Mirocha, Stephen Skory, Jack Burns}
+\author{Jordan Mirocha$^{\dagger}$, Stephen Skory, Jack Burns}
 \affil{Center for Astrophysics and Space Astronomy, University of Colorado, Campus Box 389, Boulder, CO 80309}
 \affil{NASA Lunar Science Institute, NASA Ames Research Center, Moffett Field, CA 94043, USA}
+\email{$^{\dagger}$jordan.mirocha@colorado.edu}
 \author{John Wise}
@@
 Similarly, radiation-hydrodynamic simulations may be slowed by the inclusion of photons that heat gas efficiently, since higher temperature cells have greater sound speeds, and thus require a smaller time-step.
+
 Expense due to additional energy groups is certainly method-dependent, and may even be implementation-dependent.
@@
 In \S5, we conclude.
-Details concerning the code used for this work, as well as a discussion regarding the factors that influence the solution of radiative transfer calculations, whether it be the temporal resolution, spatial resolution, frequency resolution, or the interplay of all three, can be found in the Appendix.
-
-%Reduce guesswork!  Choose frequency resolution in the same way that one chooses spatial/mass resolution.
+Verification of the code used for this work can be found in the Appendix.
 

### 2011-09-07 rt1d/mods/RadiationSource.py

@@ -13,8 +13,7 @@ Notes:
 """
 
 import numpy as np
-from scipy.integrate import quad
-from scipy.optimize import fsolve
+from Integrate import simpson as integrate
 
 h = 6.626068 * 10**-27 			                            # Planck's constant - [h] = erg*s
 k_B = 1.3806503 * 10**-16			                        # Boltzmann's constant - [k_B] = erg/K
@@ -117,7 +116,7 @@ class RadiationSource:
         else:
         
             if self.SourceType == 1 or self.SourceType == 2:
-                integral, err = quad(self.SpecificIntensity, 0, np.inf)
+                integral, err = integrate(self.SpecificIntensity, 0, np.inf)
                 
             if self.SourceType == 3:
                 if self.alpha == -1.0: 
@@ -142,8 +141,8 @@ class RadiationSource:
             return self.Lph / (np.sum(self.F / self.E / erg_per_ev))
             
         if self.SourceType == 1:
-            norm = quad(self.SpecificIntensity, 0, np.inf)[0]
-            return self.Lph / quad(lambda E: self.SpecificIntensity(E) / norm / E / erg_per_ev, 0, np.inf)[0]
+            norm = integrate(self.SpecificIntensity, 0, np.inf)[0]
+            return self.Lph / integrate(lambda E: self.SpecificIntensity(E) / norm / E / erg_per_ev, 0, np.inf)[0]
         
         if self.SourceType == 2:
             return 10**SchaererTable["Luminosity"][SchaererTable["Mass"].index(self.M)] * lsun
@@ -158,7 +157,7 @@ class RadiationSource:
         Returns cumulative energy output contributed by photons at or less than energy E.
         """    
         
-        return quad(self.Spectrum, 0, E)[0]        
+        return integrate(self.Spectrum, 0, E)[0]        
             
             
             

### 2011-09-07 rt1d/mods/ReadParameterFile.py

@@ -97,7 +97,7 @@ def ProblemType(pt):
              "ColumnDensityBinsHI": 500, "GridDimensions": 100, "LengthUnits": 6.6 * cm_per_kpc, \
              "TimeUnits": s_per_myr, "CurrentTime": 0.0, "StopTime": 500.0, \
              "StartRadius": 0.01, "dtDataDump": 5.0, "DataDumpName": 'dd', \
-             "Isothermal": 0, "MultiSpecies": 0, "SecondaryIonization": 0, "CosmologicalExpansion": 0, \
+             "Isothermal": 0, "MultiSpecies": 0, "SecondaryIonization": 1, "CosmologicalExpansion": 0, \
              "DensityProfile": 0, "InitialDensity": 1e-3, "TemperatureProfile": 0, "InitialTemperature": 1e2, \
              "IonizationProfile": 1, "InitialHIIFraction": 0, "SourceType": 0, "SourceLifetime": 1e10, \
              "SpectrumPhotonLuminosity": 5e48, "DiscreteSpectrum": 1, \
@@ -110,7 +110,7 @@ def ProblemType(pt):
              "ColumnDensityBinsHI": 100, "GridDimensions": 100, "LengthUnits": 6.6 * cm_per_kpc, \
              "TimeUnits": s_per_myr, "CurrentTime": 0.0, "StopTime": 500.0, \
              "StartRadius": 0.01, "dtDataDump": 5.0, "DataDumpName": 'dd', \
-             "Isothermal": 0, "MultiSpecies": 0, "SecondaryIonization": 0, "CosmologicalExpansion": 0, \
+             "Isothermal": 0, "MultiSpecies": 0, "SecondaryIonization": 1, "CosmologicalExpansion": 0, \
              "DensityProfile": 0, "InitialDensity": 1e-3, "TemperatureProfile": 0, "InitialTemperature": 1e2, \
              "IonizationProfile": 1, "InitialHIIFraction": 0, "SourceType": 1, "SourceLifetime": 1e10, \
              "SpectrumPhotonLuminosity": 5e48, "SpectrumMinEnergy": 13.6, "SpectrumMaxEnergy": 100., \

### 2011-09-07 rt1d/mods/SolveRateEquations.py

@@ -61,10 +61,8 @@ class SolveRateEquations:
         """
         This routine does all the work.
         
-            f: function(y, x, *args) to be solved.  Order of y and x is to be consistent with scipy convention.
-            
+            f: function(y, x, *args) to be solved.
             x0, y0: Initial conditions on x and y.  y0 may be a tuple if equations are coupled.
-            
             xf: Endpoint of integration in independent variable x.
             
         """

### 2011-10-24 doc/paper/ms.tex
Added Multiplot.py, updated a few scripts to use it.

@@
 As a result, we should expect the properties of gas surrounding a blackbody source of ultraviolet (UV) photons to be quite different from those of a power-law X-ray source such as an accreting black hole, for example.
-Even relatively small changes in SED can produce noticeable differences in surrounding gas properties, like changing the power-law index of a black-hole accretion spectrum (\cite{Kuhlen2005}; \cite{Thomas2008}).
+Holding the bolometric luminosity of an object constant, even relatively small changes in SED can produce noticeable differences in surrounding gas properties, like changing the power-law index of a black-hole accretion spectrum (\cite{Kuhlen2005}; \cite{Thomas2008}).
 But, these expectations are built on the assumption that each source is represented by a continuous SED, not one we have constructed with \nnu discrete frequency bins.
@@
 Test 1 is the expansion of an HII region in a hydrogen-only, isothermal medium surrounding a monochromatic source of 13.6 eV photons.
+We adopt the parameters used in \citet{Wise2011}: constant temperature $T = 10^4 \ \mathrm{K}$, hydrogen number density $n_H = 10^{-3} \ \mathrm{cm^{-3}}$, ionized fraction $x = 1.2 \times 10^-3$, in a box $L_{\mathrm{box}} = 6.6 \ \mathrm{kpc}$ in size, and with photon luminosity $\dot{Q} = 5\times10^{48} \ \mathrm{s^{-1}}$.
 
-\begin{figure}[htbp]
+\begin{figure*}[htbp]
 \centering
-\includegraphics[scale=0.4]{figures/RT_Test1_IfrontEvolution.ps}
-\vspace{-20pt}
-\caption{\textit{Top:} Comparison of the calculated (dashed) and analytic (solid) positions of an expanding ionization front in a hydrogen-only, isothermal medium.
-\textit{Bottom:} Ratio of the calculated and analytic solutions.}
+\includegraphics[scale=0.8]{figures/RT06_1_IfrontEvolution.eps}
+\caption{\textit{Top:} Comparison of the numerical (dashed) and analytic (solid) solutions for the position of an expanding ionization front as a function of time in a hydrogen-only, isothermal medium (RT06 problem \# 1).
+The numerical solution here was obtained using 400 grid cells.
+\textit{Bottom:} Ratio of the calculated and analytic solutions as a function of time and grid resolution.
+}
 \label{fig:RT_Test1_IfrontEvolution}
-%\end{center}
-\end{figure}
+\end{figure*}
 

### 2011-10-24 rt1d/analysis/Multiplot.py

@@ -1,0 +1,143 @@
+"""
+Multiplot.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on 2011-06-27.
+
+Description: Make multipanel plots with shared axes, with or without AxesGrid.
+     
+"""
+
+import pylab as pl
+import numpy as np
+from mpl_toolkits.axes_grid1 import AxesGrid, make_axes_locatable
+
+# Matplotlibrc defaults
+figsize = pl.rcParams['figure.figsize']
+wspace = pl.rcParams['figure.subplot.wspace']
+hspace = pl.rcParams['figure.subplot.hspace']
+left = pl.rcParams['figure.subplot.left']
+right = pl.rcParams['figure.subplot.right']
+bottom = pl.rcParams['figure.subplot.bottom']
+top = pl.rcParams['figure.subplot.top']    
+
+def AxisConstructor(nr, nc, panel): 
+    return pl.subplot(nr, nc, panel)
+
+class multiplot:
+    def __init__(self, dims = (2, 2), padding = 0, panel_size = None, useAxesGrid = True,
+        share_all = True, aspect = False):
+        """
+        dims = (rows, columns)
+        padding = separation of subplots in inches
+        panel_size = fraction of default figsize in each dimension
+        nd = dimensionality of data
+        """
+        
+        self.dims = dims
+        self.padding = padding
+        
+        if panel_size is None: self.panel_size = np.array([figsize[0], figsize[1]])
+        else: self.panel_size = np.array([panel_size[0] * figsize[0], panel_size[1] * figsize[1]])
+        
+        if share_all and not useAxesGrid:
+            pl.rcParams['figure.subplot.wspace'] = self.padding
+            pl.rcParams['figure.subplot.hspace'] = self.padding        
+                
+        self.fig = pl.figure(1, self.panel_size)
+        
+        self.Naxes = np.prod(self.dims)
+        self.elements = np.reshape(xrange(self.Naxes), (self.dims[0], self.dims[1]))
+        self.xaxes = self.elements[-1]
+        self.yaxes = zip(*self.elements)[0]                  
+        self.lowerleft = self.elements[-1][0]
+        self.lowerright = self.elements[-1][-1]
+        self.upperleft = self.elements[0][0]
+        self.upperright = self.elements[0][-1]
+                
+        # Set up grid
+        if useAxesGrid:        
+            self.grid = AxesGrid(self.fig, 111, nrows_ncols = self.dims, axes_pad = self.padding, 
[…]

### 2011-10-24 rt1d/mods/InitializeIntegralTables.py

@@ -290,7 +290,7 @@ class InitializeIntegralTables:
             integrand = lambda E: PhotoIonizationCrossSection(E, species) * self.rs.Spectrum(E) * \
                 np.exp(-self.OpticalDepth(E, n)) / (E * erg_per_ev)    
                         
-            return integrate(integrand, max(E_th[species], self.rs.Emin), self.rs.Emax, tol = 1e-16)
+            return integrate(integrand, max(E_th[species], self.rs.Emin), self.rs.Emax, tol = 1e-24)
                   
         else:
             integral = PhotoIonizationCrossSection(self.rs.E, species) * self.rs.Spectrum(self.rs.E) * \
@@ -313,7 +313,7 @@ class InitializeIntegralTables:
             integrand = lambda E: PhotoIonizationCrossSection(E, species) * (E - E_th[species]) * self.rs.Spectrum(E) * \
                 np.exp(-self.OpticalDepth(E, n)) / E
             
-            return integrate(integrand, max(E_th[species], self.rs.Emin), self.SpectrumMaxEnergy, tol = 1e-24)
+            return integrate(integrand, max(E_th[species], self.rs.Emin), self.SpectrumMaxEnergy, tol = 1e-30)
             
         else:
             integral = PhotoIonizationCrossSection(self.rs.E, species) * (self.rs.E - E_th[species]) * \

### 2011-10-24 rt1d/mods/SetDefaultParameterValues.py

@@ -33,7 +33,7 @@ def SetDefaultParameterValues():
         "ODEMaxStep": 1e-1, \
         "ODErtol": 1e-3, \
         "ODEatol": 1e-6, \
-        "ODEmaxiter": 10000, \
+        "ODEmaxiter": 1e4, \
         "RootFinder": 0, \
         
         "ParallelizationMethod": 2, \

### 2011-10-24 scripts/AnalyzeTest2.py

@@ -17,7 +17,7 @@ import rt1d.analysis as rta
 from multiplot import *
 from constants import *
 
-ds = rta.Dataset(sys.argv[1])
+ds = rta.Analyze(sys.argv[1])
 
 mp = multiplot(dims = (2, 1), panel_size = (0.5, 1))
 

### 2011-10-24 scripts/AnalyzeTest1_Resolution.py
Added resolution suite analysis script to /scripts. Removed old ones.

@@ -1,66 +1,0 @@
-"""
-AnalyzeTest1_Resolution.py
-
-Author: Jordan Mirocha
-Affiliation: University of Colorado at Boulder
-Created on 2011-07-02.
-
-Description: Analyze results of RT06 test problem #1 run with differing spatial resolutions.
-
-Notes: Supply parameter file as commmand line argument.     
-     
-"""
-
-import misc
-import pylab as pl
-
-r, x1, x2, x3, x4 = misc.readtab('c100/RT_Test1_RadialProfiles.dat')
-pl.semilogy(r, x1, ls = '-', color = 'k', label = r'$n_{\mathrm{c}} = 100$')
-pl.semilogy(r, x3, ls = '-', color = 'k')
-pl.semilogy(r, x4, ls = '-', color = 'k')
-r, x1, x2, x3, x4 = misc.readtab('c200/RT_Test1_RadialProfiles.dat')
-pl.semilogy(r, x1, ls = '--', color = 'blue', label = r'$n_{\mathrm{c}} = 200$')
-pl.semilogy(r, x3, ls = '--', color = 'blue')
-pl.semilogy(r, x4, ls = '--', color = 'blue')
-r, x1, x2, x3, x4 = misc.readtab('c400/RT_Test1_RadialProfiles.dat')
-pl.semilogy(r, x1, ls = ':', color = 'red', label = r'$n_{\mathrm{c}} = 400$')
-pl.semilogy(r, x3, ls = ':', color = 'red')
-pl.semilogy(r, x4, ls = ':', color = 'red')
-r, x1, x2, x3, x4 = misc.readtab('c800/RT_Test1_RadialProfiles.dat')
-pl.semilogy(r, x1, ls = '-.', color = 'green', label = r'$n_{\mathrm{c}} = 800$')
-pl.semilogy(r, x3, ls = '-.', color = 'green')
-pl.semilogy(r, x4, ls = '-.', color = 'green')
-r, x1, x2, x3, x4 = misc.readtab('c1600/RT_Test1_RadialProfiles.dat')
-pl.semilogy(r, x1, ls = '-', color = 'magenta', label = r'$n_{\mathrm{c}} = 1600$')
-pl.semilogy(r, x3, ls = '-', color = 'magenta')
-pl.semilogy(r, x4, ls = '-', color = 'magenta')
[…]

### 2011-10-24 scripts/AnalyzeTest2.py

@@ -13,34 +13,49 @@ Notes: Supply parameter file as commmand line argument.  Assumes data dumps are 
 """
 
 import sys, misc
+import pylab as pl
+import numpy as np
 import rt1d.analysis as rta
-from multiplot import *
-from constants import *
+
+cm_per_kpc = 3.08568e21
+s_per_myr = 365.25 * 24 * 3600 * 1e6
 
 ds = rta.Analyze(sys.argv[1])
 
-mp = multiplot(dims = (2, 1), panel_size = (0.5, 1))
+mp = rta.multiplot(dims = (2, 1), panel_size = (1, 2), useAxesGrid = False)
 
-mp.axes[0].semilogy(ds.data[2].r / cm_per_kpc / 6.6, ds.data[2].x_HI, color = 'k', ls = '-', label = r'$1 - x_i$')
-mp.axes[0].semilogy(ds.data[2].r / cm_per_kpc / 6.6, ds.data[2].x_HII, color = 'k', ls = '--', label = r'$x_i$')
-mp.axes[0].semilogy(ds.data[20].r / cm_per_kpc / 6.6, ds.data[20].x_HI, color = 'k', ls = '-')
-mp.axes[0].semilogy(ds.data[20].r / cm_per_kpc / 6.6, ds.data[20].x_HII, color = 'k', ls = '--')
-mp.axes[0].semilogy(ds.data[100].r / cm_per_kpc / 6.6, ds.data[100].x_HI, color = 'k', ls = '-')
-mp.axes[0].semilogy(ds.data[100].r / cm_per_kpc / 6.6, ds.data[100].x_HII, color = 'k', ls = '--')
-mp.axes[1].semilogy(ds.data[2].r / cm_per_kpc / 6.6, ds.data[2].T, color = 'k', ls = '-')
-mp.axes[1].semilogy(ds.data[20].r / cm_per_kpc / 6.6, ds.data[20].T, color = 'k', ls = '--')
-mp.axes[1].semilogy(ds.data[100].r / cm_per_kpc / 6.6, ds.data[100].T, color = 'k', ls = ':')
+mp.grid[0].semilogy(ds.data[2].r / cm_per_kpc / 6.6, ds.data[2].x_HI, color = 'k', ls = '-', label = r'$1 - x_i$')
[…]

### 2011-10-24 doc/manual/rt1d_manual.tex
Moved man to manual .

--- 2011-01-22 doc/man/rt1d_manual.tex
@@
 %%%
+%% Introduction
+%%%
+\section{Introduction}
+
+%%% 
 %% Methods
@@
 \subsection{\texttt{Radiate}}
-
-%%% 
-%% Start to Finish
-%%%
-\section{Start to Finish}
 
@@
 
+\section{Analysis}
+Since our datasets are small, we can read in the entire time evolution all at once.  To do this, type the following in a Python terminal:
+\begin{verbatim}
+import rt1d.analysis as rta
+import pylab as pl
+
+ds = rta.Analyze('./pf.dat')                # load parameter file and data
+pl.loglog(ds.data[10].r, ds.data[10].x_HI)  # radius vs. neutral fraction
+\end{verbatim}
+
+%%% 
+%% Start to Finish
+%%%
+\section{Start to Finish}
+
+\section{Convergence}
+In \texttt{rt1d/doc/examples} you will find directories containing convergence test suites for RT06 test problems 1 \& 2 (\texttt{RT06\_1\_ResolutionTestSuite} and \texttt{RT06\_2\_ResolutionTestSuite}).
+Each of these directories contains the parameter files necessary to run these problems on grids containing $100 \times 2^n$ cells, where $n = 0,1,...6$.
+If you want, you can run them all at once, though keep in mind the simulations with 100 cells will finish much much faster than those containing 6400 cells, so it might not be good to check out a node for this as most of the computation will be spent on 1 or 2 simulations.
+For instance, you could do:
+\begin{verbatim}
+    cd rt1d/doc/examples/RT06_1_ResolutionTestSuite/c_infinite
+    mpirun -np 4 python RT1D.py -b batch_list.dat
+\end{verbatim}
+I've included the same setup for $c \neq \infty$ in \texttt{rt1d/doc/examples/RT06\_1\_ResolutionTestSuite/c\_finite}, so you can see the differences (if any) caused by making the $c \rightarrow \infty$ approximation.
+
 \end{document}

### 2011-10-25 doc/paper/ms.tex
Added resolution suite analysis script to /scripts. Removed old ones.
Added a lot to appendix, sign errors from Thomas and Zaroubi fixed in code and paper.

@@
 \newcommand{\aHeIII}{\alpha_{\text{HeIII}}}
+\newcommand{\bHI}{\beta_{\text{HI}}} 
 \newcommand{\bHeI}{\beta_{\text{HeI}}}  
@@
 \newcommand{\Heat}{\mathcal{H}}
+\newcommand{\trec}{$t_{\text{rec}}$}
 %%% END custom commands
@@
 \section{METHODS}
-Upon careful inspection of the coupled rate equations which govern the balance of ionizations/recombinations and heating/cooling, one realizes that the integral values, i.e.
-equations \ref{HIIRateEquation}, \ref{HeIIRateEquation}, \ref{HeIIIRateEquation}, and \ref{TemperatureEvolution} can be tabulated as a function of column density (i.e.
-optical depth), allowing for a significant speed-up in run time.
-The key to our subsequent analysis was inspired by this same convenient feature: if $\Gamma_i$ and $\Heat_i$ can be tabulated for a given source \textit{a-priori}, it means that \textit{precision determination of the ionization state and temperature of the gas surrounding that source relies solely on accurate numerical computation of the photoionization and heating rate integrals in Equations \ref{PhotoionizationRate} and \ref{HeatingRate}}.
-So, for a given source, we can compute the values of these integrals numerically for the continuous spectrum, and then as discrete sums given $n_{\nu}$ frequency bins with
+Upon careful inspection of the coupled rate equations which govern the balance
+of ionizations, recombinations, heating, and cooling, one realizes that the
+integral values, i.e. equations \ref{eq:HIIRateEquation},
+\ref{eq:HeIIRateEquation}, \ref{eq:HeIIIRateEquation}, and
+\ref{eq:TemperatureEvolution} can be tabulated as a function of column density
+(i.e. optical depth), allowing for a significant speed-up in run time. The key
+to our subsequent analysis was inspired by this same convenient feature: if
+$\Gamma_i$ and $\Heat_i$ can be tabulated for a given source
+\textit{a-priori}, it means that \textit{precision determination of the
+ionization state and temperature of the gas surrounding that source relies
+solely on accurate numerical computation of the photoionization and heating
+rate integrals in Equations \ref{eq:PhotoionizationRate} and
+\ref{eq:HeatingRate}}. So, for a given source, we can compute the values of
+these integrals numerically for the continuous spectrum, and then as discrete
+sums given $n_{\nu}$ frequency bins with
 \begin{equation}
@@
 \section{RADIATIVE TRANSFER FRAMEWORK}
-
-Our radiative transfer scheme is modeled closely after the work of \cite{Fukugita1994} and \cite{Thomas2008}, though for completeness we will reiterate the aspects of these methods most pertinent to the problem at hand.
+Our radiative transfer scheme is modeled closely after the work of \cite{Fukugita1994} and \cite{Thomas2008}, though for completeness we will reiterate here the aspects of these methods most pertinent to the problem at hand.
 In general, propagating a radiation field relies on solving differential equations governing the rate of change in the number densities of all ions and the temperature of the gas.
@@
 \begin{align}
-    \frac{d \nHII}{dt} & = \gHI \nHI - \aHII \nel \nHII   \label{HIIRateEquation} \\
-    \frac{d \nHeII}{dt} & = \gHeI \nHeI - \bHeI \nel \nHeI + \bHeII \nel \nHeII \label{HeIIRateEquation} \\ \nonumber 
-    & - \aHeII \nel \nHeII + \aHeIII \nel \nHeIII - \xHeII \nel \nHeIII  \\
-    \frac{d \nHeIII}{dt} & = \gHeII \nHeII - \bHeII \nel \nHeII + \aHeIII \nel \nHeIII . \label{HeIIIRateEquation}
+    \frac{d \nHII}{dt} & = \gHI \nHI + \bHI \nHI \nel - \aHII \nel \nHII   \label{eq:HIIRateEquation} \\
+    \frac{d \nHeII}{dt} & = \gHeI \nHeI + \bHeI \nel \nHeI - \bHeII \nel \nHeII \nonumber \\ 
+    & - \aHeII \nel \nHeII + \aHeIII \nel \nHeIII - \xHeII \nel \nHeIII  \label{eq:HeIIRateEquation} \\
+    \frac{d \nHeIII}{dt} & = \gHeII \nHeII + \bHeII \nel \nHeII - \aHeIII \nel \nHeIII . \label{eq:HeIIIRateEquation}
 \end{align}
-Each of these equations represents the balance of ionizations and recombinations for ion $i = \HII, \HeII, \HeIII$, where $\Gamma_i$ denotes the ionization rate coefficient, $\alpha_i$ ($\xi_i$) the case-A (dielectric) recombination rate coefficients, and $\beta_i$ the collisional ionization rate coefficients.
-In this work, we use the formulae in Appendix B of \cite{Fukugita1994} to compute the values of $\alpha$, $\beta$, and $\xi$.
-The ionization rate coefficients are composed of terms taking into account photo-ionization, collisional ionization, and secondary ionizations due to energetic photo-electrons.
-The photoionization rate coefficients, $\Gamma_i^P$, are given by
+Each of these equations represents the balance of ionizations and recombinations for ion $i = \HII, \HeII, \HeIII$, where $\Gamma_i$ denotes the photo-ionization rate coefficient, $\alpha_i$ ($\xi_i$) the case-A (dielectric) recombination rate coefficients, and $\beta_i$ the collisional ionization rate coefficients.
+In this work, we use the formulae in Appendix B of \citet{Fukugita1994} to compute the values of $\alpha$, $\beta$, and $\xi$.
+We absorb ionizations due to to energetic photo-electrons into the  photoionization rate coefficients, $\Gamma_i$, which are given by
+\begin{align}
+    \Gamma_i = & \int_{E_i}^{\infty} \sigma_i(E) \spec(E) \frac{dE}{E} \nonumber \\  + & \sum_j f_i \left(\frac{n_j}{n_i}\right) \int_{E_i}^{\infty} \sigma_i(E)\left(\frac{E - E_j}{E_i}\right) \spec(E) \frac{dE}{E} \label{eq:PhotoionizationRate}
+\end{align}
+where $E_i$, $\sigma_i$, $n_i$, and $f_i$ are the ionization threshold energy, bound-free absorption cross section, number density, and fraction of photo-electron energy that goes into ionization of species $i$, respectively, and the sum over $j = \HI, \HeII, \HeII$ term represents ionizations due to fast secondary electrons from photoionizations of species $j$, which has number density $n_j$.
+$\spec$ is the spectral intensity, which we write as
 \begin{equation}
-    \Gamma_i = \int_{E_i}^{\infty} \sigma_i(E) \spec(E) \frac{dE}{E}    \label{PhotoionizationRate}
-\end{equation}
-where $E_i$ and $\sigma_i$ are the ionization threshold and bound-free absorption cross section for species $i$, respectively.  $\spec$ is the spectral intensity, which we write as
-\begin{equation}
-    \spec(E, r) = \frac{\Lbol}{4\pi r^2} I(E) e^{-\tau(E, r)} .
+    \spec(E, r) = \frac{\Lbol}{4\pi r^2} I(E) e^{-\tau(E, r)} . \label{eq:Spectrum}
 \end{equation}
@@
     \tau(E, r) & = \sum_i \int_r \sigma_i(E) n_i(r) dr \nonumber \\
-               & = \sum_i \sigma_i(E) \ncol(r)
+               & = \sum_i \sigma_i(E) \ncol(r) \label{eq:OpticalDepth}
 \end{align}
 where $\ncol(r)$ is the column density of species $i$ at distance $r$ from the source.
-We calculate the bound-free absorption cross-sections, $\sigma_i$, using the fits of \cite{Verner1996}.  At each time-step, we also solve for the temperature evolution, which is given by
+We calculate the bound-free absorption cross-sections, $\sigma_i$, using the fits of \citet{Verner1996}.  At each time-step, we also solve for the temperature evolution, which is given by
 \begin{align}
     \frac{3}{2}\frac{d}{dt}\left(\frac{\kB T \nb}{\mu}\right) & = \fheat \sum_i n_i \Heat_i \nonumber \\
-    & - \sum_i \zeta_i n_e n_i - \sum_i \eta_i n_e n_i - \sum_i \psi_i n_e n_i \label{TemperatureEvolution} 
+    & - \sum_i \zeta_i n_e n_i - \sum_i \eta_i n_e n_i - \sum_i \psi_i n_e n_i \label{eq:TemperatureEvolution} 
 \end{align}
@@
 \begin{equation}
-    \Heat_i = \int_{E_i}^{\infty} \sigma_i(E) \spec(E, r)(E - E_i) \frac{dE}{E} \label{HeatingRate} ,
+    \Heat_i = \int_{E_i}^{\infty} \sigma_i(E) \spec(E, r)(E - E_i) \frac{dE}{E} \label{eq:HeatingRate} ,
 \end{equation}    
 and $\zeta_i$, $\eta_i$, and $\psi_i$ are the collisional ionization, recombination, and collisional excitation cooling coefficients, respectively.
-Additional parameters include $\nb$, the total number density of baryons (check on this), the mean molecular weight $\mu$, Boltzmann's constant $\kB$, the kinetic temperature of the gas $T$, and $\fheat$, the fraction of a photon's original energy deposited as heat (see \cite{Shull1985}, \cite{Furlanetto2010}).
-For simplicity, we neglect Hubble cooling due to cosmological expansion, as well as Compton heating/cooling, and cooling by free-free emission.
+Additional parameters include $\nb$, the total number density of baryons (check on this), the mean molecular weight $\mu$, Boltzmann's constant $\kB$, the kinetic temperature of the gas $T$, and $\fheat$, the fraction of a photon's original energy deposited as heat (see \citet{Shull1985}, \citet{Furlanetto2010}).
+For simplicity, we neglect cosmological effects, including Hubble cooling due to cosmological expansion, Compton heating/cooling off CMB photons, and photo-ionization by CMB photons.
+Currently neglecting cooling by free-free emission too.
 
-It is important to note that we can pull radial and time dependencies outside of the integrals in Equations \ref{PhotoionizationRate} and \ref{HeatingRate}, leaving
+It is important to note that we can pull radial and time dependencies outside of the integrals in Equations \ref{eq:PhotoionizationRate} and \ref{eq:HeatingRate}, leaving
 \begin{equation}
@@
 It is now apparent that the values of $\Gamma_i$ and $\Heat_i$ can be tabulated as a function of column density for a given source.
-This greatly improves the speed of 1D radiative transfer calculations, like those of \cite{Thomas2008}, since the integrals in these expressions (Equations \ref{PhotoionizationRate} and \ref{HeatingRate}) no longer need to be computed numerically on-the-fly.
-However, the lookup tables will only be 1-dimensional for the case of a hydrogen-only universe.
+This greatly improves the speed of 1D radiative transfer calculations, like those of \cite{Thomas2008}, since the integrals in these expressions (Equations \ref{eq:PhotoionizationRate} and \ref{eq:HeatingRate}) no longer need to be computed numerically on-the-fly.
+However, the lookup tables will only be 1-dimensional for the case of a hydrogen-only medium.
 
@@
 \section{CODE VERIFICATION}
-Our code, \texttt{rt1d}, solves Equations \ref{HIIRateEquation}, \ref{HeIIRateEquation}, \ref{HeIIIRateEquation}, and \ref{TemperatureEvolution} using the implicit Euler method for integration and the Newton-Raphson technique for root finding.
-In order to track the propagation of ionization fronts accurately, we limit the time-step based on a maximum neutral fraction change as introduced in \cite{Shapiro2004},
+Our code, \texttt{rt1d}, solves Equations \ref{eq:HIIRateEquation}, \ref{eq:HeIIRateEquation}, \ref{eq:HeIIIRateEquation}, and \ref{eq:TemperatureEvolution} using the implicit Euler method for integration and the Newton-Raphson technique for root finding.
+In order to track the propagation of ionization fronts accurately, we limit the time-step based on a maximum neutral fraction change as introduced in \citet{Shapiro2004},
 \begin{equation}
@@
 \end{equation}
-and additionally require that the time-step increase by a factor of two at most, as in \cite{Wise2011}.
+and additionally require that the time-step increase by a factor of two at most, as in \citet{Wise2011}.
 
@@
 
-To demonstrate the functionality of the code, we repeat tests 1 and 2 from the Radiative Transfer Comparison Project (\cite{Iliev2006}; hereafter referred to as RT06).
-Test 1 is the expansion of an HII region in a hydrogen-only, isothermal medium surrounding a monochromatic source of 13.6 eV photons.
-We adopt the parameters used in \citet{Wise2011}: constant temperature $T = 10^4 \ \mathrm{K}$, hydrogen number density $n_H = 10^{-3} \ \mathrm{cm^{-3}}$, ionized fraction $x = 1.2 \times 10^-3$, in a box $L_{\mathrm{box}} = 6.6 \ \mathrm{kpc}$ in size, and with photon luminosity $\dot{Q} = 5\times10^{48} \ \mathrm{s^{-1}}$.
-
+% I-front evolution, RT06 #1
 \begin{figure*}[htbp]
@@
 \includegraphics[scale=0.8]{figures/RT06_1_IfrontEvolution.eps}
-\caption{\textit{Top:} Comparison of the numerical (dashed) and analytic (solid) solutions for the position of an expanding ionization front as a function of time in a hydrogen-only, isothermal medium (RT06 problem \# 1).
-The numerical solution here was obtained using 400 grid cells.
+\caption{\textit{Top:} Comparison of the numerical (dashed) and analytic (solid) solutions for the position of an expanding ionization front as a function of time in a hydrogen-only, isothermal medium (RT06 problem \#1).
+The numerical solution displayed here is from the highest resolution simulation (6400 grid cells, i.e.
+$\Delta x = L_{\mathrm{box}}/6400$).
 \textit{Bottom:} Ratio of the calculated and analytic solutions as a function of time and grid resolution.
-}
+Simulations with > 200 grid cells converge on the highest resolution solution to an accuracy of $\lesssim 5 \%$.}
 \label{fig:RT_Test1_IfrontEvolution}
@@
 
+To demonstrate the functionality of the code, we repeat tests 1 and 2 from the Radiative Transfer Comparison Project (\cite{Iliev2006}; hereafter referred to as RT06).
+Test 1 is the expansion of an HII region in a hydrogen-only, isothermal medium surrounding a monochromatic source of 13.6 eV photons.
+We adopt the parameters used in \citet{Wise2011}: constant temperature $T = 10^4 \ \mathrm{K}$, uniform hydrogen number density $n_H = 10^{-3} \ \mathrm{cm^{-3}}$, ionized fraction $x = 1.2 \times 10^-3$, in a box $L_{\mathrm{box}} = 6.6 \ \mathrm{kpc}$ in size, and with photon luminosity $\dot{Q} = 5\times10^{48} \ \mathrm{s^{-1}}$.
+The classical analytic solution for the radius of an ionization front is
+\begin{equation}
+    r_{\mathrm{IF}}(t) = r_s (1 - e^{-t/t_{\mathrm{rec}}})^{1/3}
+\end{equation}    
+where $r_s$ is the Str\"{o}mgren radius, 
+\begin{equation}
+    r_s = \left(\frac{3 \dot{Q}}{4\pi \alpha_{\mathrm{HII}}(T) n_{\mathrm{HI}}^2}\right)^{1/3} ,
+\end{equation}
+and the recombination time, \trec, is defined as
+\begin{equation}
+    t_{\mathrm{rec}} = \frac{1}{\alpha_{\mathrm{HII}}(T) n_{\mathrm{HI}}} .
+\end{equation}
+This solution is approximate even in isothermal media, given that it
+assumes a constant neutral hydrogen density, $\nHI$. More accurate analytic
+solutions exist (citations), and predict a departure from the classical
+solution at $t/t_{\mathrm{rec}} \simeq 1$, which grows to a $\sim 5 \%$
+difference by $t/t_{\mathrm{rec}} \simeq 4$. Our numerical solution (see
+Figure \ref{fig:RT_Test1_IfrontEvolution}) captures this behavior very well.
+In Figure \ref{fig:RT_Test1_RadialProfiles}, we show radial profiles of the ionized and neutral fractions at three stages of the I-front expansion, again in very good agreement with the calculations in \citet{Iliev2006}.
+
 \begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=0.4]{figures/RT_Test1_RadialProfiles.ps} 
-\vspace{-20pt}
-\caption[]{Radial profiles of the neutral (solid) and ionized (dashed) fractions at $t = 10$, $30$, $100$, and $500$ Myr.}
+\centering
+\includegraphics[scale=0.4]{figures/RT_Test1_RadialProfiles.eps} 
+\caption[]{Radial profiles of the neutral (solid) and ionized (dashed) fractions at $t = 10$, $30$, $100$, and $500$ Myr for Test 1.}
 \label{fig:RT_Test1_RadialProfiles}
-\end{center}
 \end{figure}
 
-Test 2 is the same, except for the temperature is allowed to evolve according to Equation \ref{TemperatureEvolution}, and the monochromatic radiation source is replaced by a $10^5$ K blackbody spectrum.
-Results for Test 1 are shown in Figures \ref{fig:RT_Test1_IfrontEvolution} and \ref{fig:RT_Test1_RadialProfiles}, and for Test 2 in Figure \ref{fig:RT_Test2_RadialProfiles}.
+Test 2 is the same as Test 1, except now the temperature is allowed to evolve according to Equation \ref{eq:TemperatureEvolution}, and the monochromatic radiation source is replaced by a $10^5$ K blackbody spectrum.
+Radial profiles of the neutral and ionized fractions and temperature can be seen in the top and bottom panels of Figure \ref{fig:RT_Test2_RadialProfiles}, respectively.
+Again, our numerical solutions are in very good agreement with previous work.
 
 \begin{figure}[htbp]
-\begin{center}
-\includegraphics[scale=0.4]{figures/RT_Test2_RadialProfiles.ps}
-\vspace{-20pt}
+\centering
+\includegraphics[scale=0.4]{figures/RT_Test2_RadialProfiles.eps}
 \caption{\textit{Top:} Radial profiles of the neutral (solid) and ionized (dashed) fractions at $t = 10$, $100$, and $500$ Myr.
-\textit{Bottom:} Radial profiles of the kinetic temperature at $t = 10$, $100$, and $500$ Myr (solid, dashed, and dotted lines, respectively).}
+\textit{Bottom:} Radial profiles of the kinetic temperature at $t = 10$, $100$, and $500$ Myr (solid, dashed, and dotted lines, respectively).
+Results in both panels from Test 2.}
 \label{fig:RT_Test2_RadialProfiles}
-\end{center}
 \end{figure}

### 2011-10-25 rt1d/mods/Radiate.py

@@ -183,9 +183,9 @@ class Radiate:
        
         # Only solve helium rate equations if self.MultiSpeces = 1
         if self.MultiSpecies:
-            newHeII = Gamma_HeI * n_HeI - Beta_HeI * n_e * n_HeI + Beta_HeII * n_e * q[1] - \
+            newHeII = Gamma_HeI * n_HeI + Beta_HeI * n_e * n_HeI - Beta_HeII * n_e * q[1] - \
                       alpha_HeII * n_e * q[1] + alpha_HeIII * n_e * n_HeIII - xi_HeII * n_e * q[1]    
-            newHeIII = Gamma_HeII * n_HeII - Beta_HeII * n_e * n_HeII + alpha_HeIII * n_e * q[2]
+            newHeIII = Gamma_HeII * n_HeII + Beta_HeII * n_e * n_HeII - alpha_HeIII * n_e * q[2]
         else:
             newHeII = q[1]
             newHeIII = q[2]
@@ -695,23 +695,24 @@ class Radiate:
         for i, n in enumerate(nabs):
             cool += n * self.CollisionalIonizationCoolingCoefficient(T, i)
                 
+        # Cooling by recombinations
+        for i, n in enumerate(nion):
+            cool += n * self.RecombinationCoolingCoefficient(T, i)
+            
+        # Cooling by dielectronic recombination
+        cool += nion[2] * self.DielectricRecombinationCoolingCoefficient(T)
+                
+                
         # Cooling by collisional excitation
         if self.CollisionalExcitation:
             for i, n in enumerate(nabs):
                 cool += n * self.CollisionalExcitationCoolingCoefficient(T, nabs, nion, i)
         
-        # Cooling by recombinations
-        for i, n in enumerate(nion):
-            cool += n * self.RecombinationCoolingCoefficient(T, i)
-                        
-        # Cooling by dielectronic recombination
-        cool += nion[2] * self.DielectricRecombinationCoolingCoefficient(T)
-        
         # Compton cooling - from FK96
         if self.ComptonCooling:
             cool += 4. * k_B * (T - T_cmb) * (np.pi**2 / 15.) * (k_B * T_cmb / hbar / c)**3 * (k_B * T_cmb / m_e / c**2) * sigma_T * c
         
-        ## Cooling by free-free emission
+        # Cooling by free-free emission
         #cool += (nion[0] + nion[1] + 4. * nion[2]) * 1.42e-27 * 1.1 * np.sqrt(T) # Check on Gaunt factor        
                 
         cool *= n_e

### 2011-10-25 scripts/AnalyzeResolutionSuite.py

@@ -28,6 +28,7 @@ import os, re, sys
 import pylab as pl
 import numpy as np
 import rt1d.analysis as rta
+from jmpy.stats import *
 
 pl.rcParams['figure.subplot.left'] = 0.12
 pl.rcParams['figure.subplot.right'] = 0.88
@@ -41,6 +42,15 @@ if sys.argv[2] == 'cinf': path = cinf
 else: path = cfin
 
 pfs = ['dx100_dt3', 'dx200_dt3', 'dx400_dt3', 'dx800_dt3', 'dx1600_dt3', 'dx3200_dt3', 'dx6400_dt3']
+
+resolution = []
+meanerror = []
+maxerror = []
+minerror = []
+
+ds = rta.Analyze("{0}/{1}.dat".format(path, 'dx6400_dt3'))
+ds.ComputeIonizationFrontEvolution()
+ref_error = ds.rIF / ds.ranl
 
 # First, the c -> infinite tests
 mp = rta.multiplot(dims = (2, 1), panel_size = (1, 1), useAxesGrid = False)
@@ -71,6 +81,11 @@ for pf in pfs:
     else:
         mp.grid[1].plot(ds.t / ds.trec, ds.rIF / ds.ranl, color = color, ls = '-', label = r'$\Delta x = 1/{0}$'.format(int(ds.pf['GridDimensions'])))
 
+        resolution.append(ds.pf['GridDimensions'])
+        meanerror.append(np.abs(np.mean(ref_error - np.resize(ds.rIF / ds.ranl, len(ref_error)))))
+        maxerror.append(np.abs(np.max(ref_error - np.resize(ds.rIF / ds.ranl, len(ref_error)))))
+        minerror.append(np.abs(np.min(ref_error - np.resize(ds.rIF / ds.ranl, len(ref_error)))))
+
 mp.grid[1].set_ylim(0.9, 1.05) 
 mp.fix_ticks()     
 mp.grid[0].set_xlim(0, max(ds.t / ds.trec))
@@ -87,5 +102,30 @@ mp.grid[0].legend(loc = 'lower right', frameon = False, ncol = 2)
 
 pl.draw()
 pl.savefig('RT06_{0}_IfrontEvolution.png'.format(int(sys.argv[1])))
+
+pl.close()
+
+# Errors
+#f = fit(np.array(np.log10(resolution)), np.array(np.log10(meanerror)))
+#ax = pl.subplot(111)
[…]

### 2011-10-26 doc/paper/ms.tex
Got rid of rt1d.py hopefully (case-folding collision).

@@
 %%%
-\section{ASSESSING THE CONSEQUENCES OF DISCRETE RADIATION FIELDS}
+\section{ASSESSING THE CONSEQUENCES OF DISCRETE RADIATION FIELDS} \label{sec:Consequences}
 Before computing the optimal multi-frequency spectra for SEDs most commonly used in numerical simulations (\S3), we will first demonstrate the difference in results brought about by varying degrees of frequency resolution.
@@
 
-\section{METHODS}
+\section{METHODS} \label{sec:Methods}
 Upon careful inspection of the coupled rate equations which govern the balance
@@
 \end{equation}
-
-% Discretization Artifacts
-\subsection{Discretization Artifacts}
 
@@
 
-\section{RESULTS}
+\section{RESULTS} \label{sec:Results}
 
@@
 \end{align}
-Each of these equations represents the balance of ionizations and recombinations for ion $i = \HII, \HeII, \HeIII$, where $\Gamma_i$ denotes the photo-ionization rate coefficient, $\alpha_i$ ($\xi_i$) the case-A (dielectric) recombination rate coefficients, and $\beta_i$ the collisional ionization rate coefficients.
+Each of these equations represents the balance of ionizations and
+recombinations for ions $\HII$, $\HeII$, $\HeIII$. We associate the index $i$
+with neutral species, $i = \HI, \HeI, \HeII$, and define $\Gamma_i$ as the
+photo-ionization rate coefficient, $\alpha_i$ ($\xi_i$) the case-A
+(dielectric) recombination rate coefficients, $\beta_i$ the collisional
+ionization rate coefficients, and $\nel$ the number density of electrons,
+\begin{equation}
+    \nel = \nHII + \nHeII + 2\nHeIII .
+\end{equation}
 In this work, we use the formulae in Appendix B of \citet{Fukugita1994} to compute the values of $\alpha$, $\beta$, and $\xi$.
@@
 \end{align}
-where $E_i$, $\sigma_i$, $n_i$, and $f_i$ are the ionization threshold energy, bound-free absorption cross section, number density, and fraction of photo-electron energy that goes into ionization of species $i$, respectively, and the sum over $j = \HI, \HeII, \HeII$ term represents ionizations due to fast secondary electrons from photoionizations of species $j$, which has number density $n_j$.
+where $E_i$, $\sigma_i$, $n_i$, and $f_i$ are the ionization threshold energy, bound-free absorption cross section, number density, and fraction of photo-electron energy that goes into ionization of species $i$, respectively.
+The term summing over $j$, where $j = \HI, \HeI, \HeII$, represents ionizations of species $i$ due to fast secondary electrons from photoionizations of species $j$, which has number density $n_j$.
 $\spec$ is the spectral intensity, which we write as
@@
 and $\zeta_i$, $\eta_i$, and $\psi_i$ are the collisional ionization, recombination, and collisional excitation cooling coefficients, respectively.
-Additional parameters include $\nb$, the total number density of baryons (check on this), the mean molecular weight $\mu$, Boltzmann's constant $\kB$, the kinetic temperature of the gas $T$, and $\fheat$, the fraction of a photon's original energy deposited as heat (see \citet{Shull1985}, \citet{Furlanetto2010}).
-For simplicity, we neglect cosmological effects, including Hubble cooling due to cosmological expansion, Compton heating/cooling off CMB photons, and photo-ionization by CMB photons.
-Currently neglecting cooling by free-free emission too.
+The total number density of baryons is given by $\nb$, the mean molecular weight, $\mu$, Boltzmann's constant, $\kB$, the kinetic temperature of the gas, $T$, and the fraction of a photon's original energy deposited as heat, $\fheat$ (see \citet{Shull1985}, \citet{Furlanetto2010}).
+For simplicity, our current treatment neglects cooling via free-free emission, as well as cosmological effects, including Hubble cooling due to cosmological expansion, Compton heating/cooling off cosmic microwave background (CMB) photons, and photo-ionization by Wien-tail CMB photons.
 
@@
 This greatly improves the speed of 1D radiative transfer calculations, like those of \cite{Thomas2008}, since the integrals in these expressions (Equations \ref{eq:PhotoionizationRate} and \ref{eq:HeatingRate}) no longer need to be computed numerically on-the-fly.
-However, the lookup tables will only be 1-dimensional for the case of a hydrogen-only medium.
+It also provides the basis for our frequency resolution optimization strategy, as presented in \S\ref{sec:Methods}.
+Note, however, that the dimensionality of these lookup tables is equal to the number of absorbing species, so the tables for simulations including only hydrogen are 1D, and those including helium are 3D.
 
@@
 
-At this juncture it is worth mentioning the numerical artifacts that can result due to poor temporal or spatial resolution.
+%At this juncture it is worth mentioning the numerical artifacts that can result due to poor temporal or spatial resolution.
 
@@
 \textit{Bottom:} Ratio of the calculated and analytic solutions as a function of time and grid resolution.
-Simulations with > 200 grid cells converge on the highest resolution solution to an accuracy of $\lesssim 5 \%$.}
+Simulations with $>$ 200 grid cells converge on the highest resolution solution to an accuracy of $\lesssim 5 \%$.}
 \label{fig:RT_Test1_IfrontEvolution}
@@
 Figure \ref{fig:RT_Test1_IfrontEvolution}) captures this behavior very well.
-In Figure \ref{fig:RT_Test1_RadialProfiles}, we show radial profiles of the ionized and neutral fractions at three stages of the I-front expansion, again in very good agreement with the calculations in \citet{Iliev2006}.
+In Figure \ref{fig:RT_Test1_RadialProfiles}, we show radial profiles of the ionized and neutral fractions at three stages of the I-front expansion, which are again in very good agreement with the calculations presented in \citet{Iliev2006}.
 

### 2011-10-26 rt1d/mods/Radiate.py

@@ -359,7 +359,7 @@ class Radiate:
         # If the speed of light = c, things are trickier    
         else:                
                                           
-            if self.HIIRestrictedTimestep: dtphot = 1e50 * np.ones_like(self.grid) #dtphot = 1e50 * np.ones(len(packs))         
+            if self.HIIRestrictedTimestep: dtphot = 1e50 * np.ones_like(self.grid)    
                                     
             # Loop over photon packages, updating values in cells: data -> newdata
             for j, pack in enumerate(packs):

### 2011-10-26 rt1d/mods/SetDefaultParameterValues.py

@@ -29,7 +29,7 @@ def SetDefaultParameterValues():
         
         # ODE Solver
         "ODEAdaptiveStep": 1, \
-        "ODEMinStep": 1e-8, \
+        "ODEMinStep": 1e-10, \
         "ODEMaxStep": 1e-1, \
         "ODErtol": 1e-3, \
         "ODEatol": 1e-6, \

### 2011-10-31 rt1d/analysis/Analysis.py
Cleaned up RT06_1_ResolutionTestSuite a bit.

@@ -9,11 +9,10 @@ Description: Functions to calculate various quantities from our rt1d datasets.
      
 """
 
-import misc
 import numpy as np
 import pylab as pl
-from constants import *
 from Multiplot import *
+from rt1d.mods.Constants import *
 from rt1d.analysis.Dataset import Dataset
 from rt1d.mods.ComputeCrossSections import PhotoIonizationCrossSection
 

### 2011-11-10 rt1d/analysis/__init__.py
Fixed compute photon timestep - now use all cells outside start radius rather than cell 1.

@@ -1,0 +1,4 @@
+from rt1d.analysis.Dataset import Dataset 
+from rt1d.analysis.Analysis import Analyze
+from rt1d.analysis.Multiplot import multiplot
+ 

### 2011-11-10 rt1d/mods/Radiate.py

@@ -1,0 +1,776 @@
+"""
+Radiate.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on 2010-10-18.
+
+Description: This routine essentially runs the show.  The method 'EvolvePhotons' is the
+driver of rt1d, calling our solvers which call all the various physics modules.
+     
+"""
+
+import copy, math
+import numpy as np
+from rt1d.mods.ComputeCrossSections import PhotoIonizationCrossSection
+from rt1d.mods.RadiationSource import RadiationSource
+from rt1d.mods.SecondaryElectrons import SecondaryElectrons
+from rt1d.mods.Interpolate import Interpolate
+from rt1d.mods.Cosmology import Cosmology
+from rt1d.mods.SolveRateEquations import SolveRateEquations
+from Integrate import simpson as integrate
+from progressbar import *
+
+try:
+    from mpi4py import MPI
+    rank = MPI.COMM_WORLD.rank
+    size = MPI.COMM_WORLD.size
+except ImportError:
+    print "Module mpi4py not found.  No worries, we'll just run in serial."
+    rank = 0
+    size = 1
+
+m_e = 9.10938188*10**-28 		# Electron mass - [m_e] = g
+m_p = 1.67262158*10**-24		# Proton mass - [m_p] = g
+m_n = 1.67492729*10**-24        # Neutron mass - [m_n] = g
+k_B = 1.3806503*10**-16			# Boltzmann's constant - [k_B] = erg/K
+sigma_T = 6.65*10**-25			# Cross section for Thomson scattering - [sigma_T] = cm^2
+h = 6.626068*10**-27 			# Planck's constant - [h] = erg*s
+hbar = h / (2 * np.pi) 			# H-bar - [h_bar] = erg*s
+c = 29979245800.0 			    # Speed of light - [c] = cm/s
+
+m_H = m_p + m_e
+m_HeI = 2.0 * (m_p + m_n + m_e)
+m_HeII = 2.0 * (m_p + m_n) + m_e
+
+E_th = [13.6, 24.6, 54.4]
+
+min_density = 1e-12
+
+# Widget for progressbar.
+widget = ["Ray Casting: ", Percentage(), ' ', Bar(marker = RotatingMarker()), ' ', ETA(), ' ']
+
+class Radiate:
+    def __init__(self, pf, data, itabs, n_col): 
+        self.rs = RadiationSource(pf)
[…]

### 2011-11-10 rt1d/mods/RadiationSource.py

@@ -1,0 +1,174 @@
+"""
+RadiationSource.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on 2010-10-26.
+
+Description: Radiation source class, contains all functions associated with initializing a 
+radiation source.  Mainly used for calculating and normalizing it's luminosity with time.
+
+Notes: 
+
+SourceType = 0: Monochromatic/Polychromatic emission of SpectrumPhotonLuminosity photons per second.
+  -The difference between this and SourceType > 0 is that the bolometric luminosity may change depending
+   on your choice of bins.
+     
+     
+"""
+
+import numpy as np
+from Integrate import simpson as integrate                  # why did scipy.integrate.quad mess up LphotNorm? 
+
+h = 6.626068e-27     			                            # Planck's constant - [h] = erg*s
+k_B = 1.3806503e-16     			                        # Boltzmann's constant - [k_B] = erg/K
+c = 29979245800.0 				                            # Speed of light - [c] = cm/s
+G = 6.673e-8     				                            # Gravitational constant - [G] = cm^3/g/s^2
+sigma_SB = 2.0 * np.pi**5 * k_B**4 / 15.0 / c**2 / h**3     # Stefan-Boltzmann constant - [sigma_SB] = erg / cm^2 / deg^4 / s
+lsun = 3.839e33                                             # Solar luminosity - erg / s
+m_p = 1.67262158e-24    		                            # Proton mass - [m_p] = g
+sigma_T = 6.65e-25	        		                        # Cross section for Thomson scattering - [sigma_T] = cm^2
+g_per_msun = 1.98892e33                                     # Mass of the sun - [g_per_sun] = g
+erg_per_ev = 1.60217646e-19 / 1e-7                          # Conversion between ergs and eV
+cm_per_rsun = 695500.0 * 1e5                                # Radius of the sun - [cm_per_rsun] = cm
+s_per_yr = 365.25 * 24 * 3600                               # Seconds per year
+t_edd = 0.45 * 1e9 * s_per_yr                               # Eddington timescale (see eq. 1 in Volonteri & Rees 2005) 
+
+np.seterr(all = 'ignore')   # exp overflow occurs when integrating BB - will return 0 as it should for x large
+
+SchaererTable = {
+                "Mass": [5, 9, 15, 25, 40, 60, 80, 120, 200, 300, 400, 500, 1000], 
[…]

### 2011-11-10 rt1d/mods/ReadParameterFile.py

@@ -1,0 +1,192 @@
+"""
+ReadParameterFile.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on 2010-10-14.
+
+Description: 
+     
+"""
+
+import copy, h5py
+import numpy as np
+from rt1d.mods.SetDefaultParameterValues import SetDefaultParameterValues
+
+cm_per_kpc = 3.08568 * 10**21
+s_per_myr = 365.25 * 24 * 3600 * 10**6
+
+def ReadParameterFile(pf):
+    """
+    Read in the parameter file, and parse the parameter names and arguments.
+    Return a dictionary that contains all parameters and their values, whether 
+    they be floats, tuples, or lists.
+    """
+    f = open(pf, "r")
+    pf_dict = SetDefaultParameterValues()
+    for line in f:
+        if not line.split(): continue
+        if line.split()[0][0] == "#": continue
+        
+        # This will prevent crashes if there is not a blank line at the end of the parameter file
+        if line[-1] != '\n': line += '\n'
+        
+        # Cleave off end-of-line comments.
+        line = line[:line.rfind("#")].strip()
+        
+        # Read in the parameter name and the parameter value(s).
+        parname, eq, parval = line.partition("=")
+                                        
+        # ProblemType option
+        if parname.strip() == 'ProblemType' and float(parval) > 0:
+            pf_new = ProblemType(float(parval))
+            for param in pf_new: pf_dict[param] = pf_new[param]
+                        
+        # Else, actually read in the parameter           
+        try: parval = float(parval)
+        except ValueError:
+            if parval.replace('_', '').replace('.', '').strip().isalnum(): 
+                parval = str(parval.strip())
+            else:
+                parval = parval.strip().split(",")
+                tmp = []                           
+                if parval[0][0] == '[':
+                    for element in parval: tmp.append(float(element.strip("[,]")))
+                    parval = list(tmp)
+                else:
[…]

### 2011-12-08 doc/manual/rt1d_manual.tex
Fixed compute photon timestep - now use all cells outside start radius rather than cell 1.
Added some stuff to the manual.

@@
 \usepackage{pslatex}
+\usepackage{natbib}
+\usepackage[pdfborder={0 0 0}]{hyperref}
+\usepackage[all]{hypcap}
 
@@
 \begin{document}
+
+%%% START custom commands
+\newcommand{\HI}{\text{HI}}
+\newcommand{\HII}{\text{HII}}
+\newcommand{\HeI}{\text{HeI}}
+\newcommand{\HeII}{\text{HeII}}
+\newcommand{\HeIII}{\text{HeIII}}
+\newcommand{\nH}{n_{\text{H}}} 
+\newcommand{\nHI}{n_{\text{HI}}} 
+\newcommand{\nHII}{n_{\text{HII}}}  
+\newcommand{\nHeI}{n_{\text{HeI}}}
+\newcommand{\nHeII}{n_{\text{HeII}}}
+\newcommand{\nHeIII}{n_{\text{HeIII}}}  
+\newcommand{\nel}{n_{\text{e}}}  
+\newcommand{\nb}{n_{\text{b}}}
+\newcommand{\nnu}{$n_{\nu}$ }
+\newcommand{\ncol}{n_i^{\text{col}}}
+\newcommand{\gHI}{\Gamma_{\text{HI}}}  
+\newcommand{\gHeI}{\Gamma_{\text{HeI}}}
+\newcommand{\gHeII}{\Gamma_{\text{HeII}}}
+\newcommand{\aHII}{\alpha_{\text{HII}}}  
+\newcommand{\aHeII}{\alpha_{\text{HeII}}}  
+\newcommand{\aHeIII}{\alpha_{\text{HeIII}}}
+\newcommand{\bHI}{\beta_{\text{HI}}} 
+\newcommand{\bHeI}{\beta_{\text{HeI}}}  
+\newcommand{\bHeII}{\beta_{\text{HeII}}}  
+\newcommand{\xHeII}{\xi_{\text{HeII}}}
+\newcommand{\kB}{k_{\text{B}}}
+\newcommand{\fheat}{f_i^{\text{heat}}}
+\newcommand{\fion}{f_i^{\text{ion}}}
+\newcommand{\Lbol}{\mathcal{L}_{\text{bol}}}
+\newcommand{\spec}{\mathcal{N}}
+\newcommand{\Heat}{\mathcal{H}}
+\newcommand{\trec}{$t_{\text{rec}}$}
+\newcommand{\Lbox}{L_{\mathrm{box}}}
+\newcommand{\dx}{\Delta x}
+\newcommand{\dd}{\text{d}}
+\newcommand{\Htwo}{\mathrm{H}_2}
+\newcommand{\eV}{\mathrm{eV}}
+%%% END custom commands
 
@@
 \section{Introduction}
+Let's see what happens to the gas around stars and black holes, shall we?
 
@@
 \section{Methods}
+Our radiative transfer scheme is modeled closely after the work of
+\cite{Fukugita1994} and \cite{Thomas2008}, though for completeness we will
+reiterate the aspects of these methods most pertinent to the problem at hand.
+In general, propagating a radiation field relies on solving differential
+equations governing the rate of change in the number densities of all ions and
+the temperature of the gas. Assuming a medium consisting of hydrogen and
+helium only, we first solve for the relative abundances of each ion via
+\begin{align}
+    \frac{d \nHII}{dt} & = \gHI \nHI + \bHI \nel \nHI - \aHII \nel \nHII   \label{eq:HIIRateEquation} \\ 
+    \frac{d \nHeII}{dt} & = \gHeI \nHeI + \bHeI \nel \nHeI - \bHeII \nel \nHeII - \aHeII \nel \nHeII + \aHeIII \nel \nHeIII - \xHeII \nel \nHeIII  \label{eq:HeIIRateEquation} \\ 
+    \frac{d \nHeIII}{dt} & = \gHeII \nHeII + \bHeII \nel \nHeII - \aHeIII \nel \nHeIII . \label{eq:HeIIIRateEquation}
+\end{align}
+Each of these equations represents the balance of ionizations and
+recombinations for ions $\HII$, $\HeII$, $\HeIII$. We associate the index $i$
+with neutral species, $i = \HI, \HeI, \HeII$, and define $\Gamma_i$ as the
+photo-ionization rate coefficient, $\alpha_i$ ($\xi_i$) as the case-A
+(dielectric) recombination rate coefficients, $\beta_i$ as the collisional
+ionization rate coefficients, and $\nel = \nHII + \nHeII + 2\nHeIII$ as the number density of electrons.
+
+In this work, we use the formulae in Appendix B of \cite{Fukugita1994} to
+compute the values of $\alpha_i$, $\beta_i$, and $\xi_i$.
+As you might expect, collisional ionization and recombination depend solely on the temperature of the gas.
+For example, collisional ionization of hydrogen (from $n = 1$) is governed by
+\begin{equation}
+    \beta_{1,\HI} = 5.85 \times 10^{-11} T^{1/2} \left[1 + \left(\frac{T}{10^5 \mathrm{K}}\right)^{1/2}\right]^{-1} \mathrm{exp}\left[-1.578 \times \left(\frac{T}{10^5 \mathrm{K}}\right)^{-1}\right] \mathrm{cm^3 s^{-1}}
+\end{equation}
+and recombinations (to $n \ge 1$) by
+\begin{equation}
+    \alpha_{\HII} = 6.28 \times 10^{-11} T^{-1/2} \left(\frac{T}{10^3\mathrm{K}}\right)^{-1/5} \left[1 + \left(\frac{T}{10^6\mathrm{K}}\right)^{0.7}\right]^{-1} \mathrm{cm^3s^{-1}}
+\end{equation}
+
+We absorb ionizations due to to energetic photo-electrons into the
+photoionization rate coefficients, $\Gamma_i$, which are given by
+\begin{equation}
+    \Gamma_i = \int_{E_i}^{\infty} \sigma_i(E) \spec(E) \frac{dE}{E} + \sum_j \fion \left(\frac{n_j}{n_i}\right) \int_{E_i}^{\infty} \sigma_i(E)\left(\frac{E - E_j}{E_i}\right) \spec(E) \frac{dE}{E} \label{eq:PhotoionizationRate}
+\end{equation}
+where $E_i$, $\sigma_i$, $n_i$, and $f_i$ are the ionization threshold energy,
+bound-free absorption cross section, number density, and fraction of
+photo-electron energy that goes into ionization of species $i$, respectively.
+The term summing over $j$, where $j = \HI, \HeI, \HeII$, represents
+ionizations of species $i$ due to fast secondary electrons from
+photoionizations of species $j$, which has number density $n_j$. $\spec$ is
+the spectral intensity, which we write as
+\begin{equation}
+    \spec(E, r) = \frac{\Lbol}{4\pi r^2} I(E) e^{-\tau(E, r)} . \label{eq:Spectrum}
+\end{equation}
+Here, $r$ is the distance from the source, $I(E)$ is the normalized spectral
+energy distribution (SED), satisfying
+\begin{equation}
+    \int_E I(E) dE = 1 ,
+\end{equation}
+and the optical depth $\tau(E, r)$ is given as a sum over absorbing species,
+\begin{align}
+    \tau(E, r) & = \sum_i \int_r \sigma_i(E) n_i(r) dr \nonumber \\
+               & = \sum_i \sigma_i(E) \ncol(r) \label{eq:OpticalDepth}
+\end{align}
+where $\ncol(r)$ is the column density of species $i$ at distance $r$ from the
+source. We calculate the bound-free absorption cross-sections, $\sigma_i$,
+using the fits of \cite{Verner1996}.
+
+To be more explicit, let's write out $\Gamma_{\HI}$ in its entirety.
+\begin{align}
+    \Gamma_{\HI} & = \int_{13.6\eV}^{\infty}\sigma_{\HI}(E) \spec(E) \frac{dE}{E} \nonumber \\
+    & + f_{\HI}^{\mathrm{ion}} \left(\frac{n_{\HI}}{n_{\HI}}\right) \int_{13.6\eV}^{\infty}\sigma_{\HI}(E) \left(\frac{E - 13.6\eV}{E}\right)\spec(E)\frac{dE}{E} \nonumber \\
+    & + f_{\HeI}^{\mathrm{ion}} \left(\frac{n_{\HeI}}{n_{\HI}}\right) \int_{24.4\eV}^{\infty}\sigma_{\HeI}(E) \left(\frac{E - 24.4\eV}{E}\right)\spec(E)\frac{dE}{E} \nonumber \\
+    & + f_{\HeII}^{\mathrm{ion}} \left(\frac{n_{\HeII}}{n_{\HI}}\right) \int_{54.4\eV}^{\infty}\sigma_{\HeI}(E) \left(\frac{E - 54.4\eV}{E}\right)\spec(E)\frac{dE}{E}
+\end{align}    
+The first term accounts for all hydrogen photoionization.
+The final three terms account for collisional ionization of hydrogen due to secondary photo-electrons from ionizations of neutral hydrogen (second term), neutral helium (third term), and singly ionized helium (fourth term).
+Collisional ionizations due to thermal electrons in the gas are accounted for separately through the $\beta_i$ coefficients.
+
+At each time-step, we also solve for the temperature evolution, $dT/dt$, which
+is given by
+\begin{equation}
+    \frac{3}{2}\frac{d}{dt}\left(\frac{\kB T \nb}{\mu}\right) = \fheat \sum_i n_i \Heat_i - \sum_i \zeta_i n_e n_i - \sum_i \eta_i n_e n_i - \sum_i \psi_i n_e n_i \label{eq:TemperatureEvolution} 
+\end{equation}
+where $\Heat_i$ is the photo-electric heating rate, given by
+\begin{equation}
+    \Heat_i = \int_{E_i}^{\infty} \sigma_i(E) \spec(E, r)(E - E_i) \frac{dE}{E} \label{eq:HeatingRate} ,
+\end{equation}    
+and $\zeta_i$, $\eta_i$, and $\psi_i$ are the collisional ionization,
+recombination, and collisional excitation cooling coefficients, respectively.
+The constants in Equation \ref{eq:HeatingRate} are the total number density of
+baryons, $\nb$, the mean molecular weight, $\mu$, Boltzmann's constant, $\kB$,
+and the fraction of a photon's original energy deposited as heat, $\fheat$. In
+the remaining sections we only include the effects of secondary electrons when
+considering X-ray sources, which emit photons in the range $10^2\mathrm{eV} <
+E < 10^4\mathrm{eV}$. In this regime, the values of $\fheat$ and $\fion$
+computed via the formulae of \cite{Shull1985} are sufficiently accurate, but
+for radiation at lower energies, the lookup tables of \cite{Furlanetto2010}
+would be more appropriate.
+
+An important simplification is that all radial and time dependencies can be
+pulled outside of the integrals in Equations \ref{eq:PhotoionizationRate} and
+\ref{eq:HeatingRate}, leaving (ignoring the secondary ionization term)
+\begin{equation}
+    \Gamma_i(n^{\mathrm{col}}) = A \int_{E_i}^{\infty} \sigma_i(E) I(E) e^{-\tau(E, n^{\mathrm{col}})} \frac{dE}{E} \label{eq:Gamma_final} ,
+\end{equation}
+and
+\begin{equation}
+    \Heat_i(n^{\mathrm{col}}) = A \int_{E_i}^{\infty} \sigma_i(E) I(E) (E - E_i) e^{-\tau(E, n^{\mathrm{col}})} \frac{dE}{E} \label{eq:Heat_final}
+\end{equation}
+where $A \equiv \Lbol / 4 \pi r^2$, and $n^{\mathrm{col}} =
+(n_{\mathrm{HI}}^{\mathrm{col}}, n_{\mathrm{HeI}}^{\mathrm{col}},
+n_{\mathrm{HeII}}^{\mathrm{col}})$.
+
+From Equations \ref{eq:Gamma_final} and \ref{eq:Heat_final} it is clear that
+the values of $\Gamma_i$ and $\Heat_i$ can be tabulated as a function of
+column density for a given source. This greatly improves the speed of 1D
+radiative transfer calculations, like those of \cite{Thomas2008}, since
+integral values in Equations \ref{eq:HIIRateEquation},
+\ref{eq:HeIIRateEquation}, \ref{eq:HeIIIRateEquation}, and
+\ref{eq:HeatingRate} no longer need to be computed numerically on-the-fly. It
+also provides the basis for our frequency resolution optimization strategy, as
+presented in \S\ref{sec:Methods}. Note, however, that the dimensionality of
+these lookup tables is equal to the number of absorbing species, so the tables
+for simulations including only hydrogen are 1D, while those including helium
+are 3D. If we choose to adopt the secondary electron treatment of
+\cite{Furlanetto2010} our lookup tables inherit an additional dimension, as
+the secondary ionization and heating factors $\fion$ and $\fheat$ depend on
+both energy and ionized fraction, $x_i$. Since integration takes place over
+energy, $x_i$ becomes the final dimension of the lookup tables.
+
+For simplicity, our current treatment neglects cooling via free-free emission,
+as well as cosmological effects, including Hubble cooling due to cosmological
+expansion, Compton heating/cooling off cosmic microwave background (CMB)
+photons, and photo-ionization by Wien-tail CMB photons.
 
@@
 \section{Classes}
-In order of appearance in \texttt{rt1d.py}.
+In order of appearance in \texttt{RT1D.py}.
 
@@
 \subsection{\texttt{Radiate}}
+\subsection{\texttt{SolveRateEquations}}
+\subsection{\texttt{Interpolate}}
+\subsection{\texttt{WriteData}}
 
@@
 \item [\texttt{ProblemType = 2}] This problem is almost exactly the same as \texttt{ProblemType = 1}.
-However, the temperature is allowed to evolve, and instead of using a monochromatic spectrum, we sample a $10^5$ K blackbody with four energy groups.
+However, the temperature is allowed to evolve, and instead of using a monochromatic spectrum, we use a $10^5$ K blackbody spectrum.
 Again, the initial density, temperature, box size, and energy groups are the same as the values used in Test 1 of Wise \& Abel 2010.
 
-\item [\texttt{ProblemType = 2.1}] Same as \texttt{ProblemType = 2} for everything, except we use a continuous blackbody spectrum rather than a discrete one.
+\item [\texttt{ProblemType = 2.1}] Same as \texttt{ProblemType = 2} for everything, except we sample the continuous blackbody spectrum with the four energy groups used by Wise \& Abel 2010.
 
+\item [\texttt{ProblemType = 3.1}] Test problem \# 3 from \cite{Iliev2006} -- I-front trapping in a dense clump and the formation of a shadow.
+The setup is the same as in \texttt{ProblemType = 2.1}, except a 1D clump is initialized at $x = 0.76 L_{\mathrm{box}}$, overdensity $\delta = 200$, radius $0.12$kpc, and temperature $T = 40$K, and the radiation field is plane-parallel.
 \end{description}
@@
 
-\item [TemperatureProfile]
+\item [\texttt{TemperatureProfile}]
+
+\item [\texttt{InitialHIIFraction}]
 
@@
 \subsection{Source Parameters}
+
+\subsection{Physics Parameters}
 
@@
 
+%%% 
+%% Convergence
+%%%
 \section{Convergence}
@@
 
+\newpage
+\bibliography{references}
+\bibliographystyle{plain}
+
 \end{document}

### 2011-12-12 doc/manual/rt1d_manual.tex
Added notes about rtol and atol to manual.

@@
 
+\begin{description}
+
+\item [\texttt{ODEatol}] Absolute tolerance of the ODE solver.
+If ionized fractions are zero to this tolerance, they are reset to \texttt{MinimumSpeciesFraction}.
+Combination of this parameter and \texttt{ODErtol} (below) determine whether or not the ODE time-step is cut in half.
+(Default = $10^{-5}$)
+
+\item [\texttt{ODErtol}]
+Relative tolerance of the ODE solver.
+Combination of this parameter and \texttt{ODEatol} determine whether or not the ODE time-step is cut in half via the formulae $|y_2 - y_1| = \texttt{ODEatol} + \texttt{ODErtol}\times y_2$.
+(Default = $10^{-5}$)
+
+\end{description}
+
 \subsection{Source Parameters}

### 2011-12-12 rt1d/analysis/Analysis.py

@@ -29,10 +29,8 @@ class Analyze:
         self.grid = np.arange(self.GridDimensions)
         self.LengthUnits = self.pf['LengthUnits']
         self.StartRadius = self.pf['StartRadius']
-
-        #self.r = self.pf['LengthUnits'] * self.grid / self.pf['GridDimensions']
-        #self.StartRadius = self.pf["StartRadius"] * self.pf['LengthUnits'] / cm_per_kpc
         
+        # Grid
         if self.pf['LogarithmicGrid']:
             self.lgrid = [0]
             self.lgrid.extend(np.logspace(0, np.log10(self.GridDimensions - 1), self.GridDimensions - 1))

### 2011-12-14 doc/manual/rt1d_manual.tex
Everything good for the paper?

@@
 
+%%% 
+%% Troubleshooting
+%%%
+\section{Troubleshooting}
+If your results look horribly, horribly wrong, here are a few places where you could've gone wrong.  We'll start with things that may occur with the stable version of the code.
+
+Now, if you modify or add anything in the source code (awesome by the way!), you're far more likely to encounter troubles.
+
+\begin{description}
+    \item[Totally Bogus Results] This is will almost certainly happen at some point in the development process.
+But, if it happens even after you're \textit{positive} the error was fixed, my guess is that you are using the integral tables that were generated with the previous, error-ridden version of the code.
+Try removing the integral tables and re-running your simulation.
+\end{description}
+
 \newpage

### 2011-12-14 rt1d/mods/RadiationSource.py

@@ -61,9 +61,11 @@ class RadiationSource:
         self.Emax = pf["SpectrumMaxEnergy"]
         self.EminNorm = pf["SpectrumMinNormEnergy"]
         self.EmaxNorm = pf["SpectrumMaxNormEnergy"]
-                    
+
         self.E = np.array(pf["DiscreteSpectrumSED"])
-        self.F = np.array(pf["DiscreteSpectrumRelLum"])
+        self.F = np.array(pf["DiscreteSpectrumRelLum"])            
+        
+        # SourceType 0, 1, 2
         self.Lph = pf["SpectrumPhotonLuminosity"]
             
         # SourceType = 1, 2
@@ -86,13 +88,14 @@ class RadiationSource:
         
         # SourceType = 4
         self.SpectrumAbsorbingColumn = pf["SpectrumAbsorbingColumn"]
-                        
+                     
         # Normalize spectrum
-        self.LuminosityNormalization = self.NormalizeLuminosity()
-        
-        # Calculate mean emission energy
-        #self.MeanEmissionEnergy = quad(lambda E: E * self.Spectrum(E), 13.6, np.inf, epsrel = 1e-15)
-                                                                
+        self.LuminosityNormalization = self.NormalizeLuminosity()    
+                        
+        # Possibly override self.F - only makes sense for monochromatic sources
+        if pf["ConserveIonizingPhotonLuminosity"]:
+            self.F[0] = self.IonizingPhotonLuminosity() * self.E[0] * erg_per_ev / self.BolometricLuminosity()    
+                                    
     def Spectrum(self, E):
         """
         Return the fraction of the bolometric luminosity emitted at this energy.  This quantity is dimensionless, and its integral should be 1.
@@ -186,7 +189,18 @@ class RadiationSource:
             if self.FixedSourceMass: Mnow = self.M
             else: Mnow = self.M * np.exp( ((1.0 - self.epsilon) / self.epsilon) * t / t_edd)
             return self.epsilon * 4.0 * np.pi * G * Mnow * g_per_msun * m_p * c / sigma_T
-            
+    
+    def IonizingPhotonLuminosity(self):
+        """
+        Return Qdot (photons / s) for this source.
+        """
+        
+        if self.SourceType < 3:
+            return self.Lph
+        else:
+            integrand = lambda E: self.Spectrum(E) / E / erg_per_ev
+            return self.BolometricLuminosity() * quad(integrand, self.Emin, self.Emax)[0]
+                            
[…]

### 2011-12-14 rt1d/mods/ReadParameterFile.py

@@ -123,7 +123,7 @@ def ProblemType(pt):
              "Isothermal": 0, "MultiSpecies": 0, "SecondaryIonization": 0, "CosmologicalExpansion": 0, \
              "DensityProfile": 0, "InitialDensity": 1e-3, "TemperatureProfile": 0, "InitialTemperature": 1e2, \
              "IonizationProfile": 0, "InitialHIIFraction": 1.2e-3, "SourceType": 1, "SourceLifetime": 1e10, \
-             "SpectrumPhotonLuminosity": 5e48, "DiscreteSpectrum": 1, \
+             "SpectrumPhotonLuminosity": 5e48, "HIColumnMax": 1e20, "DiscreteSpectrum": 1, \
              "DiscreteSpectrumSED": [16.74, 24.65, 34.49, 52.06], "DiscreteSpectrumRelLum": [0.277, 0.335, 0.2, 0.188]
             }         
     

### 2011-12-22 doc/manual/rt1d_manual.tex
Made some changes to the manual...going to try to start implementing an interactive mode.

@@
     \Gamma_{\HI} & = \int_{13.6\eV}^{\infty}\sigma_{\HI}(E) \spec(E) \frac{dE}{E} \nonumber \\
-    & + f_{\HI}^{\mathrm{ion}} \left(\frac{n_{\HI}}{n_{\HI}}\right) \int_{13.6\eV}^{\infty}\sigma_{\HI}(E) \left(\frac{E - 13.6\eV}{E}\right)\spec(E)\frac{dE}{E} \nonumber \\
-    & + f_{\HeI}^{\mathrm{ion}} \left(\frac{n_{\HeI}}{n_{\HI}}\right) \int_{24.4\eV}^{\infty}\sigma_{\HeI}(E) \left(\frac{E - 24.4\eV}{E}\right)\spec(E)\frac{dE}{E} \nonumber \\
-    & + f_{\HeII}^{\mathrm{ion}} \left(\frac{n_{\HeII}}{n_{\HI}}\right) \int_{54.4\eV}^{\infty}\sigma_{\HeI}(E) \left(\frac{E - 54.4\eV}{E}\right)\spec(E)\frac{dE}{E}
+    & + f_{\HI}^{\mathrm{ion}} \left(\frac{n_{\HI}}{n_{\HI}}\right) \int_{13.6\eV}^{\infty}\sigma_{\HI}(E) \left(\frac{E - 13.6\eV}{13.6\eV}\right)\spec(E)\frac{dE}{E} \nonumber \\
+    & + f_{\HeI}^{\mathrm{ion}} \left(\frac{n_{\HeI}}{n_{\HI}}\right) \int_{24.4\eV}^{\infty}\sigma_{\HeI}(E) \left(\frac{E - 24.4\eV}{13.6\eV}\right)\spec(E)\frac{dE}{E} \nonumber \\
+    & + f_{\HeII}^{\mathrm{ion}} \left(\frac{n_{\HeII}}{n_{\HI}}\right) \int_{54.4\eV}^{\infty}\sigma_{\HeI}(E) \left(\frac{E - 54.4\eV}{13.6\eV}\right)\spec(E)\frac{dE}{E}
 \end{align}    
@@
 
+%%%
+%% INTERACTIVE MODE
+%%%
+\section{Interactive Mode}
+Is it worth doing this?  Would want it to look something like this:
+
+\begin{verbatim}
+import rt1d
+import numpy as np
+pf = rt1d.SetDefaultParameterValues()
+r = np.linspace(3e19, 3e21, 100) # 100 grid cells between 0.01-1kpc. (optional)
+data = rt1d.Shine(pf, r)    
+\end{verbatim}
+
+This would make it easy to:
+\begin{itemize}
+    \item Allow users to write their own rt1d `scripts.'
+    \item Change source properties manually with time.
+    \item Making arbitrarily structured grids.
+    \item Very specific initial conditions.
+\end{itemize}
+
+How to change the code to allow for this?
+Define main loop in RT1D.py as Shine.  Insert if __name__ == `main' stuff?
+Would be nice to have an executable...
+Move RT1D.py to \$RT1D/bin, import it from \$RT1D/rt1d/__init__.py
+
 \newpage

### 2011-12-22 rt1d/mods/__init__.py

@@ -6,4 +6,5 @@ from rt1d.mods.InitializeIntegralTables import InitializeIntegralTables
 from rt1d.mods.InitializeGrid import InitializeGrid
 from rt1d.mods.Radiate import Radiate
 from rt1d.mods.ReadParameterFile import *
+from rt1d.mods.CheckForParameterConflicts import *
 from rt1d.mods.WriteData import WriteData

### 2011-12-22 scripts/RT1D.py

@@ -50,9 +50,11 @@ else:
 if rank == 0: 
     print "\nStarting rt1d..."
     print "Initializing {0} 1D radiative transfer calculation(s)...".format(len(all_pfs)) 
-    if IsRestart: print "Restarting from {0}".format(sys.argv[2])
+    if IsRestart: 
+        print "Restarting from {0}".format(sys.argv[2])
     print "Press ctrl-C to quit at any time.\n" 
 
+# Start clock
 start = time.time()
 
 # Loop over parameter sets. 
@@ -60,7 +62,17 @@ for i, pf in enumerate(all_pfs):
         
     if pf["ParallelizationMethod"] == 2:
         if i % size != rank: continue
+        
+    # Check for parameter conflicts    
+    conflicts, errmsg = rtm.CheckForParameterConflicts(pf)
+    if conflicts:
+        print 'ERROR -- PARAMETER VALUES IN CONFLICT:'
+        for msg in errmsg:
+            print msg
+        print '\n'    
+        raise Exception()   
           
+    # Read in a few things for convenience      
     TimeUnits = pf["TimeUnits"]
     LengthUnits = pf["LengthUnits"]
     GridDimensions = pf["GridDimensions"]
@@ -69,9 +81,10 @@ for i, pf in enumerate(all_pfs):
     
     # Print out cell-crossing and box-crossing times for convenience
     if rank == 0:
-        print "Cell-crossing time = {0} (years), {1} (code)".format(LengthUnits / GridDimensions / 29979245800.0 / 31557600.0, \
-            LengthUnits / GridDimensions / 29979245800.0 / TimeUnits)
-        print "Box-crossing time = {0} (years), {1} (code)\n".format(LengthUnits / 29979245800.0 / 31557600.0, LengthUnits / 29979245800.0 / TimeUnits)
+        print "Cell-crossing time = {0} (years), {1} (code)".format(LengthUnits / GridDimensions / \
+            29979245800.0 / 31557600.0, LengthUnits / GridDimensions / 29979245800.0 / TimeUnits)
+        print "Box-crossing time = {0} (years), {1} (code)\n".format(LengthUnits / 29979245800.0 / \
+            31557600.0, LengthUnits / 29979245800.0 / TimeUnits)
                         
     # Initialize grid and file system
     if IsRestart: 
@@ -90,7 +103,7 @@ for i, pf in enumerate(all_pfs):
         
         # Wait here if parallelizing over grid
[…]

### 2012-01-18 rt1d/mods/Radiate.py
Separate module for computing rate coefficients.  Going to rework things and add a photon-conserving algorithm too.

@@ -54,7 +54,8 @@ E_th = [13.6, 24.6, 54.4]
 neglible_column = 1
 
 # Widget for progressbar.
-widget = ["Ray Casting: ", Percentage(), ' ', Bar(marker = RotatingMarker()), ' ', ETA(), ' ']
+if pb: 
+    widget = ["Ray Casting: ", Percentage(), ' ', Bar(marker = RotatingMarker()), ' ', ETA(), ' ']
 
 class Radiate:
     def __init__(self, pf, data, itabs, n_col): 

### 2012-01-18 rt1d/mods/SetDefaultParameterValues.py

@@ -22,7 +22,7 @@ def SetDefaultParameterValues():
        {
     
         # Override
-        "ProblemType": 0, \
+        "ProblemType": 1, \
           
         # Integration
         "InterpolationMethod": 0, \
@@ -38,10 +38,12 @@ def SetDefaultParameterValues():
         
         "ParallelizationMethod": 1, \
         "Debug": 1, \
+        
+        "PhotonConserving": 0, \
           
         # Integral tabulation
         "HIColumnMin": 1e16, \
-        "HIColumnMax": 1e22, \
+        "HIColumnMax": 1e20, \
         "HeIColumnMin": 1e18, \
         "HeIColumnMax": 1e22, \
         "HeIIColumnMin": 1e16, \
@@ -63,7 +65,7 @@ def SetDefaultParameterValues():
         "CurrentTime": 0.0, \
         "StopTime": 50.0, \
         "StartRadius": 0.01, \
-        "dtDataDump": 1.0, \
+        "dtDataDump": 5.0, \
         "DataDumpName": 'dd', \
         "OutputDirectory": './', \
         "OutputFormat": 1, \
@@ -81,7 +83,7 @@ def SetDefaultParameterValues():
         "MinimumSpeciesFraction": 1e-8, \
         
         # Physics
-        "Isothermal": 0, \
+        "Isothermal": 1, \
         "MultiSpecies": 0, \
         "CollisionalIonization": 1, \
         "CollisionalExcitation": 1, \
@@ -96,8 +98,8 @@ def SetDefaultParameterValues():
         "DensityProfile": 0, \
         "InitialDensity": 0, \
         "TemperatureProfile": 0, \
-        "InitialTemperature": 100, \
-        "IonizationProfile": 1, \
+        "InitialTemperature": 1e4, \
+        "IonizationProfile": 0, \
         "InitialHIIFraction": 1e-4, \
         "Clump": 0, \
         "ClumpPosition": 0.1, \
@@ -118,10 +120,10 @@ def SetDefaultParameterValues():
         # Spectral parameters
         "DiscreteSpectrum": 0, \
         "SpectrumPowerLawIndex": 1.5, \
-        "SpectrumMinEnergy": 100, \
-        "SpectrumMaxEnergy": 1e4, \
-        "SpectrumMinNormEnergy": 100, \
-        "SpectrumMaxNormEnergy": 1e4, \
[…]

### 2012-04-03 doc/manual/rt1d_manual.tex
Separate module for computing rate coefficients.  Going to rework things and add a photon-conserving algorithm too.
SecondaryIonization methods 0-3 all work and produce sensible results for continuous and discrete SEDs.

@@
 \newcommand{\nHII}{n_{\text{HII}}}  
+\newcommand{\nHe}{n_{\text{He}}}
+\newcommand{\xHI}{x_{\text{HI}}} 
+\newcommand{\xHII}{x_{\text{HII}}} 
+\newcommand{\xHeI}{x_{\text{HeI}}} 
+\newcommand{\xHeII}{x_{\text{HeII}}} 
+\newcommand{\xHeIII}{x_{\text{HeIII}}}  
 \newcommand{\nHeI}{n_{\text{HeI}}}
@@
 \newcommand{\nel}{n_{\text{e}}}  
-\newcommand{\nb}{n_{\text{b}}}
-\newcommand{\nnu}{$n_{\nu}$ }
-\newcommand{\ncol}{n_i^{\text{col}}}
+\newcommand{\ntot}{n_{\text{tot}}}
+\newcommand{\nnu}{$n_{\nu}$}
+\newcommand{\ncol}{N_i}
 \newcommand{\gHI}{\Gamma_{\text{HI}}}  
@@
 \newcommand{\bHeII}{\beta_{\text{HeII}}}  
-\newcommand{\xHeII}{\xi_{\text{HeII}}}
+\newcommand{\xiHeII}{\xi_{\text{HeII}}}
 \newcommand{\kB}{k_{\text{B}}}
-\newcommand{\fheat}{f_i^{\text{heat}}}
+\newcommand{\fheat}{f^{\text{heat}}}
+\newcommand{\fheatE}{f_{(\nu - \nu_i)}^{\text{heat}}}
 \newcommand{\fion}{f_i^{\text{ion}}}
+\newcommand{\fionE}{f_{ij,(\nu - \nu_j)}^{\text{ion}}}
 \newcommand{\Lbol}{\mathcal{L}_{\text{bol}}}
@@
 \newcommand{\Htwo}{\mathrm{H}_2}
-\newcommand{\eV}{\mathrm{eV}}
+\newcommand{\drIF}{$\Delta r_{\mathrm{IF}}$}
+\newcommand{\dTb}{$\delta T_b$}
+\newcommand{\Nvec}{\mathbf{N}}
+\newcommand{\sh}{\mathrm{sh}}
 %%% END custom commands
@@
 \section{Methods}
-Our radiative transfer scheme is modeled closely after the work of
-\cite{Fukugita1994} and \cite{Thomas2008}, though for completeness we will
-reiterate the aspects of these methods most pertinent to the problem at hand.
-In general, propagating a radiation field relies on solving differential
-equations governing the rate of change in the number densities of all ions and
-the temperature of the gas. Assuming a medium consisting of hydrogen and
-helium only, we first solve for the relative abundances of each ion via
+\subsection{The Rate Equations}
+In general, the chemical and thermal evolution of gas surrounding a radiation
+source is governed by a set of differential equations describing the number
+densities of all ions and the temperature of the gas. Assuming a medium
+consisting of hydrogen and helium only, we first solve for the abundances of
+each ion via
 \begin{align}
-    \frac{d \nHII}{dt} & = \gHI \nHI + \bHI \nel \nHI - \aHII \nel \nHII   \label{eq:HIIRateEquation} \\ 
-    \frac{d \nHeII}{dt} & = \gHeI \nHeI + \bHeI \nel \nHeI - \bHeII \nel \nHeII - \aHeII \nel \nHeII + \aHeIII \nel \nHeIII - \xHeII \nel \nHeIII  \label{eq:HeIIRateEquation} \\ 
-    \frac{d \nHeIII}{dt} & = \gHeII \nHeII + \bHeII \nel \nHeII - \aHeIII \nel \nHeIII . \label{eq:HeIIIRateEquation}
+    \frac{d \nHII}{dt} & = (\gHI + \gamma_{\HI} + \bHI \nel) \nHI - \aHII \nel \nHII   \label{eq:HIIRateEquation} \\ 
+    \frac{d \nHeII}{dt} & = (\gHeI + \gamma_{\HeI} + \bHeI \nel) \nHeI \nonumber + \aHeIII \nel \nHeIII \\  & - (\bHeII + \aHeII + \xiHeII) \nel \nHeII \label{eq:HeIIRateEquation} \\ 
+    \frac{d \nHeIII}{dt} & = (\gHeII + \gamma_{\HeII} + \bHeII \nel) \nHeII  - \aHeIII \nel \nHeIII . \label{eq:HeIIIRateEquation}
 \end{align}
-Each of these equations represents the balance of ionizations and
-recombinations for ions $\HII$, $\HeII$, $\HeIII$. We associate the index $i$
-with neutral species, $i = \HI, \HeI, \HeII$, and define $\Gamma_i$ as the
-photo-ionization rate coefficient, $\alpha_i$ ($\xi_i$) as the case-A
-(dielectric) recombination rate coefficients, $\beta_i$ as the collisional
-ionization rate coefficients, and $\nel = \nHII + \nHeII + 2\nHeIII$ as the number density of electrons.
+Each of these equations represents the balance between ionizations of species
+$\HI$, $\HeI$, and $\HeII$, and recombinations of $\HII$, $\HeII$, and
+$\HeIII$. Associating the index $i$ with absorbing species, $i = \HI, \HeI,
+\HeII$, we define $\Gamma_i$ as the photo-ionization rate coefficient,
+$\gamma_i$ as the secondary ionization rate coefficient, $\alpha_i$ ($\xi_i$)
+as the case-B (dielectric) recombination rate coefficients, $\beta_i$ as the
+collisional ionization rate coefficients, and $\nel = \nHII + \nHeII +
+2\nHeIII$ as the number density of electrons.
 
-In this work, we use the formulae in Appendix B of \cite{Fukugita1994} to
-compute the values of $\alpha_i$, $\beta_i$, and $\xi_i$.
-As you might expect, collisional ionization and recombination depend solely on the temperature of the gas.
-For example, collisional ionization of hydrogen (from $n = 1$) is governed by
+At each time-step, we also solve for the temperature evolution, $dT_k/dt$,
+which
+is given by
+\begin{align}
+    \frac{3}{2}\frac{d}{dt}\left(\frac{\kB T_k \ntot}{\mu}\right) & = \fheat  \sum_i n_i \Heat_i - \sum_i \zeta_i n_e n_i - \sum_i \eta_i n_e n_i \nonumber \\ & - \sum_i \psi_i n_e n_i - \omega_{\HeII} n_e \nHeII \label{eq:TemperatureEvolution}
+\end{align}
+where $\Heat_i$ is the photo-electric heating rate coefficient (due to
+electrons previously bound to species $i$), $\omega_{\HeII}$ is the dielectric
+recombination cooling coefficient, and $\zeta_i$, $\eta_i$, and $\psi_i$ are
+the collisional ionization, recombination, and collisional excitation cooling
+coefficients, respectively. The constants in Equation
+\ref{eq:TemperatureEvolution} are the total number density of baryons, $\ntot
+= n_\mathrm{H} + n_{\mathrm{He}} + n_e$, the mean molecular weight, $\mu$,
+Boltzmann's constant, $\kB$, and the fraction of secondary electron energy
+deposited as heat, $\fheat$. We use the formulae in Appendix B of
+\citet{Fukugita1994} to compute the values of $\alpha_i$, $\beta_i$, $\xi_i$,
+$\zeta_i$, $\eta_i$, $\psi_i$, and $\omega_{\HeII}$.
+
+% Photon-conservation
+\subsection{Computing the Ionization and Heating Rates}
+The most critical aspect of propagating the radiation field in our 1D
+simulations is computing the ionization ($\Gamma_i$, $\gamma_{i}$) and heating
+($\Heat_i$) rate coefficients accurately. In order to directly relate our
+results to fully 3-dimensional radiative transfer calculations we have chosen
+to adopt a photon-conserving (PC) algorithm nearly identical to those employed
+by several widely used codes, like \textit{C$^2$Ray} \citep{Mellema2006}, and
+\textit{Enzo} \citep{Wise2011}. Our code is able to compute $\Gamma_i$,
+$\gamma_{i}$, and $\Heat_i$ in a non-photon-conserving (NPC) fashion as well,
+to enable comparison with previous 1D work such as \citet{Thomas2008}. The two
+formalisms are equivalent in the limit of very optically thin cells, a
+condition that can be met easily in 1D calculations but is rarely
+computationally feasible in 3D. For NPC methods, if the optical depth of an
+individual cell is substantial, the number of ionizations in that cell will
+\textit{not} equal the number of photons absorbed for that cell, i.e. photon
+number will not be conserved. This problem was remedied by \citet{Abel1999},
+who inferred the number of photo-ionizations of species $i$ in a cell from the
+radiation incident upon it and its optical depth,
 \begin{equation}
-    \beta_{1,\HI} = 5.85 \times 10^{-11} T^{1/2} \left[1 + \left(\frac{T}{10^5 \mathrm{K}}\right)^{1/2}\right]^{-1} \mathrm{exp}\left[-1.578 \times \left(\frac{T}{10^5 \mathrm{K}}\right)^{-1}\right] \mathrm{cm^3 s^{-1}}
+    \Delta \tau_{i,\nu} = n_i \sigma_{i,\nu} \Delta r .
 \end{equation}    
-and recombinations (to $n \ge 1$) by
-\begin{equation}
-    \alpha_{\HII} = 6.28 \times 10^{-11} T^{-1/2} \left(\frac{T}{10^3\mathrm{K}}\right)^{-1/5} \left[1 + \left(\frac{T}{10^6\mathrm{K}}\right)^{0.7}\right]^{-1} \mathrm{cm^3s^{-1}}
-\end{equation}
+It is most straightforward to imagine our 1D grid as a collection of
+concentric spherical shells, each having thickness $\Delta r$ and volume
+$V_{\sh}(r) = 4 \pi [(r + \Delta r)^3 - r^3] / 3$. The ionization and heating
+rates can then be related to the number of absorptions in any given shell
+(thus preserving photon number), as
+\begin{align}
+    \Gamma_i & = A_i \int_{\nu_i}^{\infty} I_{\nu} e^{-\tau_{\nu}} \left(1 - e^{-\Delta \tau_{i,\nu}}\right) \frac{d\nu}{h\nu} \label{eq:PhotoIonizationRate} \\
+    \gamma_{ij} & = A_j \int_{\nu_j}^{\infty} \left(\frac{\nu - \nu_j}{\nu_i}\right) I_{\nu} e^{-\tau_{\nu}} \left(1 - e^{-\Delta \tau_{j,\nu}}\right) \frac{d\nu}{h\nu} \label{eq:SecondaryIonizationRate} \\
+    \Heat_i & = A_i \int_{\nu_i}^{\infty} (\nu - \nu_i) I_{\nu} e^{-\tau_{\nu}} \left(1 - e^{-\Delta \tau_{i,\nu}}\right) \frac{d\nu}{\nu} , \label{eq:HeatingRate}
+\end{align}    
+where we've defined the normalization constant, $A_i \equiv
+L_{\mathrm{bol}}/n_i V_{\sh}(r)$, and denote the ionization threshold energy
+for species $i$ as $h\nu_i$.
 
-We absorb ionizations due to to energetic photo-electrons into the
-photoionization rate coefficients, $\Gamma_i$, which are given by
-\begin{equation}
-    \Gamma_i = \int_{E_i}^{\infty} \sigma_i(E) \spec(E) \frac{dE}{E} + \sum_j \fion \left(\frac{n_j}{n_i}\right) \int_{E_i}^{\infty} \sigma_i(E)\left(\frac{E - E_j}{E_i}\right) \spec(E) \frac{dE}{E} \label{eq:PhotoionizationRate}
-\end{equation}
-where $E_i$, $\sigma_i$, $n_i$, and $f_i$ are the ionization threshold energy,
-bound-free absorption cross section, number density, and fraction of
-photo-electron energy that goes into ionization of species $i$, respectively.
-The term summing over $j$, where $j = \HI, \HeI, \HeII$, represents
-ionizations of species $i$ due to fast secondary electrons from
-photoionizations of species $j$, which has number density $n_j$. $\spec$ is
-the spectral intensity, which we write as
-\begin{equation}
-    \spec(E, r) = \frac{\Lbol}{4\pi r^2} I(E) e^{-\tau(E, r)} . \label{eq:Spectrum}
-\end{equation}
-Here, $r$ is the distance from the source, $I(E)$ is the normalized spectral
-energy distribution (SED), satisfying
-\begin{equation}
-    \int_E I(E) dE = 1 ,
-\end{equation}
-and the optical depth $\tau(E, r)$ is given as a sum over absorbing species,
+Equation \ref{eq:SecondaryIonizationRate} represents ionizations of species
+$i$ due to fast secondary electrons from photoionizations of species $j$,
+which has number density $n_j$, and ionization threshold energy, $h\nu_j$.
+$\fion$ is the fraction of photo-electron energy deposited as ionizations of
+species $i$. In the remaining sections we only include the effects of
+secondary electrons when considering X-ray sources, which emit photons in the
+range $10^2\mathrm{eV} < E < 10^4\mathrm{eV}$. In this regime, the values of
+$\fheat$ and $\fion$ computed via the formulae of \citet{Shull1985} are
+sufficiently accurate, but for radiation at lower energies where $\fheat$ and
+$\fion$ have a stronger energy dependence, the lookup tables of
+\citet{Furlanetto2010} would be more appropriate. The total secondary
+ionization rate for a given species, $\gamma_i$, is the sum of ionizations due
+to the secondary electrons from all species, $\gamma_i = \fion \sum_j
+\gamma_{ij} n_j / n_i$.
+
+The optical depth, $\tau_{\nu} = \tau_{\nu}(r)$, in the above equations is
+the total optical depth at frequency $\nu$ due to all absorbing species, i.e.
 \begin{align}
-    \tau(E, r) & = \sum_i \int_r \sigma_i(E) n_i(r) dr \nonumber \\
-               & = \sum_i \sigma_i(E) \ncol(r) \label{eq:OpticalDepth}
+    \tau_{\nu}(r) & = \sum_i \int_0^r \sigma_{i,\nu} n_i(r^{\prime}) dr^{\prime} \nonumber \\
+               & = \sum_i \sigma_{i,\nu} \ncol(r) \label{eq:OpticalDepth}
 \end{align}
-where $\ncol(r)$ is the column density of species $i$ at distance $r$ from the
-source. We calculate the bound-free absorption cross-sections, $\sigma_i$,
-using the fits of \cite{Verner1996}.
+where $\ncol$ is the column density of species $i$ at distance $r$ from the
+source. We calculate the bound-free absorption cross-sections using the fits
+of \citet{Verner1996} throughout.
 
-To be more explicit, let's write out $\Gamma_{\HI}$ in its entirety.
+The values of $\Gamma_i$, $\gamma_i$, and $\Heat_i$ are completely
+predetermined for a given radiation source, and as a result, can be tabulated
+as a function of column density to avoid evaluating the
+integrals in these expressions numerically `on-the-fly' as a simulation runs
+\citep[e.g.][]{Mellema2006,Thomas2008}. Isolating the frequency dependent
+components of Equations \ref{eq:PhotoIonizationRate}-\ref{eq:HeatingRate}, we
+can define the integrals
 \begin{align}
-    \Gamma_{\HI} & = \int_{13.6\eV}^{\infty}\sigma_{\HI}(E) \spec(E) \frac{dE}{E} \nonumber \\
-    & + f_{\HI}^{\mathrm{ion}} \left(\frac{n_{\HI}}{n_{\HI}}\right) \int_{13.6\eV}^{\infty}\sigma_{\HI}(E) \left(\frac{E - 13.6\eV}{13.6\eV}\right)\spec(E)\frac{dE}{E} \nonumber \\
-    & + f_{\HeI}^{\mathrm{ion}} \left(\frac{n_{\HeI}}{n_{\HI}}\right) \int_{24.4\eV}^{\infty}\sigma_{\HeI}(E) \left(\frac{E - 24.4\eV}{13.6\eV}\right)\spec(E)\frac{dE}{E} \nonumber \\
-    & + f_{\HeII}^{\mathrm{ion}} \left(\frac{n_{\HeII}}{n_{\HI}}\right) \int_{54.4\eV}^{\infty}\sigma_{\HeI}(E) \left(\frac{E - 54.4\eV}{13.6\eV}\right)\spec(E)\frac{dE}{E}
+    \Phi_i (\tau_{\nu}) & \equiv \int_{\nu_i}^{\infty} I_{\nu} e^{-\tau_{\nu}} \frac{d\nu}{h\nu} \label{eq:PHI} \\
+    \Psi_i (\tau_{\nu}) & \equiv \int_{\nu_i}^{\infty} I_{\nu} e^{-\tau_{\nu}} d\nu \label{eq:PSI},
 \end{align}
-The first term accounts for all hydrogen photoionization.
-The final three terms account for collisional ionization of hydrogen due to secondary photo-electrons from ionizations of neutral hydrogen (second term), neutral helium (third term), and singly ionized helium (fourth term).
-Collisional ionizations due to thermal electrons in the gas are accounted for separately through the $\beta_i$ coefficients.
+allowing us to re-express the rate coefficients as
+\begin{align}
+    \Gamma_i & = A_i \left[\Phi_i(\tau_{\nu}) - \Phi_i(\tau_{i,\nu}^{\prime}) \right] \label{eq:Gamma_PhiPsi} \\
+    \gamma_{ij} & = \frac{A_j}{h\nu_i} \left\{\Psi_j(\tau_{\nu}) - \Psi_j(\tau_{j,\nu}^{\prime}) - h \nu_j \left[\Phi_j(\tau_{\nu}) - \Phi_j(\tau_{j,\nu}^{\prime}) \right] \right\}  \label{eq:gamma_PhiPsi} \\
+    \Heat_i & = A_i \left\{\Psi_i(\tau_{\nu}) - \Psi_i(\tau_{i,\nu}^{\prime}) - h\nu_i \left[\Phi_i(\tau_{\nu}) - \Phi_i(\tau_{i,\nu}^{\prime}) \right] \right\} \label{eq:Heat_PhiPsi},  
+\end{align}
+where $\tau_{i,\nu}^{\prime} \equiv \tau_{\nu} + \Delta \tau_{i, \nu}$. Later
+references to ``continuous SEDs'' signify use of this technique, where the
+integral values $\Phi_i$ and $\Psi_i$ are computed over a column density
+interval of interest a priori using a Gaussian quadrature technique, rather
+than on-the-fly via discrete summation.
 
-At each time-step, we also solve for the temperature evolution, $dT/dt$, which
-is given by
-\begin{equation}
-    \frac{3}{2}\frac{d}{dt}\left(\frac{\kB T \nb}{\mu}\right) = \fheat \sum_i n_i \Heat_i - \sum_i \zeta_i n_e n_i - \sum_i \eta_i n_e n_i - \sum_i \psi_i n_e n_i \label{eq:TemperatureEvolution} 
-\end{equation}
-where $\Heat_i$ is the photo-electric heating rate, given by
-\begin{equation}
-    \Heat_i = \int_{E_i}^{\infty} \sigma_i(E) \spec(E, r)(E - E_i) \frac{dE}{E} \label{eq:HeatingRate} ,
-\end{equation}    
-and $\zeta_i$, $\eta_i$, and $\psi_i$ are the collisional ionization,
-recombination, and collisional excitation cooling coefficients, respectively.
-The constants in Equation \ref{eq:HeatingRate} are the total number density of
-baryons, $\nb$, the mean molecular weight, $\mu$, Boltzmann's constant, $\kB$,
-and the fraction of a photon's original energy deposited as heat, $\fheat$. In
-the remaining sections we only include the effects of secondary electrons when
-considering X-ray sources, which emit photons in the range $10^2\mathrm{eV} <
-E < 10^4\mathrm{eV}$. In this regime, the values of $\fheat$ and $\fion$
-computed via the formulae of \cite{Shull1985} are sufficiently accurate, but
-for radiation at lower energies, the lookup tables of \cite{Furlanetto2010}
-would be more appropriate.
+Equations \ref{eq:Gamma_PhiPsi}-\ref{eq:Heat_PhiPsi} are completely general
+for photon-conserving algorithms, whether the source SEDs are discrete or
+continuous -- the only difference being for discrete SEDs, the integrals in
+Equations \ref{eq:PHI}-\ref{eq:PSI} become sums over the number of frequencies
+used, $n_{\nu}$. In practice, computing $\Gamma_i$, $\gamma_i$, and $\Heat_i$
+is more straightforward than this for sources with discrete SEDs, as we can
+simply count the number of ionizations caused by each individual frequency
+group, and convert this into the amount of excess electron kinetic energy
+available for further heating and ionization. When testing the accuracy of
+discrete solutions in later sections we employ this method, where radiation is
+emitted at $n_{\nu}$ frequencies, with each frequency $\nu_n$ carrying a
+fraction $I_n$ of the bolometric luminosity. The photoionization and heating
+coefficients for each individual frequency group can then be expressed as
+\begin{align}
+    \Gamma_{i,n} & = \frac{A_i I_n}{h \nu_n} e^{-\tau_{\nu_n}}(1 - e^{-\Delta \tau_{i,\nu_n}}) \label{eq:Gamma_simple}\\
+    \gamma_{ij,n} & = \Gamma_{j,\nu_n} (\nu_n-\nu_j) / \nu_i \label{eq:gamma_simple} \\
+    \Heat_{i,n} & = \Gamma_{i,\nu_n} h(\nu_n - \nu_i) \label{eq:Heat_simple}.
+\end{align}    
+The total rate coefficients can be found by summing each of these expressions
+over all frequencies, $n=1,2,3...n_{\nu}$. These equations are identical to
+Equations \ref{eq:Gamma_PhiPsi}-\ref{eq:Heat_PhiPsi} for the discrete SED
+case, but are perhaps more intuitive.
 
-An important simplification is that all radial and time dependencies can be
-pulled outside of the integrals in Equations \ref{eq:PhotoionizationRate} and
-\ref{eq:HeatingRate}, leaving (ignoring the secondary ionization term)
-\begin{equation}
-    \Gamma_i(n^{\mathrm{col}}) = A \int_{E_i}^{\infty} \sigma_i(E) I(E) e^{-\tau(E, n^{\mathrm{col}})} \frac{dE}{E} \label{eq:Gamma_final} ,
-\end{equation}
-and
-\begin{equation}
-    \Heat_i(n^{\mathrm{col}}) = A \int_{E_i}^{\infty} \sigma_i(E) I(E) (E - E_i) e^{-\tau(E, n^{\mathrm{col}})} \frac{dE}{E} \label{eq:Heat_final}
-\end{equation}
-where $A \equiv \Lbol / 4 \pi r^2$, and $n^{\mathrm{col}} =
-(n_{\mathrm{HI}}^{\mathrm{col}}, n_{\mathrm{HeI}}^{\mathrm{col}},
-n_{\mathrm{HeII}}^{\mathrm{col}})$.
+For simplicity, our current treatment neglects a few physical processes that
+are cosmological in origin, or simply do not rely on the radiation field
+directly. These include cooling via free-free emission and hydrogen ionization
+due to helium recombination photons (which depend on the gas kinetic
+temperature and electron density), and cosmological effects such as Hubble
+cooling, Compton cooling off cosmic microwave background (CMB) photons, and
+photo-ionization by Wien-tail CMB photons (which depend on kinetic
+temperature, redshift, and the Hubble parameter). Our optimized SEDs
+(\S\ref{sec:Results}) will be unchanged upon including such physics in later
+work.
 
-From Equations \ref{eq:Gamma_final} and \ref{eq:Heat_final} it is clear that
-the values of $\Gamma_i$ and $\Heat_i$ can be tabulated as a function of
-column density for a given source. This greatly improves the speed of 1D
-radiative transfer calculations, like those of \cite{Thomas2008}, since
-integral values in Equations \ref{eq:HIIRateEquation},
-\ref{eq:HeIIRateEquation}, \ref{eq:HeIIIRateEquation}, and
-\ref{eq:HeatingRate} no longer need to be computed numerically on-the-fly. It
-also provides the basis for our frequency resolution optimization strategy, as
-presented in \S\ref{sec:Methods}. Note, however, that the dimensionality of
-these lookup tables is equal to the number of absorbing species, so the tables
-for simulations including only hydrogen are 1D, while those including helium
-are 3D. If we choose to adopt the secondary electron treatment of
-\cite{Furlanetto2010} our lookup tables inherit an additional dimension, as
-the secondary ionization and heating factors $\fion$ and $\fheat$ depend on
-both energy and ionized fraction, $x_i$. Since integration takes place over
-energy, $x_i$ becomes the final dimension of the lookup tables.
+\subsection{Additional Complexity: Energy Dependent Secondary Electron Treatment}
+In the previous section (and in all of Mirocha et al.
+2012), we considered the asymptotic limit of \citet{Shull1985}, in which the fractional energy deposition of secondary electrons as heat, ionization, and excitation depends only on the hydrogen ionized fraction.
+\citet{Furlanetto2010} updated this work, providing lookup tables of deposition fraction as a function of both ionized fraction and electron energy.
+These effects are most important at $E \lesssim 10^2$ eV.
 
-For simplicity, our current treatment neglects cooling via free-free emission,
-as well as cosmological effects, including Hubble cooling due to cosmological
-expansion, Compton heating/cooling off cosmic microwave background (CMB)
-photons, and photo-ionization by Wien-tail CMB photons.
+To accommodate this model, we must redefine the quantities $\Phi_i$ and $\Psi_i$.  We will use a tilde to denote the new quantities related to secondary ionization, 
+\begin{align}
+    \widetilde{\Phi}_{ij} & \equiv \int_{\nu_j}^{\infty} \fionE I_{\nu} e^{-\tau_{\nu}} \frac{d\nu}{h\nu} \\
+    \widetilde{\Psi}_{ij} & \equiv \int_{\nu_j}^{\infty} \fionE I_{\nu} e^{-\tau_{\nu}} d\nu ,
+\end{align}
+and a hat to denote the new quantities related to photo-electric heating,
+\begin{align}
+    \widehat{\Phi}_i & \equiv \int_{\nu_i}^{\infty} \fheatE I_{\nu} e^{-\tau_{\nu}} \frac{d\nu}{h\nu} \\
+    \widehat{\Psi}_i & \equiv \int_{\nu_i}^{\infty} \fheatE I_{\nu} e^{-\tau_{\nu}} d\nu .
+\end{align}
+Now, our ionization and heating coefficients read
+\begin{align}
+    \Gamma_i & = A_i \left[\Phi_i(\tau_{\nu}) - \Phi_i(\tau_{i,\nu}^{\prime}) \right] \label{eq:Gamma_PhiPsi_complex} \\
+    \gamma_{ij} & = \frac{A_j}{h\nu_i} \left\{\widetilde{\Psi}_{ij}(\tau_{\nu}) - \widetilde{\Psi}_{ij}(\tau_{j,\nu}^{\prime}) - h \nu_j \left[\widetilde{\Phi}_{ij}(\tau_{\nu}) - \widetilde{\Phi}_{ij}(\tau_{j,\nu}^{\prime}) \right] \right\}  \label{eq:gamma_PhiPsi_complex} \\
+    \Heat_i & = A_i \left\{\widehat{\Psi}_i(\tau_{\nu}) - \widehat{\Psi}_i(\tau_{i,\nu}^{\prime}) - h\nu_i \left[\widehat{\Phi}_i(\tau_{\nu}) - \widehat{\Phi}_i(\tau_{i,\nu}^{\prime}) \right] \right\} \label{eq:Heat_PhiPsi_complex} .
+\end{align}
 
@@
 
-\item [\texttt{ProblemType = 2.1}] Same as \texttt{ProblemType = 2} for everything, except we sample the continuous blackbody spectrum with the four energy groups used by Wise \& Abel 2010.
+\item [\texttt{ProblemType = 2.1}] Same as \texttt{ProblemType = 2} for everything, except we sample the continuous blackbody spectrum with the four energy groups used by \citet{Wise2011}
+
+\item [\texttt{ProblemType = 2.2}] Same as \texttt{ProblemType = 2} for everything, except we sample the continuous blackbody spectrum with the four energy groups used by Mirocha et al. 2012.
 
@@
 
-How to change the code to allow for this?
-Define main loop in RT1D.py as Shine.  Insert if __name__ == `main' stuff?
-Would be nice to have an executable...
-Move RT1D.py to \$RT1D/bin, import it from \$RT1D/rt1d/__init__.py
+%How to change the code to allow for this?
+%Define main loop in RT1D.py as Shine.  Insert if __name__ == `main' stuff?
+%Would be nice to have an executable...
+%Move RT1D.py to \$RT1D/bin, import it from \$RT1D/rt1d/__init__.py
+
+What parameters must be set?
+\begin{itemize}
+    \item ProblemType
+    \item GridDimensions
+    \item MultiSpecies
+    \item LengthUnits
+    \item InitialDensity
+    \item SourceParameters
+\end{itemize}
+
+%In this work, we use the formulae in Appendix B of \cite{Fukugita1994} to
+%compute the values of $\alpha_i$, $\beta_i$, and $\xi_i$.
+As you might expect, collisional ionization and recombination depend solely on the temperature of the gas.
+For example, collisional ionization of hydrogen (from $n = 1$) is governed by
+%\begin{equation}
+%    \beta_{1,\HI} = 5.85 \times 10^{-11} T^{1/2} \left[1 + \left(\frac{T}{10^5 \mathrm{K}}\right)^{1/2}\right]^{-1} \mathrm{exp}\left[-1.578 \times \left(\frac{T}{10^5 \mathrm{K}}\right)^{-1}\right] \mathrm{cm^3 s^{-1}}
+%\end{equation}
+%and recombinations (to $n \ge 1$) by
+%\begin{equation}
+%    \alpha_{\HII} = 6.28 \times 10^{-11} T^{-1/2} \left(\frac{T}{10^3\mathrm{K}}\right)^{-1/5} \left[1 + \left(\frac{T}{10^6\mathrm{K}}\right)^{0.7}\right]^{-1} \mathrm{cm^3s^{-1}}
+%\end{equation}
 
@@
 \bibliography{references}
-\bibliographystyle{plain}
+\bibliographystyle{apj}
 

### 2012-04-03 rt1d/analysis/Analysis.py

@@ -23,7 +23,7 @@ from ..mods.ComputeCrossSections import PhotoIonizationCrossSection
 from ..mods.InitializeIntegralTables import InitializeIntegralTables
 
 class Analyze:
-    def __init__(self, pf, retabulate = False):
+    def __init__(self, pf, retabulate = True):
         self.ds = Dataset(pf)
         self.data = self.ds.data
         self.pf = self.ds.pf    # dictionary
@@ -59,7 +59,7 @@ class Analyze:
         
         # Read integral table if it exists.
         self.tname = self.iits.DetermineTableName()
-        if os.path.exists('%s' % self.tname) and retabulate:
+        if os.path.exists('%s/%s' % (self.ds.od, self.tname)) and retabulate:
             self.itabs = self.iits.TabulateRateIntegrals()
             self.interp = Interpolate(self.pf, [self.iits.HIColumn, self.iits.HeIColumn, self.iits.HeIIColumn], 
                 self.itabs)
@@ -158,7 +158,7 @@ class Analyze:
                 continue
             
             exec('self.ax.loglog(self.data[%i].r / self.pf[\'LengthUnits\'], \
-                self.data[%i].T, ls = \'%s\', color = \'%s\')' % (dd, dd, '-', color))                
+                self.data[%i].T, ls = \'%s\', color = \'%s\')' % (dd, dd, ls, color))                
             
         self.ax.set_xscale(xscale)    
         self.ax.set_xlabel(r'$r / L_{\mathrm{box}}$') 
@@ -429,9 +429,7 @@ class Analyze:
         # Save heating and cooling rates
         self.heat = heat
         self.cool = cool
-        
-    
-                    
+            
     def ComputeDistributionFunctions(self, field, normalize = True, bins = 20, volume = False):
         """
         Histogram all fields.

### 2012-04-03 rt1d/analysis/Inspection.py

@@ -24,7 +24,10 @@ class Inspect:
         self.itabs = self.anl.itabs
         self.interp = self.anl.interp
         
-    def InspectIntegralTable(self, integral = 0, species = 0, nHI = None, nHeI = 0.0, nHeII = 0.0,
+        self.MultiSpecies = self.pf["MultiSpecies"]
+        
+    def InspectIntegralTable(self, intnum = 1, species = 0, donor_species = 0,
+        nHI = None, nHeI = 0.0, nHeII = 0.0,
         color = 'k', ls = '-', annotate = True):
         """
         Plot integral values...or something.
@@ -38,16 +41,28 @@ class Inspect:
             s = 'HeII'
         
         # Convert integral int to string
-        if integral == 0:
-            integral = 'PhotoIonizationRate%i' % species 
+        if intnum == 1:
+            integral = 'Phi%i' % species 
             ylabel = r'$\Phi_{\mathrm{%s}}$' % s
-        elif integral == 1:
-            integral = 'ElectronHeatingRate%i' % species
+        elif intnum == 1.1:
+            integral = 'PhiWiggle%i%i' % (species, donor_species)
+            ylabel = r'$\widetilde{\Phi}_{\mathrm{%s}}$' % s 
+        elif intnum == 1.2:
+            integral = 'PhiHat%i' % species
+            ylabel = r'$\widetilde{\Phi}_{\mathrm{%s}}$' % s        
+        elif intnum == 2:
+            integral = 'Psi%i' % species
             ylabel = r'$\Psi_{\mathrm{%s}}$' % s
+        elif intnum == 2.1:
+            integral = 'PsiWiggle%i%i' % (species, donor_species)
+            ylabel = r'$\widetilde{\Psi}_{\mathrm{%s}}$' % s 
+        elif intnum == 2.2:
+            integral = 'PsiHat%i' % species
+            ylabel = r'$\widetilde{\Psi}_{\mathrm{%s}}$' % s            
         else:
             integral = 'TotalOpticalDepth' 
             ylabel = r'$\sum_i\int_{\nu} \tau_{i,\nu} d\nu$'
-            pl.rcParams['figure.subplot.left'] = 0.2
+            pl.rcParams['figure.subplot.left'] = 0.2012
         
         # Figure out axes
         if nHI is None:
@@ -57,8 +72,13 @@ class Inspect:
                 i2 = np.argmin(np.abs(self.itabs['HeIIColumnValues_z'] - nHeII))
[…]

### 2012-04-03 rt1d/mods/ComputeRateCoefficients.py

@@ -113,8 +113,8 @@ class RateCoefficients:
                 
                 # A few quantities that are better to just compute once   
                 nout = np.log10(10**ncol + ncell * self.mask[i])
-                indices_out = self.Interpolate.GetIndices3D(nout)
-                                                           
+                indices_out = self.Interpolate.GetIndices(nout, np.log10(x_HII))
+                                                                                                                      
                 # Photo-Ionization - keep integral values too to be used in heating/secondary ionization
                 Gamma[i], Phi_N[i], Phi_N_dN[i] = self.PhotoIonizationRate(species = i, indices_in = indices,  
                     Lbol = Lbol, r = r, ncol = ncol, nabs = nabs, dr = dr,
@@ -141,17 +141,23 @@ class RateCoefficients:
                     psi[i] = self.CollisionalExcitationCoolingRate(species = i, T = T, nabs = nabs, nion = nion)    
             
             # Secondary ionization - do separately to take advantage of already knowing Psi and Phi
+            # Unless SecondaryIonization = 2 -- then knowing Phi and Psi for Gamma won't matter
             if self.SecondaryIonization:
+                
                 for i in xrange(3):
                     if not self.MultiSpecies and i > 0:
                         continue  
                     
                     # Ionizing species i with electrons from species j                        
                     for j in self.donors:
+                        if not self.MultiSpecies and j > 0:
+                            continue
+                            
                         gamma[i][j] += self.SecondaryIonizationRate(Psi_N = Psi_N[j], Psi_N_dN = Psi_N_dN[j],
                             Phi_N = Phi_N[j], Phi_N_dN = Phi_N_dN[j], 
                             Lbol = Lbol, r = r, ncol = ncol, nabs = nabs, dr = dr,
-                            species = i, donor_species = j, x_HII = x_HII, A = A[j])
+                            species = i, donor_species = j, x_HII = x_HII, A = A[j],
+                            indices_in = indices, indices_out = indices_out, nout = nout)
                                     
                     gamma[i] /= (E_th[i] * erg_per_ev)
                                                         
@@ -199,7 +205,8 @@ class RateCoefficients:
                     if self.SecondaryIonization:
                         
[…]

### 2012-04-03 rt1d/mods/ControlSimulation.py

@@ -69,10 +69,9 @@ class ControlSimulation:
         nout = np.log10(ncell)
         Vsh = r.coeff.ShellVolume(self.R0, r.dx[0])    
         Lbol = r.rs.BolometricLuminosity(0)
+        logxHII = np.log10(nion[0] / n_H)
                                                                     
-        indices_in = None
-        if self.pf['MultiSpecies'] > 0 and self.pf['TabulateIntegrals']: 
-            indices_in = r.coeff.Interpolate.GetIndices3D(ncol)  
+        indices_in = r.coeff.Interpolate.GetIndices(ncol, logxHII)
 
         Gamma = r.coeff.ConstructArgs([nabs, nion, n_H, n_He, n_e], 
             indices_in, Lbol, self.R0, ncol, T, r.dx[0], 0.)[0]        

### 2012-04-03 rt1d/mods/ProblemTypes.py

@@ -59,7 +59,6 @@ def ProblemType(ptype):
              "LengthUnits": 6.6 * cm_per_kpc,
              "StopTime": 500.0,
              "dtDataDump": 10.0, 
-             "ColumnDensityBinsHI": 500, 
              "MinimumSpeciesFraction": 1e-6, 
              "DensityProfile": 0, 
              "InitialDensity": 1e-3, 

### 2012-04-03 rt1d/mods/SecondaryElectrons.py

@@ -24,7 +24,7 @@ class SecondaryElectrons:
         self.NumberOfEnergyBins = 258
         self.NumberOfXiBins = 14
         
-        if self.Method == 2:
+        if self.Method >= 2:
             rt1d = os.environ.get("RT1D")
             if rt1d:
                 f = h5py.File("{0}/input/secondary_electron_data.h5".format(rt1d), 'r')
@@ -34,12 +34,19 @@ class SecondaryElectrons:
                 raise Exception('Error loading secondary electron data.')
                 
             # Read in Furlanetto & Stoever lookup tables (not yet implemented)
-            self.Energies = f["ElectronEnergy"].value
-            self.IonizedFractions = f["IonizedFraction"].value
-            self.fheat = f["Heat"].value
-            self.fion_HI = f["IonizationHI"].value
-            self.fion_HeI = f["IonizationHeI"].value
-            self.fion_HeII = f["IonizationHeII"].value
+            self.Energies = f["electron_energy"].value
+            self.IonizedFractions = f["ionized_fraction"].value
+            self.LogIonizedFractions = np.log10(self.IonizedFractions)
+            self.fheat = f["f_heat"].value
+            self.fion_HI = f["fion_HI"].value
+            self.fion_HeI = f["fion_HeI"].value
+            self.fion_HeII = f["fion_HeII"].value
+            self.fexc = f['fexc'].value
+            self.f_Lya = f['f_Lya'].value
+            self.fion = f['fion'].value
+            
+            self.xmin = min(self.IonizedFractions)
+            self.xmax = max(self.IonizedFractions)
             
             f.close()        
         
@@ -51,17 +58,16 @@ class SecondaryElectrons:
             channel = 0: heat
             channel = 1: ionization of HI
             channel = 2: ionization of HeI
-            channel = 3: ionization of HeII (not implemented)
+            channel = 3: ionization of HeII
+            channel = 4: lyman-alpha excitation
             
         and
         
-            Method = 0: OFF - all secondary electron energy goes to heat
-            Method = 1: Empirical fits of Shull & vanSteenberg 1985
-            Method = 2: Lookup tables of Furlanetto & Stoever 2010
+            Method = 0: OFF - all secondary electron energy goes to heat.
+            Method = 1: Empirical fits of Shull & vanSteenberg 1985.
+            Method = 2: Lookup tables of Furlanetto & Stoever 2010.
+            Method = 3: Empirical Fits of Ricotti et al. 2002.
             
[…]

### 2012-04-03 rt1d/mods/__init__.py

@@ -2,10 +2,10 @@
 Import everything.
 """
 
-from rt1d.mods.InitializeIntegralTables import InitializeIntegralTables
-from rt1d.mods.InitializeGrid import InitializeGrid
-from rt1d.mods.Radiate import Radiate
-from rt1d.mods.ReadParameterFile import *
-from rt1d.mods.CheckForParameterConflicts import *
-from rt1d.mods.WriteData import WriteData
-from rt1d.mods.SecondaryElectrons import SecondaryElectrons
+from .InitializeIntegralTables import InitializeIntegralTables
+from .InitializeGrid import InitializeGrid
+from .Radiate import Radiate
+from .ReadParameterFile import *
+from .CheckForParameterConflicts import *
+from .WriteData import WriteData
+from .SecondaryElectrons import SecondaryElectrons

### 2012-04-15 doc/manual/rt1d_manual.tex
General cleanup + test problem 0 + recombination cooling bug fix + other stuff.

@@
 \usepackage{pslatex}
+\usepackage{epsfig}
 \usepackage{natbib}
@@
 \usepackage[all]{hypcap}
-
-%%% Beginning of Document
-
-\begin{document}
 
@@
 
+%%% Beginning of Document
+\begin{document}
 \author{Jordan Mirocha}	
-
 \title{\Large {\bf rt1d: User's Manual}}
@@
 \section{Introduction}
-Let's see what happens to the gas around stars and black holes, shall we?
+Let's see what happens to the gas around stars and black holes, shall we? The
+primary reference for this code is
+\href{http://arxiv.org/abs/1204.1944}{arxiv:1204.1944} -- much of the text has
+been extracted from there.
 
@@
 \section{Methods}
+
+% Rate Equations
 \subsection{The Rate Equations}
@@
 
-% Photon-conservation
+% Ionization and Heating Rates
 \subsection{Computing the Ionization and Heating Rates}
@@
 concentric spherical shells, each having thickness $\Delta r$ and volume
-$V_{\sh}(r) = 4 \pi [(r + \Delta r)^3 - r^3] / 3$. The ionization and heating
-rates can then be related to the number of absorptions in any given shell
-(thus preserving photon number), as
+$V_{\sh}(r) = 4 \pi [(r + \Delta r)^3 - r^3] / 3$ (i.e. radii correspond to
+cell interfaces). The ionization and heating rates can then be related to the
+number of absorptions in any given shell (thus preserving photon number), as
 \begin{align}
@@
 
-For simplicity, our current treatment neglects a few physical processes that
-are cosmological in origin, or simply do not rely on the radiation field
-directly. These include cooling via free-free emission and hydrogen ionization
-due to helium recombination photons (which depend on the gas kinetic
-temperature and electron density), and cosmological effects such as Hubble
-cooling, Compton cooling off cosmic microwave background (CMB) photons, and
-photo-ionization by Wien-tail CMB photons (which depend on kinetic
-temperature, redshift, and the Hubble parameter). Our optimized SEDs
-(\S\ref{sec:Results}) will be unchanged upon including such physics in later
-work.
-
+% Secondary Electrons
 \subsection{Additional Complexity: Energy Dependent Secondary Electron Treatment}
@@
 \end{align}    
+Various methods for handling secondary electrons are compared for a $10^5$ K blackbody source of $\dot{Q} = 5 \times 10^{48}$ ionizing photons $s^{-1}$ in Figure \ref{fig:SecondaryIonizationMethods}.
+
+% Secondary electron methods
+\begin{figure*}[htbp!]
+\begin{center}
+\includegraphics[width=0.98\textwidth]{figures/secondary_ionization_methods.eps}
+\caption{Comparison of different methods for secondary ionization.  See \S\ref{sec:PhysicsParameters} for details on setting the secondary ionization method.}
+\label{fig:SecondaryIonizationMethods}
+\end{center}
+\end{figure*}
+
+\subsection{Time-stepping Techniques}
+
+\subsection{Parallelism}
 
@@
 \subsection{\texttt{InitializeGrid}}
+Initialize the temperature and ion densities for each cell in the grid.
 \subsection{\texttt{InitializeIntegralTables}}
+If \texttt{IntegralTabulation} = 1 (necessary when \texttt{DiscreteSpectrum} = 0), tabulates the quantities $\Phi_i$ and $\Psi_i$.
+If \texttt{MultiSpecies} = 1, tabulation occurs for hydrogen and helium ions, and if \texttt{SecondaryIonization} $\ge 2$, also tabulate $\widetilde{\Phi}_i$, $\widetilde{\Psi}_i$, $\widehat{\Phi}_i$, and $\widehat{\Psi}_i$, adding an additional dimension for $x_{\mathrm{HII}}$.
 \subsection{\texttt{RadiationSource}}
+Initialize the radiation source.  Most importantly the \texttt{Spectrum} function, which returns the fraction of the bolometric luminosity emitted at energy $E$ (only argument).
 \subsection{\texttt{Radiate}}
+Calls \texttt{EvolvePhotons} routines, which loop over grid, call solver, and update ion densities and the temperature in each cell.
 \subsection{\texttt{SolveRateEquations}}
+ODE solver.
 \subsection{\texttt{Interpolate}}
+Performs linear, bi-linear, tri-linear, and eventually quad-linear interpolation on integral tables to extract ionization and heating coefficients as a function of column density.
 \subsection{\texttt{WriteData}}
+Write out data to HDF5.  Would be easy to add support for ASCII write-out.
 
@@
 %%%
-\section{Parameters}
+\section{Parameters} \label{sec:Parameters}
 
@@
 
-\item [\texttt{ProblemType = 2.1}] Same as \texttt{ProblemType = 2} for everything, except we sample the continuous blackbody spectrum with the four energy groups used by \citet{Wise2011}
+\item [\texttt{ProblemType = 2.1}] Same as \texttt{ProblemType = 2}, except we sample the continuous blackbody spectrum with the four energy groups used by \citet{Wise2011}.
 
-\item [\texttt{ProblemType = 2.2}] Same as \texttt{ProblemType = 2} for everything, except we sample the continuous blackbody spectrum with the four energy groups used by Mirocha et al. 2012.
+\item [\texttt{ProblemType = 2.2}] Same as \texttt{ProblemType = 2}, except we sample the continuous blackbody spectrum with the four energy groups used by Mirocha et al. (2012).
 
-\item [\texttt{ProblemType = 3.1}] Test problem \# 3 from \cite{Iliev2006} -- I-front trapping in a dense clump and the formation of a shadow.
-The setup is the same as in \texttt{ProblemType = 2.1}, except a 1D clump is initialized at $x = 0.76 L_{\mathrm{box}}$, overdensity $\delta = 200$, radius $0.12$kpc, and temperature $T = 40$K, and the radiation field is plane-parallel.
+\item [\texttt{ProblemType = 3}] Test problem \# 3 from \cite{Iliev2006} -- I-front trapping in a dense clump and the formation of a shadow.
+The setup is the same as in \texttt{ProblemType = 2}, except a 1D clump is initialized at $x = 0.76 L_{\mathrm{box}}$, overdensity $\delta = 200$, radius $0.12$kpc, and temperature $T = 40$K, and the radiation field is plane-parallel.
+
+\item [\texttt{ProblemType = 3.1}] Same as \texttt{ProblemType = 3}, except we sample the continuous blackbody spectrum with the four energy groups used by \citet{Wise2011}.
+
 \end{description}
@@
 
-\subsection{Physics Parameters}
+\subsection{Physics Parameters} \label{sec:PhysicsParameters}
 
@@
 \end{verbatim}
+
+%%%
+%% Test Problems
+%%%
+\section{Test Problems}
+In this section we focus on test problems 1-3 of the Radiative Transfer Comparison Project \citep[hereafter RT06;][]{Iliev2006}.
+
+\subsection{Test \#0}
+Single-zone ionization, heating, and cooling.
+
+\begin{figure*}[htbp!]
+\begin{center}
+\includegraphics[width=0.98\textwidth]{figures/rt06_0.eps}
+\caption{Convergence test for RT06 \#1.}
+\label{fig:RT06_1}
+\end{center}
+\end{figure*}
+
+\subsection{Test \#1}
+
+\begin{figure*}[htbp!]
+\begin{center}
+\includegraphics[width=0.98\textwidth]{figures/rt06_1_convergence.eps}
+\caption{Convergence test for RT06 \#1.}
+\label{fig:RT06_1}
+\end{center}
+\end{figure*}
+
+\subsection{Test \#2}
+
+\begin{figure*}[htbp!]
+\begin{center}
+\includegraphics[width=0.98\textwidth]{figures/rt06_2.eps}
+\caption{RT06 \#3.}
+\label{fig:RT06_2}
+\end{center}
+\end{figure*}
+
+\subsection{Test \#3}
+
+\begin{figure*}[htbp!]
+\begin{center}
+\includegraphics[width=0.98\textwidth]{figures/rt06_3.eps}
+\caption{RT06 \#3.}
+\label{fig:RT06_3}
+\end{center}
+\end{figure*}
 
@@
 
-%How to change the code to allow for this?
-%Define main loop in RT1D.py as Shine.  Insert if __name__ == `main' stuff?
-%Would be nice to have an executable...
-%Move RT1D.py to \$RT1D/bin, import it from \$RT1D/rt1d/__init__.py
-
 What parameters must be set?
@@
 
-%In this work, we use the formulae in Appendix B of \cite{Fukugita1994} to
-%compute the values of $\alpha_i$, $\beta_i$, and $\xi_i$.
-As you might expect, collisional ionization and recombination depend solely on the temperature of the gas.
-For example, collisional ionization of hydrogen (from $n = 1$) is governed by
-%\begin{equation}
-%    \beta_{1,\HI} = 5.85 \times 10^{-11} T^{1/2} \left[1 + \left(\frac{T}{10^5 \mathrm{K}}\right)^{1/2}\right]^{-1} \mathrm{exp}\left[-1.578 \times \left(\frac{T}{10^5 \mathrm{K}}\right)^{-1}\right] \mathrm{cm^3 s^{-1}}
-%\end{equation}
-%and recombinations (to $n \ge 1$) by
-%\begin{equation}
-%    \alpha_{\HII} = 6.28 \times 10^{-11} T^{-1/2} \left(\frac{T}{10^3\mathrm{K}}\right)^{-1/5} \left[1 + \left(\frac{T}{10^6\mathrm{K}}\right)^{0.7}\right]^{-1} \mathrm{cm^3s^{-1}}
-%\end{equation}
-
 \newpage

### 2012-04-15 rt1d/analysis/DataDump.py

@@ -74,16 +74,16 @@ class DataDump:
         self.nabs = np.array([self.n_HI, self.n_HeI, self.n_HeII])
         self.nion = np.array([self.n_HII, self.n_HeII, self.n_HeIII])
         
-        self.Gamma = np.zeros([3, self.GridDimensions])
-        self.k_H = np.zeros([3, self.GridDimensions])
-        self.gamma = np.zeros([3, 3, self.GridDimensions])
-        self.Beta = np.zeros([3, self.GridDimensions])
-        self.alpha = np.zeros([3, self.GridDimensions])
-        self.xi = np.zeros([3, self.GridDimensions])
-        self.zeta = np.zeros([3, self.GridDimensions])
-        self.eta = np.zeros([3, self.GridDimensions])
-        self.psi = np.zeros([3, self.GridDimensions])
-        self.omega = np.zeros([3, self.GridDimensions])
+        self.Gamma = np.zeros([self.GridDimensions, 3])
+        self.k_H = np.zeros_like(self.Gamma)
+        self.gamma = np.zeros_like(self.Gamma)
+        self.Beta = np.zeros_like(self.Gamma)
+        self.alpha = np.zeros_like(self.Gamma)
+        self.xi = np.zeros_like(self.Gamma)
+        self.zeta = np.zeros_like(self.Gamma)
+        self.eta = np.zeros_like(self.Gamma)
+        self.psi = np.zeros_like(self.Gamma)
+        self.omega = np.zeros_like(self.Gamma)
         
         # try-except to ensure backward compatibility
         try:
@@ -95,23 +95,22 @@ class DataDump:
             # odeit is more interesting than rootit alone.
             self.rootit = dd['RootFinderIterations'].value
         
-        
             if pf["OutputRates"].value:
                 for i in xrange(3):
-                    self.Gamma[i] = dd['PhotoIonizationRate%i' % i].value
-                    self.k_H[i] = dd['PhotoHeatingRate%i' % i].value
-                    self.Beta[i] = dd['CollisionalIonizationRate%i' % i].value
-                    self.alpha[i] = dd['RadiativeRecombinationRate%i' % i].value
-                    self.zeta[i] = dd['CollisionalIonzationCoolingRate%i' % i].value
-                    self.eta[i] = dd['RecombinationCoolingRate%i' % i].value
-                    self.psi[i] = dd['CollisionalExcitationCoolingRate%i' % i].value
+                    self.Gamma[:,i] = dd['PhotoIonizationRate%i' % i].value
[…]

### 2012-04-15 rt1d/mods/ComputeCrossSections.py

@@ -5,8 +5,8 @@ Author: Jordan Mirocha
 Affiliation: University of Colorado at Boulder
 Created on 2010-10-25.
 
-Description: Compute photo and collisional ionization cross sections as well as recombination rate 
-coefficients and anything else like this that may crop up.
+Description: Compute the bound-free absorption cross-sections via the fitting
+formulae of Verner et al. 1996, and eventually others.
      
 """
 

### 2012-04-15 rt1d/mods/ControlSimulation.py

@@ -40,12 +40,14 @@ class ControlSimulation:
         self.MaxHeIIChange = pf["MaxHeIIChange"]
         self.MaxHeIIIChange = pf["MaxHeIIIChange"]
         self.MaxElectronChange = pf["MaxElectronChange"]
+        self.MaxTemperatureChange = pf["MaxTemperatureChange"]
         self.HIRestrictedTimestep = pf["HIRestrictedTimestep"]
         self.HeIRestrictedTimestep = pf["HeIRestrictedTimestep"]
         self.HeIIRestrictedTimestep = pf["HeIIRestrictedTimestep"]
         self.HeIIIRestrictedTimestep = pf["HeIIIRestrictedTimestep"]
         self.OpticalDepthDefiningIFront = pf["OpticalDepthDefiningIFront"]
-        self.ElectronFractionRestrictedTimestep = pf["ElectronFractionRestrictedTimestep"]
+        self.ElectronRestrictedTimestep = pf["ElectronRestrictedTimestep"]
+        self.TemperatureRestrictedTimestep = pf["TemperatureRestrictedTimestep"]
         self.MinimumSpeciesFraction = pf["MinimumSpeciesFraction"]
         
         self.mask = np.array([[1, 0, 0], [0, 1, 0], [0, 0, 1]])
@@ -95,14 +97,18 @@ class ControlSimulation:
             dHeIIIdt = nabs[2] * Gamma[2] 
             dtHeIII = self.MaxHeIIIChange * nion[2] / dHeIIdt
                   
-        dtef = 1e50  
-        if self.ElectronFractionRestrictedTimestep:
-            defdt = np.abs(dHIdt + dHeIIdt + 2. * dHeIIIdt)
-            dtef = self.MaxElectronChange * n_e / (n_e + n_H + n_He) / defdt 
-        
-        return min(dtHI, dtHeII, dtHeIII, dtef)
+        dtne = 1e50     
+        if self.ElectronRestrictedTimestep:  
+            dHIIdt = nabs[0] * Gamma[0]
+            dHeIIdt = dHeIIdt = nabs[1] * Gamma[1]
+            dHeIIIdt = dHeIIIdt = nabs[2] * Gamma[2]              
+            dedt = np.abs(dHIIdt + dHeIIdt + 2. * dHeIIIdt)
+            dtne = self.MaxElectronChange * n_e / dedt 
+                  
+        return min(dtHI, dtHeII, dtHeIII)
     
[…]

### 2012-04-15 rt1d/mods/InitializeIntegralTables.py

@@ -95,7 +95,7 @@ class InitializeIntegralTables:
         self.HIColumnMin = np.floor(np.log10(pf["MinimumSpeciesFraction"] * np.min(self.n_H * grid.dx)))
         self.HIColumnMax = np.ceil(np.log10(pf["LengthUnits"] * np.max(self.n_H)))
         self.HeIColumnMin = self.HeIIColumnMin = np.floor(np.log10(pf["MinimumSpeciesFraction"] * np.min(self.n_He * grid.dx)))
-        self.HeIColumnMax = self.HeIIColumnMax = np.ceil(np.log10(pf["LengthUnits"] * np.max(self.n_He)))
+        self.HeIColumnMax = self.HeIIColumnMax = np.ceil(np.log10(pf["LengthUnits"] * np.max(self.n_He)))            
         
         self.HINBins = pf["ColumnDensityBinsHI"]
         self.HeINBins = pf["ColumnDensityBinsHeI"]
@@ -528,8 +528,8 @@ class InitializeIntegralTables:
     def Phi(self, ncol, species = 0):
         """
         Equation 10 in Mirocha et al. 2012.
-        """        
-        
+        """      
+                
         # Otherwise, continuous spectrum                
         if self.PhotonConserving:
             integrand = lambda E: 1e10 * self.rs.Spectrum(E, Lbol = self.Lbol) * \

### 2012-04-15 rt1d/mods/ProblemTypes.py

@@ -22,14 +22,47 @@ way:
 
 import numpy as np
 
-cm_per_kpc = 3.08568 * 10**21
-s_per_myr = 365.25 * 24 * 3600 * 10**6
+cm_per_kpc = 3.08568e21
+s_per_myr = 365.25 * 24 * 3600 * 1e6
 
 def ProblemType(ptype):
     """
-    Storage bin for predefined problem types, 'pt's, like those used in the 
+    Storage bin for predefined problem types, like those used in the 
     radiative transfer comparison project ('RT06').
     """
+    
+    # RT06-0.3, Single zone ionization/heating, then source switches off.
+    if ptype == 0:
+        pf = {"ProblemType": 0, 
+              "TabulateIntegrals": 1, 
+              "LengthUnits": 1e-4 * cm_per_kpc, # 100 milliparsecs 
+              "GridDimensions": 1, 
+              "StartRadius": 0.99, # cell = 1 milliparsec across
+              "StopTime": 10, 
+              "LogarithmicDataDump": 1,
+              "ElectronRestrictedTimestep": 1,
+              "TemperatureRestrictedTimestep": 0,
+              "NlogDataDumps": 500, 
+              "ODEMinStep": 1e-15,
+              "MinimumSpeciesFraction": 1e-9, 
+              "ColumnDensityBinsHI": 20,
+              "DensityProfile": 0,
+              "InitialDensity": 1., 
+              "TemperatureProfile": 0, 
+              "InitialTemperature": 1e2,
+              "IonizationProfile": 0, 
+              "InitialHIIFraction": 1e-6, 
+              "SourceType": 1, 
+              "SourceLifetime": 0.5,
+              "SpectrumPhotonLuminosity": 1e12, 
+              "SpectrumMinEnergy": 13.6, 
+              "SpectrumMaxEnergy": 100.,
+              "SpectrumMinNormEnergy": 0.1, 
+              "SpectrumMaxNormEnergy": 100., 
+              "DiscreteSpectrum": 0,
+              "Isothermal": 0,
+              "PlaneParallelField": 1              
+             }        
     
     # RT06-1, RT1: Pure hydrogen, isothermal HII region expansion, monochromatic spectrum at 13.6 eV
     if ptype in [1, 1.1]:
@@ -119,17 +152,15 @@ def ProblemType(ptype):
              "DiscreteSpectrumRelLum": [0.24, 0.35, 0.23, 0.06],
              "Isothermal": 0
             }    
-    
-    
-    
[…]

### 2012-04-15 rt1d/mods/Radiate.py

@@ -85,13 +85,13 @@ class Radiate:
         self.HeIRestrictedTimestep = pf["HeIRestrictedTimestep"]
         self.HeIIRestrictedTimestep = pf["HeIIRestrictedTimestep"]
         self.HeIIIRestrictedTimestep = pf["HeIIIRestrictedTimestep"]
-        self.ElectronFractionRestrictedTimestep = pf["ElectronFractionRestrictedTimestep"]
+        self.ElectronRestrictedTimestep = pf["ElectronRestrictedTimestep"]
         self.LightCrossingTimeRestrictedTimestep = pf["LightCrossingTimeRestrictedTimestep"]
         self.AdaptiveODEStep = pf["ODEAdaptiveStep"]
         self.MaxStep = pf["ODEMaxStep"] * self.TimeUnits
         self.MinStep = pf["ODEMinStep"] * self.TimeUnits
         
-        self.AdaptiveGlobalStep = self.HIRestrictedTimestep or self.ElectronFractionRestrictedTimestep
+        self.AdaptiveGlobalStep = self.HIRestrictedTimestep or self.ElectronRestrictedTimestep
         if self.MultiSpecies:
             self.AdaptiveGlobalStep |= (self.HeIRestrictedTimestep or \
                 self.HeIIRestrictedTimestep or self.HeIIIRestrictedTimestep)
@@ -290,7 +290,7 @@ class Radiate:
         
         # Could change with time for accreting black holes
         Lbol = self.rs.BolometricLuminosity(t)
-                
+                        
         # Loop over cells radially, solve rate equations, update values in data -> newdata
         for cell in self.grid:   
                         
@@ -379,7 +379,8 @@ class Radiate:
             # Adjust timestep based on maximum allowed neutral fraction change     
             if self.AdaptiveGlobalStep:
                 dtphot[cell] = self.control.ComputePhotonTimestep(tau, 
-                    nabs, nion, ncol, n_H, n_He, n_e, n_B, Gamma, gamma, Beta, alpha, xi, dt) 
+                    nabs, nion, ncol, n_H, n_He, n_e, n_B, Gamma, gamma, Beta, alpha, k_H, 
+                        zeta, eta, psi, xi, omega, newT, dt) 
         
         return newdata, dtphot
         
@@ -560,7 +561,7 @@ class Radiate:
                 dtphot[cell] = self.control.ComputePhotonTimestep(self.tau_all_arr[:,cell], 
[…]

### 2012-04-15 rt1d/mods/RadiationSource.py

@@ -53,9 +53,9 @@ class RadiationSource:
         self.pf = pf
         self.SourceType = pf["SourceType"]
         self.DiscreteSpectrum = pf["DiscreteSpectrum"]
-        self.tau = pf["SourceLifetime"]
         self.TimeUnits = pf["TimeUnits"]
         self.PlaneParallelField = pf["PlaneParallelField"]
+        self.tau = pf["SourceLifetime"] * self.TimeUnits
                 
         # Spectrum bounds
         self.Emin = pf["SpectrumMinEnergy"]
@@ -199,7 +199,7 @@ class RadiationSource:
         bolometric luminosity will increase with time, hence the optional 't' argument.
         """
         
-        if (t / self.TimeUnits) > self.tau: 
+        if t >= self.tau: 
             return 0.0
         
         if self.SourceType == 0: 

### 2012-04-15 rt1d/mods/ReadParameterFile.py

@@ -17,8 +17,8 @@ monochromatic source of photons).
 
 import re, copy, h5py
 import numpy as np
-from rt1d.mods.SetDefaultParameterValues import SetDefaultParameterValues
-from rt1d.mods.ProblemTypes import ProblemType
+from .SetDefaultParameterValues import SetDefaultParameterValues
+from .ProblemTypes import ProblemType
 
 def ReadParameterFile(pf):
     """
@@ -42,7 +42,7 @@ def ReadParameterFile(pf):
         parname, eq, parval = line.partition("=")
                         
         # ProblemType option
-        if parname.strip() == 'ProblemType' and float(parval) > 0:
+        if parname.strip() == 'ProblemType' and float(parval) > -1:
             pf_new = ProblemType(float(parval))
             for param in pf_new: 
                 pf_dict[param] = pf_new[param]

### 2012-04-15 rt1d/mods/SecondaryElectrons.py

@@ -29,7 +29,7 @@ class SecondaryElectrons:
         self.NumberOfEnergyBins = 258
         self.NumberOfXiBins = 14
         
-        if self.Method == 2:
+        if self.Method >= 2:
             rt1d = os.environ.get("RT1D")
             if rt1d:
                 f = h5py.File("{0}/input/secondary_electron_data.h5".format(rt1d), 'r')
@@ -54,12 +54,7 @@ class SecondaryElectrons:
             self.xmax = max(self.IonizedFractions)
             
             f.close()        
-        
-        if self.Method == 3:
-            self.IonizedFractions = x_HII
-            self.LogIonizedFractions = np.log10(self.IonizedFractions)    
-            
-        
+
     def DepositionFraction(self, E, xHII, channel = 0):
         """
         Return the fraction of secondary electron energy deposited as heat, or further ionizations.
@@ -75,8 +70,8 @@ class SecondaryElectrons:
         
             Method = 0: OFF - all secondary electron energy goes to heat.
             Method = 1: Empirical fits of Shull & vanSteenberg 1985.
-            Method = 2: Lookup tables of Furlanetto & Stoever 2010.
-            Method = 3: Empirical Fits of Ricotti et al. 2002.
+            Method = 2: Empirical Fits of Ricotti et al. 2002.
+            Method = 3: Lookup tables of Furlanetto & Stoever 2010.
             
         """
         
@@ -102,7 +97,38 @@ class SecondaryElectrons:
             if channel >= 3: 
                 return 0.0
             
+        # Ricotti, Gnedin, & Shull (2002)
         if self.Method == 2:
+            if channel == 0: 
+                if xHII <= 1e-4: 
+                    return 0.15 # This is what they do in Thomas & Zaroubi (2008).
+                else: 
+                    if E >= 11:
+                        return 3.9811 * (11. / E)**0.7 * pow(xHII, 0.4) * \
+                            (1. - pow(xHII, 0.34))**2 + \
+                            (1. - (1. - pow(xHII, 0.2663))**1.3163)
+                    else:
+                        return 0.0
+                    
+            if channel == 1: 
+                if E >= 28:
+                    return -0.6941 * (28. / E)**0.4 * pow(xHII, 0.2) * \
+                        (1. - pow(xHII, 0.38))**2 + \
+                        0.3908 * (1. - pow(xHII, 0.4092))**1.7592
+                else:
+                    return 0.0    
+            if channel == 2: 
+                if E >= 28:
[…]

### 2012-04-18 rt1d/analysis/Analysis.py
Added wiki/Home.wiki

@@ -308,13 +308,17 @@ class Analyze:
             t, nabs = self.CellTimeEvolution(cell = cell, field = 'nabs')
             t, ne = self.CellTimeEvolution(cell = cell, field = 'n_e')
             val *= zip(*nabs)[species] * ne
-        elif field in ['zeta', 'eta', 'psi']:
+        elif field in ['zeta', 'psi']:
             t, nabs = self.CellTimeEvolution(cell = cell, field = 'nabs')
             t, ne = self.CellTimeEvolution(cell = cell, field = 'n_e')
             val *= zip(*nabs)[species] * ne
+        elif field in ['eta']:
+            t, nion = self.CellTimeEvolution(cell = cell, field = 'nion')
+            t, ne = self.CellTimeEvolution(cell = cell, field = 'n_e')
+            val *= zip(*nion)[species] * ne
             
         self.ax = pl.subplot(111)
-        self.ax.loglog(t / s_per_yr, val, color = color)  
+        self.ax.loglog(t / s_per_yr, val, color = color, ls = ls)  
         
         pl.draw()  
             

### 2012-04-24 doc/manual/rt1d_manual.tex
Added wiki/Home.wiki
Added documentation PDF.

@@
 \author{Jordan Mirocha}	
-\title{\Large {\bf rt1d: User's Manual}}
+\title{\Large {\bf rt1d: Documentation}}
 \date{Last Updated: \today}
@@
 
+\item [\texttt{ProblemType = 0}] Single-zone.
+
 \item [\texttt{ProblemType = 1}] This will set up a standard test with known analytic solution: the propagation of an ionization front in an isothermal, hydrogen-only, static universe.
@@
 \section{Test Problems}
-In this section we focus on test problems 1-3 of the Radiative Transfer Comparison Project \citep[hereafter RT06;][]{Iliev2006}.
+In this section we focus on test problems 0-3 of the Radiative Transfer Comparison Project \citep[hereafter RT06;][]{Iliev2006}.
 
@@
 \includegraphics[width=0.98\textwidth]{figures/rt06_0.eps}
-\caption{Convergence test for RT06 \#1.}
-\label{fig:RT06_1}
+\caption{Single-zone ionization, heating, and cooling RT06 \#0.}
+\label{fig:RT06_0}
 \end{center}
@@
 \section{Start to Finish}
+
+%%%
+%% Advanced usage
+%%%
+\section{Advanced Usage}
 

### 2012-05-01 doc/manual/rt1d_manual.tex
Added README file to root directory.
In the process of implementing cosmological expansion...not going so great.

@@
 %%% START custom commands
+\newcommand{\H}{\text{H}}
+\newcommand{\He}{\text{He}}
 \newcommand{\HI}{\text{HI}}
@@
 
+% Cosmology
+\subsection{Additional Complexity: Cosmological Expansion}
+For an expanding universe, the number density of species $i$ (neglecting all processes except for the cosmological expansion) is
+\begin{equation}
+    n_i(z) = \bar{n}_i^0 (1 + z)^3 ,
+\end{equation}    
+where $\bar{n}_i^0$ is the number density today.  The rate of change in the number density of species $i$ is then
+\begin{align}
+    \frac{dn_i}{dt} & = 3 \bar{n}_i^0 (1 + z)^2 \frac{dz}{dt} \nonumber \\
+    & = 3 \bar{n}_i^0 (1 + z)^3 H(z)
+\end{align}
+It will be advantageous to work with ionized fractions now, i.e.
+\begin{align}
+    \dot{x}_{\HII} & = (\gHI + \gamma_{\HI} + \bHI \nel) (1 - \xHII) - \aHII \nel \xHII   \label{eq:HIIRateEquation} \\ 
+    \dot{x}_{\HeII} & = (\gHeI + \gamma_{\HeI} + \bHeI \nel) (1 - \xHeI - \xHeII) \nonumber + \aHeIII \nel \xHeIII \\  & - (\bHeII + \aHeII + \xiHeII) \nel \xHeII \label{eq:HeIIRateEquation} \\ 
+    \dot{x}_{\HeIII} & = (\gHeII + \gamma_{\HeII} + \bHeII \nel) \xHeII  - \aHeIII \nel \xHeIII . \label{eq:HeIIIRateEquation}
+\end{align}
+where
+\begin{equation}
+    n_e = \bar{n}_{\H}^0 (1 + \xHII) (1 + z)^3 + \bar{n}_{\He}^0 (1 + \xHeII + 2 \xHeIII) (1 + z)^3
+\end{equation}
+
 \subsection{Time-stepping Techniques}
@@
 %%%
-%% Start to Finish
-%%%
-\section{Start to Finish}
-
-%%%
 %% Advanced usage

### 2012-05-01 rt1d/analysis/Analysis.py

@@ -16,6 +16,7 @@ from .Multiplot import *
 from .Dataset import Dataset
 from .Inspection import Inspect
 from ..mods.Constants import *
+from ..mods.Cosmology import *
 from ..mods.Interpolate import Interpolate
 from ..mods.InitializeGrid import InitializeGrid
 from ..mods.SecondaryElectrons import SecondaryElectrons
@@ -28,6 +29,7 @@ class Analyze:
         self.data = self.ds.data
         self.pf = self.ds.pf    # dictionary
         self.g = InitializeGrid(self.pf)   
+        self.cosm = Cosmology(self.pf)
         self.iits = InitializeIntegralTables(self.pf, self.data[0], self.g)      
         self.esec = SecondaryElectrons(self.pf)         
         
@@ -207,7 +209,8 @@ class Analyze:
         
         pl.draw()
         
-    def RadialProfileMovie(self, field = 'x_HI', out = 'frames', scale = 'linear'):
+    def RadialProfileMovie(self, field = 'x_HI', out = 'frames', xscale = 'linear',
+        title = True):
         """
         Save time-series images of 'field' to 'out' directory.
         
@@ -235,12 +238,54 @@ class Analyze:
 
             ax.set_xlim(self.data[0].r[0] / self.pf['LengthUnits'], 1)        
             ax.set_ylim(mi, ma)    
-            ax.set_xscale(scale)     
+            ax.set_xscale(xscale) 
+            
+            if title:
+                ax.set_title(r'$t = %g \ \mathrm{Myr}$' % self.data[dd].t / self.pf['TimeUnits'])
                     
             pl.savefig('%s/dd%s_%s.png' % (out, str(dd).zfill(4), field))                        
             ax.clear()
             
         pl.close()    
+        
+    def Ionization_Temperature_Movie(self, out = 'frames', xscale = 'linear', title = True):
+        """
+        Meant to answer Eric's question.
+        """    
+                
+        for dd in self.data.keys():
+            
+            mp = multiplot(dims = (2, 1), useAxesGrid = False)
+            
+            exec('mp.grid[0].semilogy(self.data[%i].r / self.pf[\'LengthUnits\'], \
+                self.data[%i].x_HI, ls = \'-\', color = \'k\', label = r\'$x_{\mathrm{HI}}$\')' % (dd, dd))
+            exec('mp.grid[0].semilogy(self.data[%i].r / self.pf[\'LengthUnits\'], \
[…]

### 2012-05-01 rt1d/mods/__init__.py

@@ -2,6 +2,7 @@
 Import everything.
 """
 
+from .Cosmology import Cosmology
 from .InitializeIntegralTables import InitializeIntegralTables
 from .InitializeGrid import InitializeGrid
 from .Radiate import Radiate

### 2012-05-01 doc/manual/rt1d_manual.tex
Hubble cooling working - must be careful with decoupling redshift - need to work on Compton scattering in order for this to be solved self-consistently.

--- 2012-04-24 doc/manual/rt1d_manual.tex
@@
 
+% Cosmology
+\subsection{Additional Complexity: Cosmological Expansion}
+For an expanding universe, things are slightly more complicated, as ion densities will have an additional term $\propto (1 + z)^3$, and we must take into account Hubble cooling, and Compton scattering of CMB photons.
+
+We must also modify the minimum and maximum column densities for our integral tables.
+To calculate what these should be for a simulation spanning the redshift interval $(z_i, z_f)$, minimum species fraction $x_{\mathrm{min}}$, and grid resolution (in proper coordinates) $dx$, do
+\begin{align}
+    N_{\HI}^{\mathrm{min}} & = \bar{n}_H^0 (1 + z_f)^3 dx x_{\mathrm{min}}   \\
+    N_{\HI}^{\mathrm{max}} & = \bar{n}_H^0 (1 + z_i)^3 \Lbox .
+\end{align}
+
 \subsection{Time-stepping Techniques}

### 2012-05-01 rt1d/mods/ControlSimulation.py

@@ -119,7 +119,7 @@ class ControlSimulation:
             
         dtT = 1e50
         if self.TemperatureRestrictedTimestep:
-            dTdt = np.abs(2. * T * self.cosm.HubbleParameter(self.z0))
+            dTdt = np.abs(2. * T * hubble)
             dtT = self.MaxTemperatureChange * T / dTdt    
 
         return min(dtHI, dtHeII, dtHeIII, dtne, dtT)
@@ -183,7 +183,7 @@ class ControlSimulation:
         if self.TemperatureRestrictedTimestep:
             dTdt = np.abs(np.sum(k_H * nabs) - n_e * (np.sum(zeta * nabs) + \
                 np.sum(eta * nabs) + np.sum(psi * nabs) + nion[2] * omega[1]) - \
-                3. * theta * T / 2.)
+                3. * hubble * T / 2.)
             dtT = self.MaxTemperatureChange * T / dTdt
                         
         return min(dtHI, dtHeII, dtHeIII, dtHeI, dtne, dtT)        

### 2012-05-01 rt1d/mods/Radiate.py

@@ -701,10 +701,10 @@ class Radiate:
                 + np.sum(eta * nion) + np.sum(psi * nabs) + q[2] * omega[1])      
                                
             if self.CosmologicalExpansion:
-                dqdt[3] = -4. * hubble * q[3] / 3.
+                dqdt[3] -=2. * hubble * q[3]
                 
                 if self.ComptonHeating:
                     dqdt[3] += compton
-                
+                                                        
         return dqdt
         

### 2012-05-07 doc/manual/rt1d_manual.tex
Started implementing MCD+PL SED.  Need to think more about time-dependent SEDs.  Pausing to dotdictify pfs in major classes.

@@
 \begin{align}
-    \frac{3}{2}\frac{d}{dt}\left(\frac{\kB T_k \ntot}{\mu}\right) & = \fheat  \sum_i n_i \Heat_i - \sum_i \zeta_i n_e n_i - \sum_i \eta_i n_e n_i \nonumber \\ & - \sum_i \psi_i n_e n_i - \omega_{\HeII} n_e \nHeII \label{eq:TemperatureEvolution}
+    \frac{3}{2}\frac{d}{dt}\left(\frac{\kB T_k \ntot}{\mu}\right) & = \fheat  \sum_i n_i \Heat_i - \sum_i \zeta_i n_e n_i - \sum_{i^{\prime}} \eta_i n_e n_i \nonumber \\ & - \sum_i \psi_i n_e n_i - \omega_{\HeII} n_e \nHeII \label{eq:TemperatureEvolution}
 \end{align}
-where $\Heat_i$ is the photo-electric heating rate coefficient (due to
-electrons previously bound to species $i$), $\omega_{\HeII}$ is the dielectric
-recombination cooling coefficient, and $\zeta_i$, $\eta_i$, and $\psi_i$ are
-the collisional ionization, recombination, and collisional excitation cooling
-coefficients, respectively. The constants in Equation
-\ref{eq:TemperatureEvolution} are the total number density of baryons, $\ntot
-= n_\mathrm{H} + n_{\mathrm{He}} + n_e$, the mean molecular weight, $\mu$,
-Boltzmann's constant, $\kB$, and the fraction of secondary electron energy
-deposited as heat, $\fheat$. We use the formulae in Appendix B of
-\citet{Fukugita1994} to compute the values of $\alpha_i$, $\beta_i$, $\xi_i$,
-$\zeta_i$, $\eta_i$, $\psi_i$, and $\omega_{\HeII}$.
+where the index $i$ corresponds as usual to species that have bound electrons,
+and $i^{\prime}$ corresponds to ionized species. $\Heat_i$ is the
+photo-electric heating rate coefficient (due to electrons previously bound to
+species $i$), $\omega_{\HeII}$ is the dielectric recombination cooling
+coefficient, and $\zeta_i$, $\eta_i$, and $\psi_i$ are the collisional
+ionization, recombination, and collisional excitation cooling coefficients,
+respectively. The constants in Equation \ref{eq:TemperatureEvolution} are the
+total number density of baryons, $\ntot = n_\mathrm{H} + n_{\mathrm{He}} +
+n_e$, the mean molecular weight, $\mu$, Boltzmann's constant, $\kB$, and the
+fraction of secondary electron energy deposited as heat, $\fheat$. We use the
+formulae in Appendix B of \citet{Fukugita1994} to compute the values of
+$\alpha_i$, $\beta_i$, $\xi_i$, $\zeta_i$, $\eta_i$, $\psi_i$, and
+$\omega_{\HeII}$.
 
@@
 \subsection{Additional Complexity: Energy Dependent Secondary Electron Treatment}
-In the previous section (and in all of Mirocha et al.
-2012), we considered the asymptotic limit of \citet{Shull1985}, in which the fractional energy deposition of secondary electrons as heat, ionization, and excitation depends only on the hydrogen ionized fraction.
+In the previous section (and in all of \citet{Mirocha2012}), we considered the asymptotic limit of \citet{Shull1985}, in which the fractional energy deposition of secondary electrons as heat, ionization, and excitation depends only on the hydrogen ionized fraction.
 \citet{Furlanetto2010} updated this work, providing lookup tables of deposition fraction as a function of both ionized fraction and electron energy.
@@
 
-\item [\texttt{ProblemType = 2.2}] Same as \texttt{ProblemType = 2}, except we sample the continuous blackbody spectrum with the four energy groups used by Mirocha et al. (2012).
+\item [\texttt{ProblemType = 2.2}] Same as \texttt{ProblemType = 2}, except we sample the continuous blackbody spectrum with the four energy groups used by \citet{Mirocha2012}.
 
@@
 \subsection{Source Parameters}
+
+\begin{description}
+
+\item \texttt{SourceType}
+\begin{description}
+    \item [0] Monochromatic source.
+    \item [1] Blackbody.  
+    \item [2] PopIII star.  
+    \item [3] Power-law.        
+    \item [4] Absorbed power-law (column density given by \texttt{SpectrumAbsorbingColumn} parameter).
+    \item [5] Multi-color disk + power-law. Relative normalization given by \texttt{SpectrumDiskFraction} parameter.                
+\end{description}
+
+\end{description}
 

### 2012-05-07 rt1d/mods/ComputeRateCoefficients.py

@@ -481,12 +481,12 @@ class RateCoefficients:
             Tcmb = self.cosm.TCMB(z)
             ucmb = sigma_SB * Tcmb**4. / 4. / np.pi / c
             
-            return xHII * k_B * nH * 8. * sigma_T * ucmb * \
-                (Tcmb - Tk) / 2. / (1. + self.cosm.y + xHII) / m_e / c   
+            return 4. * sigma_T * ne * ucmb * k_B * (Tcmb - Tk) / m_e / c
+            #return xHII * k_B * (nH + nHe + ne) * 4. * sigma_T * ucmb * \
+            #    (Tcmb - Tk) / (1. + self.cosm.y + xHII) / m_e / c   
         else:
             return 0.0
             # Implement source-dependent compton heating        
-                 
     
     def ShellVolume(self, r, dr):
         """

### 2012-05-07 rt1d/mods/RadiationSource.py

@@ -36,6 +36,7 @@ cm_per_rsun = 695500.0 * 1e5                                # Radius of the sun 
 s_per_yr = 365.25 * 24 * 3600                               # Seconds per year
 t_edd = 0.45 * 1e9 * s_per_yr                               # Eddington timescale (see eq. 1 in Volonteri & Rees 2005) 
 y_He = 0.08
+xi = (3. / 7.)**0.5 * (6. / 7.)**3                
 
 np.seterr(all = 'ignore')   # exp overflow occurs when integrating BB - will return 0 as it should for x large
 
@@ -82,6 +83,21 @@ class RadiationSource:
         # SourceType = 4
         self.SpectrumAbsorbingColumn = pf["SpectrumAbsorbingColumn"]
         
+        # SourceType = 5
+        self.SourceISCO = pf["SourceISCO"]
+        self.SpectrumDiskFraction = pf["SpectrumDiskFraction"]
+        self.SpectrumColorCorrection = self.fcol = pf["SpectrumColorCorrection"]
+        self.SpectrumTimeEvolution = pf["SpectrumTimeEvolution"]
+        self.Rg = G * self.M * g_per_msun / c**2
+        
+        if not self.SourceISCO:
+            self.r_in = 6. * self.Rg
+        else:
+            self.r_in = self.SourceISCO    
+            
+        self.Lbol = self.BolometricLuminosity(0.0)
+        self.T_in = (self.Lbol * xi**2 * self.fcol**4 / 4. * np.pi / sigma_SB / self.r_in**2)**0.25    
+             
         # Number of ionizing photons per cm^2 of surface area for BB of temperature self.T.  
         # Use to solve for stellar radius (which we need to get Lbol).  The factor of pi gets rid of the / sr units
         if self.SourceType == 0:
@@ -94,16 +110,13 @@ class RadiationSource:
             self.R = np.sqrt(self.Lph / 4. / np.pi / self.LphNorm)        
             self.Lbol = 4. * np.pi * self.R**2 * sigma_SB * self.T**4
             self.Qdot = self.Lbol * self.F / self.E / erg_per_ev 
+             
+        # Join MCD and PL spectra
+        if self.SourceType == 5:
+            self.MCDNormalization, self.PLNormalization = self.NormalizeDisk()
             
         # Normalize spectrum
-        self.LuminosityNormalization = self.NormalizeLuminosity()   
-        
-        # This is just for sanity checks
-        if self.DiscreteSpectrum and self.SourceType in [3]:
[…]

### 2012-05-11 doc/manual/rt1d_manual.tex
cleaning up a bit.

@@
     \item [0] Monochromatic source.
-    \item [1] Blackbody.  
+    \item [1] Star.  
     \item [2] PopIII star.  
-    \item [3] Power-law.        
+    \item [3] Black hole.        
     \item [4] Absorbed power-law (column density given by \texttt{SpectrumAbsorbingColumn} parameter).
@@
 \end{description}    
+\end{description}
 
-\end{description}
+\subsection{Spectrum Parameters}
+We keep these parameters separate from Source parameters to distinguish which parameters are allowed to vary with time, mass, luminosity, etc.
+This separation was useful after renovating the \texttt{RadiationSource} class in May 2012, which made it so any parameter beginning with the word `Spectrum' can be a list, to allow construction of spectra consisting of an arbitrary number of components, each with uniquely specified parameters.
 
@@
 %%% 
-%% Convergence
-%%%
-\section{Convergence}
-In \texttt{rt1d/doc/examples} you will find directories containing convergence test suites for RT06 test problems 1 \& 2 (\texttt{RT06\_1\_ResolutionTestSuite} and \texttt{RT06\_2\_ResolutionTestSuite}).
-Each of these directories contains the parameter files necessary to run these problems on grids containing $100 \times 2^n$ cells, where $n = 0,1,...6$.
-If you want, you can run them all at once, though keep in mind the simulations with 100 cells will finish much much faster than those containing 6400 cells, so it might not be good to check out a node for this as most of the computation will be spent on 1 or 2 simulations.
-For instance, you could do:
-\begin{verbatim}
-    cd rt1d/doc/examples/RT06_1_ResolutionTestSuite/c_infinite
-    mpirun -np 4 python RT1D.py -b batch_list.dat
-\end{verbatim}
-I've included the same setup for $c \neq \infty$ in \texttt{rt1d/doc/examples/RT06\_1\_ResolutionTestSuite/c\_finite}, so you can see the differences (if any) caused by making the $c \rightarrow \infty$ approximation.
-
-%%% 
 %% Troubleshooting

### 2012-05-11 rt1d/analysis/Analysis.py

@@ -36,19 +36,16 @@ class Analyze:
         self.esec = SecondaryElectrons(self.pf)         
         
         # Convenience
-        self.GridDimensions = int(self.pf['GridDimensions'])
+        self.GridDimensions = int(self.pf.GridDimensions)
         self.grid = np.arange(self.GridDimensions)
-        self.LengthUnits = self.pf['LengthUnits']
-        self.StartRadius = self.pf['StartRadius']
-        self.MultiSpecies = self.pf["MultiSpecies"]
                 
         # Deal with log-grid
-        if self.pf["LogarithmicGrid"]:
-            self.r = np.logspace(np.log10(self.StartRadius * self.LengthUnits), \
-                np.log10(self.LengthUnits), self.GridDimensions + 1)
+        if self.pf.LogarithmicGrid:
+            self.r = np.logspace(np.log10(pf.StartRadius * pf.LengthUnits),
+                np.log10(pf.LengthUnits), self.GridDimensions + 1)
         else:
-            rmin = self.StartRadius * self.LengthUnits
-            self.r = np.linspace(rmin, self.LengthUnits, self.GridDimensions + 1)
+            self.r = np.linspace(pf.StartRadius * pf.LengthUnits, 
+                pf.LengthUnits, self.GridDimensions + 1)
         
         self.dx = np.diff(self.r)   
         self.r = self.r[0:-1]
@@ -84,12 +81,12 @@ class Analyze:
         if sol == 0:
             if T0 is not None: T = T0
             else: T = self.data[0].T[0]
-            self.Ndot = self.pf["SpectrumPhotonLuminosity"]
+            self.Ndot = self.pf.SpectrumPhotonLuminosity
             self.alpha_HII = 2.6e-13 * (T / 1.e4)**-0.85
             self.trec = 1. / self.alpha_HII / self.data[0].n_HI[-1]                                         # s
             self.rs = (3. * self.Ndot / 4. / np.pi / self.alpha_HII / self.data[0].n_HI[-1]**2)**(1. / 3.)  # cm
         
-        return self.rs * (1. - np.exp(-t / self.trec))**(1. / 3.) + self.StartRadius
+        return self.rs * (1. - np.exp(-t / self.trec))**(1. / 3.) + self.pf.StartRadius
         
     def LocateIonizationFront(self, dd, species = 0):
         """
@@ -237,7 +234,7 @@ class Analyze:
[…]

### 2012-05-11 rt1d/analysis/Inspection.py

@@ -23,9 +23,7 @@ class Inspect:
         self.data = self.anl.data
         self.itabs = self.anl.itabs
         self.interp = self.anl.interp
-        
-        self.MultiSpecies = self.pf["MultiSpecies"]
-        
+                
     def InspectIntegralTable(self, intnum = 1, species = 0, donor_species = 0,
         nHI = None, nHeI = 0.0, nHeII = 0.0,
         color = 'k', ls = '-', annotate = True):
@@ -67,7 +65,7 @@ class Inspect:
         # Figure out axes
         if nHI is None:
             x = self.itabs['HIColumnValues_x']
-            if self.MultiSpecies:
+            if self.pf.MultiSpecies:
                 i1 = np.argmin(np.abs(self.itabs['HeIColumnValues_y'] - nHeI))
                 i2 = np.argmin(np.abs(self.itabs['HeIIColumnValues_z'] - nHeII))
                 y = self.itabs[integral][0:,i1,i2]
@@ -179,7 +177,7 @@ class Inspect:
         
         self.dt = np.zeros([5, self.GridDimensions])
         for dd in self.data.keys():
-            if self.data[dd].t / self.pf['TimeUnits'] != t: 
+            if self.data[dd].t / self.pf.TimeUnits != t: 
                 continue
         
             dHIIdt = self.data[dd].n_HI * \
@@ -206,11 +204,11 @@ class Inspect:
                 
             defdt = dHIIdt + dHeIIdt + 2. * dHeIIIdt
                                         
-            self.dt[0] = self.pf['MaxHIIChange'] * self.data[dd].n_HI / np.abs(dHIIdt) / self.pf['TimeUnits']
-            self.dt[1] = self.pf['MaxHeIChange'] * self.data[dd].n_HeI / np.abs(dHeIdt) / self.pf['TimeUnits']
-            self.dt[2] = self.pf['MaxHeIIChange'] * self.data[dd].n_HeII / np.abs(dHeIIdt) / self.pf['TimeUnits']
-            self.dt[3] = self.pf['MaxHeIIIChange'] * self.data[dd].n_HeIII / np.abs(dHeIIIdt) / self.pf['TimeUnits']
-            self.dt[4] = self.pf["MaxElectronChange"] * (self.data[dd].n_e / self.data[dd].n_B) / np.abs(defdt) / self.pf['TimeUnits']  
[…]

### 2012-05-11 rt1d/mods/Cosmology.py

@@ -16,7 +16,7 @@ Notes:
 """
 
 import numpy as np
-from .Constants import c, G, km_per_mpc, m_H
+from .Constants import c, G, km_per_mpc, m_H, m_He
 from .Integrate import simpson
 
 class Cosmology:

### 2012-05-11 rt1d/mods/InitializeIntegralTables.py

@@ -142,8 +142,8 @@ class InitializeIntegralTables:
             prop = "M%g" % int(self.pf.SourceMass)
             
         elif self.pf.SourceType >= 3:
-            src = ''
-            prop = ''
+            src = 'bh_M%i' % int(self.pf.SourceMass)
+            prop = '' 
             if 3 in self.rs.SpectrumPars.Type:
                 src += 'mcd'
                 prop += "f%g" % self.rs.SpectrumPars.Fraction[self.rs.SpectrumPars.Type.index(3)]
@@ -157,16 +157,7 @@ class InitializeIntegralTables:
                 prop += "_in%g" % self.pf.SpectrumPowerLawIndex[self.rs.SpectrumPars.Type.index(4)]
                 if self.rs.SpectrumPars.AbsorbingColumn[self.rs.SpectrumPars.Type.index(3)] > 0:
                     prop += "_logN%g" % np.log10(self.rs.SpectrumPars.AbsorbingColumn[self.rs.SpectrumPars.Type.index(3)])
-            if 5 in self.rs.SpectrumPars.Type:
-                src += 'apl' 
-                if 3 in self.rs.SpectrumPars.Type:
-                    prop += '_'
-                prop += "f%g" % self.rs.SpectrumPars.Fraction[self.rs.SpectrumPars.Type.index(4)]
-                prop += "_in%g" % self.pf.SpectrumPowerLawIndex[self.rs.SpectrumPars.Type.index(4)]
-                if self.rs.SpectrumPars.AbsorbingColumn[self.rs.SpectrumPars.Type.index(3)] > 0:
-                    prop += "_logN%g" % np.log10(self.rs.SpectrumPars.AbsorbingColumn[self.rs.SpectrumPars.Type.index(3)])
-        
-        
+
         # Limits
         Hlim = '%i%i' % (self.HIColumn[0], self.HIColumn[-1])
         Helim = '%i%i' % (self.HeIColumn[0], self.HeIColumn[-1])        
@@ -205,10 +196,10 @@ class InitializeIntegralTables:
             table_from_pf = True
         else:
             if rank == 0:
-                print "\nDid not find a pre-existing integral table.  Generating {0}/{1} now...".format(self.pf.OutputDirectory, filename)
-                print "10^%g < ncol_HI < 10^%g" % (self.HIColumnMin, self.HIColumnMax)
+                print "Did not find a pre-existing integral table.  Generating {0}/{1} now...".format(self.pf.OutputDirectory, filename)
[…]

### 2012-05-11 rt1d/mods/RadiationSource.py

@@ -17,7 +17,6 @@ SourceType = 0: Monochromatic/Polychromatic emission of SpectrumPhotonLuminosity
      
 """
 
-import re
 import numpy as np
 from scipy.integrate import quad, fixed_quad
 from Integrate import simpson as integrate                  # why did scipy.integrate.quad mess up LphotNorm? 
@@ -38,7 +37,7 @@ big_number = 1e5
 ls = ['-', '--', ':', '-.']
 
 """
-SourceType = 0  (monochromatic)     Just need DiscreteSpectrum__ and PhotonLuminosity
+SourceType = 0  (monochromatic)     Just need DiscreteSpectrum and PhotonLuminosity
 SourceType = 1  (star)              Need temperature, PhotonLuminosity
 SourceType = 2  (popIII star)       Need mass
 SourceType = 3  (BH)                Need Mass, epsilon
@@ -69,6 +68,7 @@ class RadiationSource:
         self.E = np.array(pf.DiscreteSpectrumSED)
         self.F = np.array(pf.DiscreteSpectrumRelLum)      
         
+        self.last_renormalized = 0
         self.tau = pf.SourceLifetime * pf.TimeUnits
         
         # SourceType 0, 1, 2
@@ -81,11 +81,13 @@ class RadiationSource:
         self.M = pf.SourceMass 
         self.M0 = pf.SourceMass
         self.epsilon = pf.SourceRadiativeEfficiency
-
+        
         if 3 in self.SpectrumPars.Type:
             self.r_in = self.DiskInnermostRadius(self.M0)
+            self.r_out = pf.SourceDiskMaxRadius * self.GravitationalRadius(self.M0)
             self.fcol = self.SpectrumPars.ColorCorrectionFactor[self.SpectrumPars.Type.index(3)]
             self.T_in = self.DiskInnermostTemperature(self.M0)
+            self.T_out = self.DiskTemperature(self.M0, self.r_out)
 
         ### PUT THIS STUFF ELSEWHERE
                                  
@@ -111,10 +113,13 @@ class RadiationSource:
         """
         Half the Schwartzchild radius.
         """
-        return G * M * g_per_msun / c**2
+        return G * M * g_per_msun / c**2    
         
     def SchwartzchildRadius(self, M):
         return 2. * self.GravitationalRadius(M)    
+        
+    def MassAccretionRate(self, M = None):
+        return self.BolometricLuminosity(0, M = M) / self.epsilon / c**2    
         
     def DiskInnermostRadius(self, M):      
[…]

### 2012-05-11 rt1d/mods/SetDefaultParameterValues.py

@@ -129,6 +129,7 @@ def SetDefaultParameterValues():
         "SourceLifetime": 1e10,  
         "SourceRadiativeEfficiency": 0.1,
         "SourceISCO": 0,  
+        "SourceDiskMaxRadius": 1e3,
         "SourceTimeEvolution": 0,
           
         # Spectral parameters
@@ -142,7 +143,6 @@ def SetDefaultParameterValues():
         "SpectrumMaxNormEnergy": 5e2,  
         "SpectrumPhotonLuminosity": 5e48,  
         "SpectrumAbsorbingColumn": 0,  
-        "SpectrumDiskFraction": 1.0,
         "SpectrumColorCorrectionFactor": 1.7,
           
         # Set discrete spectrum manually  

### 2012-05-22 doc/manual/rt1d_manual.tex
Added script to compare optimal monochromatic and effective column density solutions to RT06 #2.

@@
 \usepackage{natbib}
-\usepackage[pdfborder={0 0 0}]{hyperref}
-\usepackage[all]{hypcap}
+\usepackage[pdfborder={0 0 1}]{hyperref}
 
@@
 Let's see what happens to the gas around stars and black holes, shall we? The
-primary reference for this code is
-\href{http://arxiv.org/abs/1204.1944}{arxiv:1204.1944} -- much of the text has
+primary reference for this code is \cite[\href{http://arxiv.org/abs/1204.1944}{arxiv:1204.1944}]{Mirocha2012} -- much of the text has
 been extracted from there.
@@
 In the previous section (and in all of \citet{Mirocha2012}), we considered the asymptotic limit of \citet{Shull1985}, in which the fractional energy deposition of secondary electrons as heat, ionization, and excitation depends only on the hydrogen ionized fraction.
-\citet{Furlanetto2010} updated this work, providing lookup tables of deposition fraction as a function of both ionized fraction and electron energy.
+\citet{Ricotti2002} and \citet{Furlanetto2010} updated this work, providing (respectively) empirical fits and lookup tables of deposition fraction as a function of both ionized fraction and electron energy.
 These effects are most important at $E \lesssim 10^2$ eV.
@@
 
+Let's first account for dilution of ion densities by expansion.  The number density of atoms at the cosmic mean density evolves as
+\begin{align}
+    \frac{dn}{dt} & = \frac{dn}{dz}\frac{dz}{dt} \nonumber \\
+    & = \frac{d}{dz}\left(\bar{n}^0 (1 + z)^3 \right) \frac{dz}{dt} \nonumber \\
+    & = 3 \bar{n}^0 (1 + z)^2 \frac{dz}{dt} .
+\end{align}    
+But, $dz/dt = H(z) (1 +z)$ and $\bar{n}^0 (1 + z)^3 = n$, so
+\begin{equation}
+    \frac{dn}{dt} = -3 n H(z) , \label{eq:dilution}
+\end{equation}    
+where we've inserted a minus sign to indicate that densities decrease with increasing time, or decreasing redshift.
+The right-hand side of Equation \ref{eq:dilution} is added to Equations \ref{eq:HIIRateEquation}-\ref{eq:HeIIIRateEquation} when \texttt{CosmologicalExpansion = 1}.
+
 We must also modify the minimum and maximum column densities for our integral tables.
@@
 \begin{align}
-    N_{\HI}^{\mathrm{min}} & = \bar{n}_H^0 (1 + z_f)^3 dx x_{\mathrm{min}}   \\
-    N_{\HI}^{\mathrm{max}} & = \bar{n}_H^0 (1 + z_i)^3 \Lbox .
+    N_{\HI}^{\mathrm{min}} & = \bar{n}_H^0 (1 + z_f)^3 x_{\mathrm{min}} dx   \\
+    N_{\HI}^{\mathrm{max}} & = \bar{n}_H^0 (1 + z_i)^3 \Lbox
 \end{align}
@@
 %%%
+%% SED studies
+%%%
+\section{Verifying Discrete SEDs with \texttt{rt1d}}
+In \citet{Mirocha2012}, we used \texttt{rt1d} to verify new, optimally selected discrete spectral energy distributions for sources of ultraviolet and X-ray photons.
+As we optimize more and more SEDs for fully 3D cosmological  radiation-hydro simulations, \texttt{rt1d} remains a quick and easy tool for verifying that the SEDs we're using are acceptable.
+
+%%%
 %% Advanced usage
@@
 \section{Advanced Usage}
+In \cite{Mirocha2012}
 

### 2012-05-22 rt1d/analysis/Multiplot.py

@@ -59,7 +59,7 @@ class multiplot:
             
             self.panel_size = np.array([xsize, ysize])
                     
-        if share_all and not useAxesGrid:
+        if (share_all and not useAxesGrid) or not share_all:
             pl.rcParams['figure.subplot.wspace'] = self.padding
             pl.rcParams['figure.subplot.hspace'] = self.padding        
                 

### 2012-05-22 rt1d/mods/RadiationSource.py

@@ -65,6 +65,7 @@ class RadiationSource:
         if self.N == 1:
             self.Type = self.SpectrumPars.Type[0]
                     
+        # Correct later if using multi-group approach
         self.E = np.array(pf.DiscreteSpectrumSED)
         self.F = np.array(pf.DiscreteSpectrumRelLum)      
         
@@ -104,11 +105,45 @@ class RadiationSource:
             self.Lbol = 4. * np.pi * self.R**2 * sigma_SB * self.T**4
             self.Qdot = self.Lbol * self.F / self.E / erg_per_ev
         else:
-            self.Lbol = self.BolometricLuminosity(0.0)           
+            self.Lbol = self.BolometricLuminosity(0.0)    
+                
+        # Parameters for average AGN spectrum of SOS04.
+        self.Alpha = 0.24
+        self.Beta = 1.60
+        self.Gamma = 1.06
+        self.E_1 = 83.
+        self.K = 0.0041
+        self.E_0 = (self.Beta - self.Alpha) * self.E_1
+        self.A = np.exp(2.0 / self.E_1) * 2.0**self.Alpha
+        self.B = ((self.E_0**(self.Beta - self.Alpha)) * np.exp(-(self.Beta - self.Alpha))) / \
+            (1.0 + (self.K * self.E_0**(self.Beta - self.Gamma)))
+            
+        # Normalization constants to make the SOS04 spectrum continuous.
+        self.SX_Normalization = 1.0
+        self.UV_Normalization = self.SX_Normalization * ((self.A * 2000.0**-self.Alpha) * \
+            np.exp(-2000.0 / self.E_1)) / ((1.2 * 2000**-1.7) * np.exp(2000.0 / 2000.0))
+        self.IR_Normalization = self.UV_Normalization * ((1.2 * 10**-1.7) * np.exp(10.0 / 2000.0)) / \
+            (1.2 * 159 * 10**-0.6)
+        self.HX_Normalization = self.SX_Normalization * (self.A * self.E_0**-self.Alpha * \
+            np.exp(-self.E_0 / self.E_1)) / (self.A * self.B * (1.0 + self.K * self.E_0**(self.Beta - self.Gamma)) * \
+            self.E_0**-self.Beta)               
              
         # Normalize spectrum
         self.LuminosityNormalizations = self.NormalizeSpectrumComponents(0.0)
-          
+        
+        # Multi-group treatment
+        if pf.FrequencyAveragedCrossSections:
+            self.E = np.zeros(pf.FrequencyGroups)
[…]

### 2012-05-22 rt1d/mods/SetDefaultParameterValues.py

@@ -134,6 +134,9 @@ def SetDefaultParameterValues():
           
         # Spectral parameters
         "DiscreteSpectrum": 0,
+        "FrequencyAveragedCrossSections": 0,
+        "FrequencyGroups": 1,
+        "FrequencyBands": [13.6, 24.6, 54.4],
         "SpectrumType": 1, 
         "SpectrumFraction": 1,   
         "SpectrumPowerLawIndex": 1.5,  

### 2012-05-22 scripts/CompareDiscretizationSchemes.py

@@ -1,0 +1,79 @@
+"""
+
+CompareDiscretizationSchemes.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on: Tue May 22 09:25:38 2012
+
+Description: Compare our optimal discrete SEDs to those constructed via
+effective column density scheme of Aubert & Teyssier (2008). Plot modeled after
+A&T Figure 4 (Test #2 from RT06).
+
+"""
+
+import numpy as np
+import pylab as pl
+import rt1d.analysis as rta
+
+continuous = rta.Analyze('./continuous.dat')
+optimal_mono = rta.Analyze('./optimal_monochromatic.dat')
+effective_mono = rta.Analyze('./effective_xsection_monochromatic.dat')
+
+Lbox = continuous.pf.LengthUnits
+
+# Construct mask
+mask = np.arange(0, len(continuous.data[0].r) - 1)
+mask[mask % 2 != 0] = 0
+mask[mask > 0] = 1
+mask[1] = 1    
+mask[-1] = 1  
+
+mp = rta.multiplot(dims = (1, 2), useAxesGrid = False, panel_size = (2, 1), 
+        share_all = False, aspect = False, padding = 0.2)
+
+mp.grid[0].plot(continuous.data[1].r / Lbox, np.log10(continuous.data[1].T), color = 'k', ls = '-')
+#mp.grid[0].plot(continuous.data[7].r / Lbox, np.log10(continuous.data[7].T), color = 'k', ls = '--')
+mp.grid[0].plot(continuous.data[20].r / Lbox, np.log10(continuous.data[20].T), color = 'k', ls = ':')
+
+mp.grid[0].scatter(optimal_mono.data[1].r[mask == 1] / Lbox, np.log10(optimal_mono.data[1].T[mask == 1]), color = 'b', marker = 'o', s = 20, facecolors = 'none')
+#mp.grid[0].scatter(optimal_mono.data[7].r[mask == 1] / Lbox, np.log10(optimal_mono.data[7].T[mask == 1]), color = 'b', marker = '^', s = 20, facecolors = 'none')
[…]

### 2012-05-23 doc/manual/rt1d_manual.tex
Minor changes. About to rework lookup table initialization and interpolation to accomodate higher (4-5) dimensional tables.

@@
 \usepackage[pdfborder={0 0 1}]{hyperref}
+\usepackage[htt]{hyphenat}
 
@@
 
+% Table of contents
 \setcounter{tocdepth}{2}
@@
 Let's see what happens to the gas around stars and black holes, shall we? The
-primary reference for this code is \cite[\href{http://arxiv.org/abs/1204.1944}{arxiv:1204.1944}]{Mirocha2012} -- much of the text has
-been extracted from there.
+primary reference for this code is
+\cite[\href{http://arxiv.org/abs/1204.1944}{arxiv:1204.1944}]{Mirocha2012} --
+much of the text has been extracted from there.
 
@@
 $\HeIII$. Associating the index $i$ with absorbing species, $i = \HI, \HeI,
-\HeII$, we define $\Gamma_i$ as the photo-ionization rate coefficient,
-$\gamma_i$ as the secondary ionization rate coefficient, $\alpha_i$ ($\xi_i$)
-as the case-B (dielectric) recombination rate coefficients, $\beta_i$ as the
-collisional ionization rate coefficients, and $\nel = \nHII + \nHeII +
-2\nHeIII$ as the number density of electrons.
+\HeII$, and the index $i^{\prime}$ with ions, $i^{\prime} = \HII, \HeII,
+\HeIII$, we define $\Gamma_i$ as the photo-ionization rate coefficient,
+$\gamma_i$ as the secondary ionization rate coefficient, $\alpha_{i^{\prime}}$
+($\xi_{i^{\prime}}$) as the case-B (dielectric) recombination rate
+coefficients, $\beta_i$ as the collisional ionization rate coefficients, and
+$\nel = \nHII + \nHeII + 2\nHeIII$ as the number density of electrons.
 
@@
 \begin{align}
-    \frac{3}{2}\frac{d}{dt}\left(\frac{\kB T_k \ntot}{\mu}\right) & = \fheat  \sum_i n_i \Heat_i - \sum_i \zeta_i n_e n_i - \sum_{i^{\prime}} \eta_i n_e n_i \nonumber \\ & - \sum_i \psi_i n_e n_i - \omega_{\HeII} n_e \nHeII \label{eq:TemperatureEvolution}
+    \frac{3}{2}\frac{d}{dt}\left(\frac{\kB T_k \ntot}{\mu}\right) & = \fheat  \sum_i n_i \Heat_i - \sum_i \zeta_i n_e n_i - \sum_{i^{\prime}} \eta_{i^{\prime}} n_e n_{i^{\prime}} \nonumber \\ & - \sum_i \psi_i n_e n_i - \omega_{\HeII} n_e \nHeII \label{eq:TemperatureEvolution}
 \end{align}
-where the index $i$ corresponds as usual to species that have bound electrons,
-and $i^{\prime}$ corresponds to ionized species. $\Heat_i$ is the
-photo-electric heating rate coefficient (due to electrons previously bound to
-species $i$), $\omega_{\HeII}$ is the dielectric recombination cooling
-coefficient, and $\zeta_i$, $\eta_i$, and $\psi_i$ are the collisional
-ionization, recombination, and collisional excitation cooling coefficients,
-respectively. The constants in Equation \ref{eq:TemperatureEvolution} are the
-total number density of baryons, $\ntot = n_\mathrm{H} + n_{\mathrm{He}} +
-n_e$, the mean molecular weight, $\mu$, Boltzmann's constant, $\kB$, and the
-fraction of secondary electron energy deposited as heat, $\fheat$. We use the
-formulae in Appendix B of \citet{Fukugita1994} to compute the values of
-$\alpha_i$, $\beta_i$, $\xi_i$, $\zeta_i$, $\eta_i$, $\psi_i$, and
-$\omega_{\HeII}$.
+where $\Heat_i$ is the photo-electric heating rate coefficient (due to
+electrons previously bound to species $i$), $\omega_{\HeII}$ is the dielectric
+recombination cooling coefficient, and $\zeta_i$, $\eta_{i^{\prime}}$, and
+$\psi_i$ are the collisional ionization, recombination, and collisional
+excitation cooling coefficients, respectively. The constants in Equation
+\ref{eq:TemperatureEvolution} are the total number density of baryons, $\ntot
+= n_\mathrm{H} + n_{\mathrm{He}} + n_e$, the mean molecular weight, $\mu$,
+Boltzmann's constant, $\kB$, and the fraction of secondary electron energy
+deposited as heat, $\fheat$. We use the formulae in Appendix B of
+\citet{Fukugita1994} to compute the values of $\alpha_i$, $\beta_i$, $\xi_i$,
+$\zeta_i$, $\eta_{i^{\prime}}$, $\psi_i$, and $\omega_{\HeII}$.
 
@@
 concentric spherical shells, each having thickness $\Delta r$ and volume
-$V_{\sh}(r) = 4 \pi [(r + \Delta r)^3 - r^3] / 3$ (i.e. radii correspond to
-cell interfaces). The ionization and heating rates can then be related to the
-number of absorptions in any given shell (thus preserving photon number), as
+$V_{\sh}(r) = 4 \pi [(r + \Delta r)^3 - r^3] / 3$, where $r$ is the distance
+between the origin and the inner interface if each shell. The ionization and
+heating rates can then be related to the number of absorptions in any given
+shell (thus preserving photon number), as
 \begin{align}
@@
 \end{align}    
-where we've defined the normalization constant, $A_i \equiv
+where we have defined the normalization constant $A_i \equiv
 L_{\mathrm{bol}}/n_i V_{\sh}(r)$, and denote the ionization threshold energy
-for species $i$ as $h\nu_i$.
+for species $i$ as $h\nu_i$.  $I_{\nu}$ represents the SED of radiation sources, and satisfies 
+\begin{equation}
+\int_{\nu} I_{\nu} d\nu = 1,
+\end{equation}
+such that $A_i I_{\nu}$ is the number of photons of frequency $\nu$ per atom (of species $i$) in a shell at radius $r$.
 
@@
 
+% Time-dependent source SEDs
+\subsection{Additional Complexity: Time-Dependent Source SEDs}
+Our lookup tables will inherit an additional dimension.  Goodie.
+
+% Spectrum Normalization
+\subsection{Spectrum Normalization}
+For convenience, we generally work with the normalized SED $I_{\nu}$, such that 
+\begin{equation}
+    L_{\nu} = L_{\mathrm{bol}} I_{\nu} ,
+\end{equation}    
+and 
+\begin{equation}
+    \dot{Q}_{\nu} = L_{\nu} / h\nu .
+\end{equation}    
+However, in \S\ref{sec:Discretization}, we will discuss how \texttt{rt1d} can be used to verify choices of discrete SEDs -- when \texttt{DiscreteSpectrum = 1}, we must be careful about how we normalize a spectrum.
+
+There are two primary ways we can go about normalizing a discrete spectrum consisting of $n_{\nu}$ frequency bins.
+
+The first way is to manually supply the location of each bin (in eV) and the fraction of the bolometric luminosity allocated to that bin via the parameters \texttt{DiscreteSpectrumSED} and \texttt{DiscreteSpectrumRelLum}.
+In \citet{Mirocha2012}, we denote these quantities as $h\nu_n$ and $I_n$, respectively, where $n = 1, 2,...n_{\nu}$, and do \textit{not} require the bolometric luminosity to remain conserved.
+That is, we allow
+\begin{equation}
+    \sum_n I_n \neq 1.
+\end{equation}    
+This is because very high energy photons (depending on the spectrum) may not be absorbed at all over the column density regime or volume of interest.
+It is important to note that even if $\sum_n I_n = 1$ exactly, the photon emissivity of a source constructed in this way will \textit{not} be conserved.
+
+For example, when we simulate a $10^5$ K blackbody source emitting $5\times 10^{48}$ ionizing photons per second (as in RT06 problem \#2), the \texttt{RadiationSource} class figures out the bolometric luminosity (show equations here for reference), and assigns a fraction $I_n$ of that to bin $n$.
+In doing this,
+\begin{equation}
+    L_{\mathrm{bol}} \sum_n I_n / h\nu_n \neq 5 \times 10^{48} \ \mathrm{s}^{-1} .
+\end{equation}
+
+The second approach we can use is that used in previous work such as \citet{Aubert2008}, which requires us to define bandpasses, each defined by an average photon energy and effective bound-free absorption cross-section.
+The details will be discussed in \S\ref{sec:Discretization}, but for now it is enough to know that the ionizing photon luminosity \textit{will} be preserved with this method, assuming that the bandpasses span the full emission interval of the radiation source of interest, and the bolometric luminosity will \textit{not} be preserved.
+That is,
+\begin{equation}
+    \sum_n \dot{Q}_n h \nu_n \neq L_{\mathrm{bol}} .
+\end{equation}
+
 \subsection{Time-stepping Techniques}
@@
 \subsection{Parallelism}
+
+% Lookup Tables
+\subsection{Lookup Tables}
+How does the dimensionality of our lookup tables depend on input parameters?
 
@@
 %%%
-\section{Verifying Discrete SEDs with \texttt{rt1d}}
-In \citet{Mirocha2012}, we used \texttt{rt1d} to verify new, optimally selected discrete spectral energy distributions for sources of ultraviolet and X-ray photons.
-As we optimize more and more SEDs for fully 3D cosmological  radiation-hydro simulations, \texttt{rt1d} remains a quick and easy tool for verifying that the SEDs we're using are acceptable.
+\section{Verifying Discrete SEDs with \texttt{rt1d}} \label{sec:Discretization}
+In \citet{Mirocha2012}, we used \texttt{rt1d} to verify new, optimally
+selected discrete spectral energy distributions for sources of ultraviolet and
+X-ray photons. As we optimize more and more SEDs for fully 3D cosmological
+radiation-hydro simulations, \texttt{rt1d} remains a quick and easy tool for
+verifying that the SEDs we're using are acceptable.
+
+In the literature, discrete SEDs have been constructed using a `multi-group'
+approach, which defines new absorption cross sections which represent
+spectrum-weighted averages over broad bandpasses,
+\begin{equation}
+    \tilde{\sigma}_{i,n} \equiv \int_{\nu_n}^{\nu_{n+1}} \sigma_{i,\nu} \frac{I_{\nu}}{h\nu}  d\nu \left(\int_{\nu_n}^{\nu_{n+1}}\frac{I_{\nu}}{h\nu}d\nu \right)^{-1} ,
+\end{equation}    
+where the interval $(\nu_n, \nu_{n+1})$ represents the bandpass bounds for
+discrete photon group $n$.
+
+The emission energies are chosen in a similar way, 
+\begin{equation}
+    h\nu_n = \int_{h\nu_n}^{h\nu_{n+1}} I_{\nu} d\nu \left(\int_{h\nu_n}^{h\nu_{n+1}} \frac{I_{\nu}}{h\nu} d\nu \right)^{-1} . \label{eq:MeanPhotonEnergy}
+\end{equation}    
+It's worth keeping in mind that there are different ways to compute the mean
+photon energy in each interval. Here, we've calculated by dividing a quantity
+proportional to the bolometric energy output by the total number of ionizing
+photons. We could also compute the expectation value, median value, or even
+the peak of emission (e.g. for a blackbody). Equation
+\ref{eq:MeanPhotonEnergy} is most often used in the literature, so we'll stick
+to that.
 

### 2012-05-23 rt1d/mods/CheckForParameterConflicts.py

@@ -38,8 +38,12 @@ def CheckForParameterConflicts(pf):
         
     if probs:
         msg = []
-        for con in probs:
-            msg.append('%s = %g and %s = %g' % (con[0][0], con[0][1], con[1][0], con[1][1]))
+        for i, con in enumerate(probs):
+            for element in con:
+                msg.append('%s = %g' % (element[0], element[1]))
+            
+            if len(probs) > 1 and i != len(probs):
+                msg.append('\nAND\n')
         
         return True, msg
     else:

### 2012-05-23 rt1d/mods/Constants.py

@@ -62,8 +62,8 @@ erg_per_j = 1e-7
 erg_per_ev = e/erg_per_j
 erg_per_kev = erg_per_ev*1e3
 
-lsun = 3.839e33                                             # Solar luminosity - erg / s
-cm_per_rsun = 695500.0 * 1e5                                # Radius of the sun - [cm_per_rsun] = cm
-t_edd = 0.45 * 1e9 * s_per_yr                               # Eddington timescale (see eq. 1 in Volonteri & Rees 2005) 
-
-xi = (3. / 7.)**0.5 * (6. / 7.)**3
+lsun = 3.839e33                     # Solar luminosity - erg / s
+cm_per_rsun = 695500.0 * 1e5        # Radius of the sun - [cm_per_rsun] = cm
+t_edd = 0.45 * 1e9 * s_per_yr       # Eddington timescale (see eq. 1 in Volonteri & Rees 2005) 
+                                    
+xi = (3. / 7.)**0.5 * (6. / 7.)**3  # Correction factor for peak temperature location in disk                        

### 2012-05-23 rt1d/mods/InitializeIntegralTables.py

@@ -109,7 +109,7 @@ class InitializeIntegralTables:
         Returns the filename following the convention:
                 
         filename = SourceType_UniqueSourceProperties_PhotonConserving_MultiSpecies_
-            DiscreteOrContinuous_TableDims_NHlimits_NHelimits.h5
+            DiscreteOrContinuous_TimeDependent?_TableDims_NHlimits_NHelimits.h5
         
         """
         
@@ -117,7 +117,8 @@ class InitializeIntegralTables:
             return self.pf.IntegralTableName
               
         ms = 'ms%i' % self.pf.MultiSpecies
-        pc = 'pc%i' % self.pf.PhotonConserving        
+        pc = 'pc%i' % self.pf.PhotonConserving 
+        td = 'tdep%i' % self.pf.SourceTimeEvolution       
         
         if self.pf.DiscreteSpectrum:
             sed = 'D'
@@ -162,7 +163,7 @@ class InitializeIntegralTables:
         Hlim = '%i%i' % (self.HIColumn[0], self.HIColumn[-1])
         Helim = '%i%i' % (self.HeIColumn[0], self.HeIColumn[-1])        
                     
-        return "%s_%s_%s_%s_%s_%s_%s_%s.h5" % (src, prop, pc, ms, sed, dims, Hlim, Helim)
+        return "%s_%s_%s_%s_%s_%s_%s_%s_%s.h5" % (src, prop, pc, ms, sed, td, dims, Hlim, Helim)
             
     def DatasetName(self, integral, species, donor_species = None):
         """
@@ -450,6 +451,28 @@ class InitializeIntegralTables:
             integrals.extend(['PhiWiggle', 'PsiWiggle', 'PhiHat', 'PsiHat'])
         
         return integrals
+    
+    def Dimensionality(self):
+        """
+        Return dimensionality of lookup table.
+        """    
+        
+        dims = 0
+          
+        if self.pf.SecondaryIonization >= 2:
+            dims += 1
+            
+        if self.pf.SourceTimeEvolution:
+            dims += 1
+            
+        if self.pf.MultiSpecies >= 1:
+            dims += 2            
+            
+        # Reserved for H2    
+        if self.pf.MultiSpecies == 2: 
+            dims += 1   
+        
+        return dims    
         
     def TotalOpticalDepth(self, ncol, species = None):
         """

### 2012-05-23 rt1d/mods/RadiationSource.py

@@ -18,7 +18,7 @@ SourceType = 0: Monochromatic/Polychromatic emission of SpectrumPhotonLuminosity
 """
 
 import numpy as np
-from scipy.integrate import quad, fixed_quad
+from scipy.integrate import quad
 from Integrate import simpson as integrate                  # why did scipy.integrate.quad mess up LphotNorm? 
 from .Constants import *
 from .ReadParameterFile import dotdictify
@@ -60,7 +60,9 @@ class RadiationSource:
         self.SpectrumPars.Type = map(int, self.SpectrumPars.Type)
         
         self.Emin = min(self.SpectrumPars.MinEnergy)
-        self.Emax = min(self.SpectrumPars.MaxEnergy)        
+        self.Emax = min(self.SpectrumPars.MaxEnergy)   
+        self.EminNorm = min(self.SpectrumPars.MinNormEnergy)
+        self.EmaxNorm = min(self.SpectrumPars.MaxNormEnergy)
         
         if self.N == 1:
             self.Type = self.SpectrumPars.Type[0]
@@ -132,18 +134,24 @@ class RadiationSource:
         self.LuminosityNormalizations = self.NormalizeSpectrumComponents(0.0)
         
         # Multi-group treatment
+        Qnorm = pf.SpectrumPhotonLuminosity / self.Lbol / \
+            quad(lambda x: self.Spectrum(x) / x, self.EminNorm, self.EmaxNorm)[0]
         if pf.FrequencyAveragedCrossSections:
             self.E = np.zeros(pf.FrequencyGroups)
             self.F = np.ones_like(self.E)
+            self.Qdot = np.zeros_like(self.E)
             self.bands = pf.FrequencyBands
+            
             if len(self.bands) == (len(self.E) - 1):
                 self.bands.append(self.Emax)
             
             for i in xrange(pf.FrequencyGroups):
-                self.E[i] = quad(lambda x: x * self.Spectrum(x), 
-                    self.bands[i], self.bands[i + 1])[0]
-                self.Qdot = np.array([pf.SpectrumPhotonLuminosity]) #self.Lbol * self.F / self.E / erg_per_ev    
-                              
+                L = quad(lambda x: self.Spectrum(x), self.bands[i], self.bands[i + 1])[0]
+                Q = quad(lambda x: self.Spectrum(x) / x, self.bands[i], self.bands[i + 1])[0]
+                
+                self.E[i] = L / Q
+                self.Qdot[i] = Qnorm * self.Lbol * Q
[…]

### 2012-05-23 scripts/CompareDiscretizationSchemes.py

@@ -33,8 +33,8 @@ mp = rta.multiplot(dims = (1, 2), useAxesGrid = False, panel_size = (2, 1),
         share_all = False, aspect = False, padding = 0.2)
 
 mp.grid[0].plot(continuous.data[1].r / Lbox, np.log10(continuous.data[1].T), color = 'k', ls = '-')
-#mp.grid[0].plot(continuous.data[7].r / Lbox, np.log10(continuous.data[7].T), color = 'k', ls = '--')
-mp.grid[0].plot(continuous.data[20].r / Lbox, np.log10(continuous.data[20].T), color = 'k', ls = ':')
+#mp.grid[0].plot(continuous.data[7].r / Lbox, np.log10(continuous.data[7].T), color = 'k', ls = '-')
+mp.grid[0].plot(continuous.data[20].r / Lbox, np.log10(continuous.data[20].T), color = 'k', ls = '-')
 
 mp.grid[0].scatter(optimal_mono.data[1].r[mask == 1] / Lbox, np.log10(optimal_mono.data[1].T[mask == 1]), color = 'b', marker = 'o', s = 20, facecolors = 'none')
 #mp.grid[0].scatter(optimal_mono.data[7].r[mask == 1] / Lbox, np.log10(optimal_mono.data[7].T[mask == 1]), color = 'b', marker = '^', s = 20, facecolors = 'none')

### 2012-06-03 rt1d/__init__.py
Trying to get finite-c working again...runs, needs more testing.  Other random cleanup too.

@@ -1,5 +1,6 @@
 """
-Import main driver module.
+Import everything.
 """
 
-from .mods import RT1D
+from .mods import *
+from .mods.RT1D import Shine

### 2012-06-03 rt1d/mods/RT1D.py

@@ -24,7 +24,7 @@ except ImportError:
 # Class definitions
 import rt1d.mods as rtm
     
-def Shine(pf, r = None, IsRestart = False):
+def Shine(pf, r = None, data = None, IsRestart = False):
     """
     Initialize grid, radiation source, integral tables (maybe), and evolve
     radiation from CurrentTime to StopTime.
@@ -64,6 +64,9 @@ def Shine(pf, r = None, IsRestart = False):
             return
               
         # Initialize grid and file system
+        if data is not None:
+            data = data
+            g = rtm.InitializeGrid(pf)
         if IsRestart: 
             data = ICs
             g = rtm.InitializeGrid(pf)
@@ -84,7 +87,7 @@ def Shine(pf, r = None, IsRestart = False):
     
         # Initialize integral tables
         iits = rtm.InitializeIntegralTables(pf)
-        if pf.TabulateIntegrals:
+        if pf.TabulateIntegrals and iits.itabs:
             itabs = iits.TabulateRateIntegrals()        
             
             if pf.ExitAfterIntegralTabulation: 
@@ -211,8 +214,11 @@ def Shine(pf, r = None, IsRestart = False):
         print >> f, elapsed
         print >> f, ''
         f.close()
-    
-    print "Successful run. Exiting."
+            
+    if __name__ != '__main__':
+        return data
+    else:
+        print "Successful run. Exiting."
 
 if __name__ == '__main__':
     

### 2012-06-03 rt1d/mods/Radiate.py

@@ -138,6 +138,7 @@ class Radiate:
         for i, col in enumerate(self.ncol_all):
             self.indices_all.append(self.coeff.Interpolate.GetIndices([col[0], col[1], col[2], np.log10(self.x_HII_arr[i]), t]))
                                                 
+        # PUT THIS ELSEWHERE                                        
         # Compute optical depths *between* source and all cells. Do we only use this for timestep calculation?
         self.tau_all_arr = np.zeros([3, self.GridDimensions])    
         if not self.pf['TabulateIntegrals']:
@@ -351,12 +352,12 @@ class Radiate:
         else: 
             packs = list(self.data['PhotonPackages']) 
         
-        if packs and self.OnePhotonPackagePerCell:
+        if packs and self.pf.OnePhotonPackagePerCell:
             t_packs = np.array(zip(*packs)[0])    # Birth times for all packages
             r_packs = (t - t_packs) * c
             
             # If there are already photon packages in the first cell, add energy in would-be packet to preexisting one
-            if r_packs[-1] < (self.StartRadius * self.LengthUnits):
+            if r_packs[-1] < (self.pf.StartRadius * self.pf.LengthUnits):
                 packs[-1][-1] += Lbol * dt
             # Launch new photon package
             else:
@@ -371,8 +372,8 @@ class Radiate:
             r_max = r_pack + dt * c           # Furthest this package will get this timestep
                                                                              
             # Cells we need to know about - not necessarily integer
-            cell_pack = r_pack * self.GridDimensions / self.LengthUnits
-            cell_pack_max = r_max * self.GridDimensions / self.LengthUnits
+            cell_pack = r_pack * self.pf.GridDimensions / self.pf.LengthUnits
+            cell_pack_max = r_max * self.pf.GridDimensions / self.pf.LengthUnits
                           
             Lbol = pack[-1] / pack[1]          
                                             
@@ -382,7 +383,7 @@ class Radiate:
                 # What cell are we in
                 cell = int(round(cell_pack))
                 
-                if cell >= self.GridDimensions: 
+                if cell >= self.pf.GridDimensions: 
                     break
                 
                 # Compute dc (like dx but in fractional cell units)
@@ -398,9 +399,9 @@ class Radiate:
                 else: 
                     subdt = dt
[…]

### 2012-06-03 rt1d/mods/SetDefaultParameterValues.py

@@ -12,6 +12,7 @@ the parameter space.
 """
 
 from .Constants import *
+from .Extras import dotdictify
 
 def SetDefaultParameterValues():
     pf = \
@@ -107,6 +108,7 @@ def SetDefaultParameterValues():
         "InfiniteSpeedOfLight": 1,  
         "PlaneParallelField": 0,  
         "ClumpingFactor": 1,
+        "RecombinationMethod": 'CaseB',
         
         # Initial conditions
         "InitialRedshift": 20.0, 
@@ -165,4 +167,4 @@ def SetDefaultParameterValues():
           
         }
 
-    return pf
+    return dotdictify(pf)

### 2012-06-03 rt1d/mods/RT1D.py
Working on finite c solver - error found in treatment of first cell.

@@ -216,7 +216,7 @@ def Shine(pf, r = None, data = None, IsRestart = False):
         f.close()
             
     if __name__ != '__main__':
-        return data
+        return data, dt
     else:
         print "Successful run. Exiting."
 

### 2012-06-03 rt1d/mods/Radiate.py

@@ -138,32 +138,9 @@ class Radiate:
         for i, col in enumerate(self.ncol_all):
             self.indices_all.append(self.coeff.Interpolate.GetIndices([col[0], col[1], col[2], np.log10(self.x_HII_arr[i]), t]))
                                                 
-        # PUT THIS ELSEWHERE                                        
-        # Compute optical depths *between* source and all cells. Do we only use this for timestep calculation?
-        self.tau_all_arr = np.zeros([3, self.GridDimensions])    
-        if not self.pf['TabulateIntegrals']:
-            sigma0 = PhotoIonizationCrossSection(self.rs.E, species = 0)
-            tmp_nHI = np.transpose(len(self.rs.E) * [self.ncol_HI])            
-            self.tau_all_arr[0] = np.sum(tmp_nHI * sigma0, axis = 1)
-                        
-            if self.pf.MultiSpecies:
-                sigma1 = PhotoIonizationCrossSection(self.rs.E, species = 1)
-                sigma2 = PhotoIonizationCrossSection(self.rs.E, species = 2)
-                tmp_nHeI = np.transpose(len(self.rs.E) * [self.ncol_HeI])
-                tmp_nHeII = np.transpose(len(self.rs.E) * [self.ncol_HeII])
-                self.tau_all_arr[1] = np.sum(tmp_nHeI * sigma1, axis = 1)
-                self.tau_all_arr[2] = np.sum(tmp_nHeII * sigma2, axis = 1)
-        else:
-            for i, col in enumerate(np.log10(self.ncol_HI)):
-                self.tau_all_arr[0][i] = self.coeff.Interpolate.OpticalDepth(col, 0)
-                
-                if self.pf.MultiSpecies:
-                    self.tau_all_arr[1][i] = self.coeff.Interpolate.OpticalDepth(np.log10(self.ncol_HeI[i]), 1)
-                    self.tau_all_arr[2][i] = self.coeff.Interpolate.OpticalDepth(np.log10(self.ncol_HeII[i]), 2)
-                    
-            self.tau_all_arr[0][0] = self.tau_all_arr[1][0] = self.tau_all_arr[2][0] = neglible_tau 
-
-        self.tau_all = zip(*self.tau_all_arr) 
+        # Compute tau *between* source and all cells
+        tau_all_arr = self.ComputeOpticalDepths([self.ncol_HI, self.ncol_HeI, self.ncol_HeII])
+        self.tau_all = zip(*tau_all_arr) 
                                         
[…]

### 2012-06-06 doc/manual/rt1d_manual.tex
Trying to get finite-c working again...runs, needs more testing.  Other random cleanup too.
Working on finite c solver - error found in treatment of first cell.
Time-dependent SED stuff should now be handled via object age, instead of BH mass or whathaveyou.

@@
 import rt1d
-import numpy as np
+
 pf = rt1d.SetDefaultParameterValues()
-r = np.linspace(3e19, 3e21, 100) # 100 grid cells between 0.01-1kpc. (optional)
-data = rt1d.Shine(pf, r)    
+prob = rt1d.ProblemTypes.ProblemType(1)
+
+for key in pf.keys():
+    if key in prob.keys():
+        pf[key] = prob[key]
+
+pf["StopTime"] = 10
+
+data, dt = rt1d.Shine(pf)   
 \end{verbatim}
@@
     \item Very specific initial conditions.
+    \item Run rad-hydro with \texttt{proteus}.
 \end{itemize}
@@
 
+%%%
+%% Diffuse Radiation
+%%%
+\section{Diffuse Radiation}
+Let's figure out where the on-the-spot approximation is pretty good.  This will depend on photon frequency, gas neutral fraction (of hydrogen and helium), and our grid resolution.
+
+The optical depth along a ray is
+\begin{align}
+    \tau_{i,\nu} & = N_{i,\nu} \sigma_{i,\nu} \nonumber \\
+    & = n_i dx \sigma_{i,\nu}
+\end{align}    
+where $n_i$ is the number density of absorbing species $i$, $\sigma_{i, \nu}$ is the bound-free absorption cross section for species $i$, and $dx$ is the length of the ray segment (i.e.
+our grid resolution).
+So, the total optical depth along this ray segment is
+\begin{equation}
+    \tau_{\nu} = \sum_i \tau_{i, \nu} .
+\end{equation}    
+If we assume that $\xHII = \xHeII$ and $\xHeIII = 0$, we can plot the ionized fraction vs. $\tau_{\nu}$ to get a handle on how good the OTSA is as a function of grid resolution and frequency.
+
 \newpage

### 2012-06-06 rt1d/mods/RadiationSource.py

@@ -154,12 +154,7 @@ class RadiationSource:
                 
         # Time evolution
         if pf.SourceTimeEvolution:
-            self.logMarr = np.linspace(np.log10(pf.SourceMass), np.log10(pf.SourceFinalMass), pf.MassBins)
-            self.Marr = 10**self.logMarr     
-            self.Age = self.BlackHoleAge(self.Marr)
-            self.logAge = np.log10(self.Age)
-            self.Age[0] = 0
-            self.logAge[0] = 1  # seconds
+            self.Age = np.linspace(0, pf.StopTime * pf.TimeUnits, pf.AgeBins)
                                               
     def GravitationalRadius(self, M):
         """

### 2012-06-08 rt1d/analysis/Inspection.py
Need to update rs.Spectrum attributes for t == 0 sometimes (like in integral tabulation)...

@@ -100,7 +100,7 @@ class Inspect:
         pl.draw()
         
     def InspectInterpolation(self, intnum = 1, species = 0, donor_species = 0,
-        nHI = np.linspace(12, 19, 1000), 
+        nHI = np.linspace(12, 19, 1000), t = 0,
         nHeI = 0.0, nHeII = 0.0, color = 'b', ls = '-', x_HII = 1e-4):
         """
         Now check how good our interpolation is...
@@ -143,10 +143,10 @@ class Inspect:
         if type(nHI) not in [int, float]:
             x = nHI
             for col in nHI:
-                tmp = [col, nHeI, nHeII]                
-                indices = self.interp.GetIndices(tmp, x_HII = x_HII)                
+                tmp = [col, nHeI, nHeII, np.log10(x_HII), t]                
+                indices = self.interp.GetIndices(tmp)                
                 result.append(self.interp.interp(indices, 
-                    '%s' % integral, tmp, x_HII = x_HII))
+                    '%s' % integral, tmp))
             
         elif type(nHeI) not in [int, float]:
             x = nHeI

### 2012-06-08 rt1d/mods/ControlSimulation.py

@@ -49,7 +49,7 @@ class ControlSimulation:
         nion = np.array([data['HIIDensity'][0], data['HeIIDensity'][0], data['HeIIIDensity'][0]])
         T = data['Temperature'][0]
         
-        ncol = 11 * np.ones(3)
+        ncol = 1. * np.ones(3)
         ncell = r.dx[0] * nabs    
         nout = np.log10(ncell)
         Vsh = r.coeff.ShellVolume(self.R0, r.dx[0])    
@@ -101,7 +101,8 @@ class ControlSimulation:
             dTdt = np.abs(2. * T * hubble)
             dtT = self.pf.MaxTemperatureChange * T / dTdt    
 
-        return min(dtHI, dtHeII, dtHeIII, dtne, dtT)
+        return min(dtHI, dtHeII, dtHeIII, dtne, dtT, 
+            self.pf.MaximumGlobalTimestep * self.pf.TimeUnits)
     
     def ComputePhotonTimestep(self, tau, nabs, nion, ncol, n_H, n_He, n_e, n_B, 
         Gamma, gamma, Beta, alpha, k_H, zeta, eta, psi, xi, omega, hubble, compton, 
@@ -120,7 +121,7 @@ class ControlSimulation:
                 dHIIdt -= 3. * nabs[0] * hubble
             if tau[0] >= self.pf.OpticalDepthDefiningIFront[0]:
                 dtHI = self.pf.MaxHIIChange * nabs[0] / abs(dHIIdt)
-        
+                    
         dtHeII = 1e50 
         if self.pf.MultiSpecies and self.pf.HeIIRestrictedTimestep:
             dHeIIdt = nabs[1] * (Gamma[1] + Beta[1] * n_e) + \

### 2012-06-08 rt1d/mods/Interpolate.py

@@ -51,7 +51,7 @@ class Interpolate:
     def OpticalDepth(self, value, species):
         return 10**np.interp(value, self.columns[species], self.itabs['logOpticalDepth%i' % species])
                 
-    def InterpolateLinear(self, indices, integral, value, x_HII = None):
+    def InterpolateLinear(self, indices, integral, value):
         """
         Use this technique for hydrogen-only calculations.  For consistency with MultiSpecies > 0 
         methods, 'value' should still be a 3-element list.  
@@ -66,7 +66,7 @@ class Interpolate:
         """    
         
         i_x, i_y = indices      
-                                                                                
+                                                                                        
         x1 = self.columns[0][i_x]
         x2 = self.columns[0][i_x + 1]
         y1 = self.columns[1][i_y]
@@ -82,9 +82,7 @@ class Interpolate:
             f12 * (x2 - value[self.locs[0]]) * (value[self.locs[1]] - y1) + \
             f22 * (value[self.locs[0]] - x1) * (value[self.locs[1]] - y1)) / \
             (x2 - x1) / (y2 - y1)
-                        
-        print integral, value[self.locs[0]], value[self.locs[1]], x1, x2, y1, y2, f11, f21, f12, f22, final
-                        
+                                                                                         
         return 10**final    
         
     def InterpolateTriLinear(self, indices, integral, value):
@@ -134,7 +132,7 @@ class Interpolate:
                 
         i1 = int((value[self.locs[0]] - self.colmin[0]) / self.dcolumns[0])
         i2 = int((value[self.locs[1]] - self.colmin[1]) / self.dcolumns[1])
-        
+                
         return min(max(i1, 0), self.dims[0] - 2), min(max(i2, 0), self.dims[1] - 2)
         
     def GetIndices3D(self, value):

### 2012-06-08 rt1d/mods/Radiate.py

@@ -43,7 +43,7 @@ except ImportError:
 E_th = [13.6, 24.6, 54.4]
 
 neglible_tau = 1e-12
-neglible_column = 1e11
+neglible_column = 10
 
 class Radiate:
     def __init__(self, pf, iits): 
@@ -262,7 +262,7 @@ class Radiate:
             # Unpack so we have everything by name
             nabs, nion, n_H, n_He, n_e, Gamma, gamma, Beta, alpha, \
                 k_H, zeta, eta, psi, xi, omega, hubble, compton = args                                                   
-                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         
+                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               
             ######################################
             ######## Solve Rate Equations ########
             ######################################        
@@ -628,7 +628,7 @@ class Radiate:
         dqdt[0] = (Gamma[0] + Beta[0] * n_e) * nHI + \
                   (gamma[0][0] * nHI + gamma[0][1] * nHeI + gamma[0][2] * nHeII) - \
                    alpha[0] * n_e * q[0]
-
+                
         if self.pf.CosmologicalExpansion:
             dqdt[0] -= 3. * q[0] * hubble
 

### 2012-12-18 doc/tests/rt06_2_He/AnalyzeTest2_He.py
Changed to README.rst

@@ -1,0 +1,78 @@
+"""
+
+AnalyzeTest2_He.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on: Wed Aug 29 11:53:44 2012
+
+Description: Compare RT06 #2 solutions with and without including helium.
+
+"""
+
+import numpy as np
+import pylab as pl
+import rt1d.analysis as rta
+
+dsH = rta.Analyze('./rt06_2_H.dat')
+dsHe = rta.Analyze('./rt06_2_He.dat')
+
+# 4-panel plot: ionization (top) and heating (bottom), 10 Myr (left), 100 Myr (right)
+mp = rta.multiplot(dims = (2, 2), panel_size = (2, 2), useAxesGrid = False)
+
+### UPPER PANELS
+mp.grid[0].semilogy(dsH.data[0].r / rta.cm_per_kpc / 6.6, dsH.data[1].x_HII,   
+    color = 'k', ls = '--')
+mp.grid[0].semilogy(dsH.data[0].r / rta.cm_per_kpc / 6.6, dsH.data[1].x_HI,   
+    color = 'k', ls = '-')        
+mp.grid[0].semilogy(dsHe.data[0].r / rta.cm_per_kpc / 6.6, dsHe.data[1].x_HII,   
+    color = 'b', ls = '--')
+mp.grid[0].semilogy(dsHe.data[0].r / rta.cm_per_kpc / 6.6, dsHe.data[1].x_HI,   
+    color = 'b', ls = '-')                          
+mp.grid[1].semilogy(dsH.data[0].r / rta.cm_per_kpc / 6.6, dsH.data[10].x_HI,   
+    color = 'k', ls = '-', label = r'$x_{\mathrm{HI}}$')   
+mp.grid[1].semilogy(dsH.data[0].r / rta.cm_per_kpc / 6.6, dsH.data[10].x_HII,   
+    color = 'k', ls = '--', label = r'$x_{\mathrm{HII}}$')
+mp.grid[1].semilogy(dsHe.data[0].r / rta.cm_per_kpc / 6.6, dsHe.data[10].x_HI,   
+    color = 'b', ls = '-')   
+mp.grid[1].semilogy(dsHe.data[0].r / rta.cm_per_kpc / 6.6, dsHe.data[10].x_HII,   
+    color = 'b', ls = '--')    
+  
+for i in xrange(2):
[…]

### 2012-12-18 rt1d/analysis/Analysis.py

@@ -1,0 +1,573 @@
+"""
+Analysis.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on 2011-06-17.
+
+Description: Functions to calculate various quantities from our rt1d datasets.
+     
+"""
+
+import os
+import numpy as np
+import pylab as pl
+from .Multiplot import *
+from .Dataset import Dataset
+from .Inspection import Inspect
+from ..mods.Constants import *
+from ..mods.Cosmology import *
+from ..mods.Interpolate import Interpolate
+from ..mods.InitializeGrid import InitializeGrid
+from ..mods.RadiationSources import RadiationSources
+from ..mods.SecondaryElectrons import SecondaryElectrons
+from ..mods.ComputeCrossSections import PhotoIonizationCrossSection
+from ..mods.InitializeIntegralTables import InitializeIntegralTables
+
+class Analyze:
+    def __init__(self, pf, retabulate = False):
+        self.ds = Dataset(pf)
+        self.data = self.ds.data
+        self.t = self.ds.t
+        self.dt = self.ds.dt
+        
+        self.pf = self.ds.pf        # dict
+        self.g = InitializeGrid(self.pf)   
+        self.cosm = Cosmology(self.pf)
+        self.rs = RadiationSources(self.pf, retabulate = retabulate, path = self.ds.path_to_output)
+        self.esec = SecondaryElectrons(self.pf)         
+        
+        # Convenience
+        self.GridDimensions = int(self.pf['GridDimensions'])
+        self.grid = np.arange(self.GridDimensions)
+                
+        # Deal with log-grid
+        if self.pf['LogarithmicGrid']:
+            self.r = np.logspace(np.log10(self.pf['StartRadius'] * self.pf['LengthUnits']),
+                np.log10(self.pf['LengthUnits']), self.GridDimensions + 1)
+        else:
+            self.r = np.linspace(self.pf['StartRadius'] * self.pf['LengthUnits'], 
+                self.pf['LengthUnits'], self.GridDimensions + 1)
+        
+        self.dx = np.diff(self.r)   
+        self.r = self.r[0:-1]
+                
+        self.Vsh = 4. * np.pi * ((self.r + self.dx)**3 - self.r**3) / 3. / cm_per_mpc**3
+                
+        # Shift radii to cell-centered values
+        self.r += self.dx / 2.   
+                                
[…]

### 2012-12-18 rt1d/analysis/Dataset.py

@@ -1,0 +1,115 @@
+"""
+Dataset.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on 2010-12-01.
+
+Description: General module for command line analysis of rt1d data.
+     
+"""
+
+import os, re, h5py
+import numpy as np
+import pylab as pl
+from .DataDump import DataDump
+from ..mods.ReadParameterFile import ReadParameterFile
+        
+class Dataset:
+    def __init__(self, pf):
+        """
+        Initialize our analysis environment.  The variable 'dataset' can be either the master
+        parameter file for a series of runs, or an individual parameter file for a single run.
+        The parameter 'gd' is the global directory, which will default to the directory where
+        we launched python from.
+        
+        Some jargon:
+            'ds' = dataset, this is the dictionary we'll ultimately return
+            'dsn' = dataset name, a string that is the name of the directory all datadumps live in
+            'dd' = datadump, referring to the data in a specific time output in the entire dataset
+            'ddf' = datadump file, just the filename of a particular dd, like 'dd0000.h5'
+        
+        """   
+        
+        # Global directory
+        self.gd = os.getcwd()
+                
+        # Run directory
+        self.rd = pf.rpartition('/')[0]
+        
+        if not self.rd.strip():
+            self.rd = '.'        
+                
+        # Read in parameter file
+        self.pf = ReadParameterFile(pf)
+                        
+        # Also need path to parameter file (not including the parameter file itself)
+        # NOTE: THIS IS RELATIVE TO RUN DIRECTORY
+        self.od = self.pf['OutputDirectory']
+        
+        # Path to all output - ABSOLUTE PATH
+        self.path_to_output = '%s/%s/%s' % (self.gd, self.rd, self.od)
+        self.path_to_output = self.path_to_output.strip().rstrip('.')
+        
+        self.data = self.load()
+        self.t, self.dt = self.read_timestep_evolution()
+                                     
+    def load(self):
+        """
+        Return object containing access to all datadumps for the run we've specified.
+        """
+        
+        # List all data*dumps* in this data*set*.
+        alldds = []
+        for f in os.listdir(self.path_to_output):
+            if not re.search('.h5', f): 
+                continue
+            if not re.search('dd', f): 
+                continue # temporary hack
+            alldds.append(f)
+            
+        ds = {}
+        for ddf in alldds:
+            f = h5py.File("%s/%s" % (self.path_to_output, ddf))
[…]

### 2012-12-18 rt1d/analysis/Multiplot.py

@@ -1,0 +1,174 @@
+"""
+Multiplot.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on 2011-06-27.
+
+Description: Make multipanel plots with shared axes, with or without AxesGrid.
+     
+"""
+
+import pylab as pl
+import numpy as np
+from mpl_toolkits.axes_grid1 import AxesGrid, make_axes_locatable
+
+# Matplotlibrc defaults
+figsize = pl.rcParams['figure.figsize']
+wspace = pl.rcParams['figure.subplot.wspace']
+hspace = pl.rcParams['figure.subplot.hspace']
+left = pl.rcParams['figure.subplot.left']
+right = pl.rcParams['figure.subplot.right']
+bottom = pl.rcParams['figure.subplot.bottom']
+top = pl.rcParams['figure.subplot.top']
+dtop = 1 - top
+dbottom = bottom
+dright = 1 - right
+dleft = left
+
+def AxisConstructor(nr, nc, panel): 
+    return pl.subplot(nr, nc, panel)
+
+class multiplot:
+    def __init__(self, dims = (2, 2), padding = 0, panel_size = None, useAxesGrid = True,
+        share_all = True, aspect = False):
+        """
+        dims = (rows, columns)
+        padding = separation of subplots in inches
+        panel_size = fraction of default figsize in each dimension
+        nd = dimensionality of data
+        """
+        
+        self.dims = dims
+        self.padding = padding
+        
+        if panel_size is None: 
+            self.panel_size = np.array([figsize[0], figsize[1]])
+        else: 
+            
+            xsize = figsize[0]
+            if panel_size[0] > 1:
+                
+                for i in xrange(panel_size[0] - 1):
+                    xsize += (right - left) * figsize[0]
+            
+            ysize = figsize[1]    
+            if panel_size[1] > 1:
+                for i in xrange(panel_size[1] - 1):
+                    ysize += (top - bottom) * figsize[1]
+            
+            self.panel_size = np.array([xsize, ysize])
+                    
+        if (share_all and not useAxesGrid) or not share_all:
+            pl.rcParams['figure.subplot.wspace'] = self.padding
+            pl.rcParams['figure.subplot.hspace'] = self.padding        
+                
+        self.fig = pl.figure(1, self.panel_size)
+        
+        self.Naxes = np.prod(self.dims)
[…]

### 2012-12-18 rt1d/mods/ComputeCrossSections.py

@@ -1,0 +1,87 @@
+"""
+ComputeCrossSections.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on 2010-10-25.
+
+Description: Compute the bound-free absorption cross-sections via the fitting
+formulae of Verner et al. 1996, and eventually others.
+     
+"""
+
+import numpy as np
+from .Constants import h
+from scipy.integrate import quad
+
+E_th = [13.6, 24.6, 54.4]
+params = [[4.298e-1, 5.475e4, 3.288e1, 2.963, 0.0, 0.0, 0.0],
+          [13.61, 9.492e2, 1.469, 3.188, 2.039, 4.434e-1, 2.136],
+          [1.72, 1.369e4, 3.288e1, 2.963, 0.0, 0.0, 0.0]]
+
+def PhotoIonizationCrossSection(E, species = 0):
+    """ 
+    Returns photoionization cross section for HI, HeI, or HeII from the fits of
+    Verner et al. 1996.  HI is the first 7-element sub-array in 'params', HeI
+    is the second, and HeII is the third.  In order, the coefficients in these arrays are:
+        
+        E_0, sigma_0, y_a, P, y_w, y_0, y_1
+        
+        Also:
+        species = 0 for HI
+        species = 1 for HeI
+        species = 2 for HeII
+        
+        Note: The units are cm^2.
+    
+    """                            
+    
+    x = (E / params[species][0]) - params[species][5]
+    y = np.sqrt(x**2 + params[species][6]**2)
+    F_y = ((x - 1.0)**2 + params[species][4]**2) * \
+        y**(0.5 * params[species][3] - 5.5) * \
+        (1.0 + np.sqrt(y / params[species][2]))**-params[species][3]
+                                
+    return params[species][1] * F_y * 1e-18
+                                                                                                               
+def EffectiveCrossSection(rs, E1, E2, species = 0):
+    """
+    Compute frequency averaged cross-section, as in Aubert & Teyssier 2008.
+    
+        rs: Instance of RadiationSource class.
+    """
+        
+    if rs._name == 'RadiationSourceIdealized':
+        Q12 = quad(lambda x: rs.Spectrum(x) / x, E1, E2)[0]
[…]

### 2012-12-18 rt1d/mods/ComputeRateCoefficients.py

@@ -1,0 +1,559 @@
+"""
+
+ComputeRateCoefficients.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on: Tue Jan 17 16:57:48 2012
+
+Description: 
+
+"""
+
+import numpy as np
+
+from .Constants import erg_per_ev, k_B, c, m_e, sigma_SB, \
+    sigma_T, h, nu_alpha, E_LyA
+from .Cosmology import Cosmology
+from .Interpolate import Interpolate
+from .RadiationSource import RadiationSource
+from .SecondaryElectrons import SecondaryElectrons
+from .ComputeCrossSections import *
+from scipy.integrate import quad
+
+E_th = np.array([13.6, 24.6, 54.4])
+
+class RateCoefficients:
+    def __init__(self, pf, rs = None, grid = None):
+        self.pf = pf        
+        self.rs = rs
+        self.cosm = Cosmology(pf)
+        self.esec = SecondaryElectrons(pf)
+        
+        self.mask = np.array([[1, 0, 0], [0, 1, 0], [0, 0, 1]])    
+        if pf['MultiSpecies']:
+            self.donors = np.arange(3)
+        else:
+            self.donors = np.array([0]) 
+        
+        self.Vsh = 4. * np.pi * ((grid.r + grid.dx)**3 - grid.r**3) / 3.
+        
+        self.smallcol = pf['OpticallyThinColumn']
+                                    
+    def ConstructArgs(self, args, indices, Lbol, r, ncol, T, dr, t, z, cell):
+        """
+        Make list of rate coefficients that we'll pass to solver.
+        
+            incoming args: [nabs, nion, n_H, n_He, n_e]
+            outgoing args: [nabs, n_H, n_He, n_e, Gamma, gamma, Beta, alpha, 
+                k_H, zeta, eta, psi, xi, omega, theta, Jc, Ji]
+            
+            ***REMEMBER: ncol is really np.log10(ncol)!
+            
+        """    
+                
+        nabs = args[0]
+        nion = args[1]
+        n_H = args[2]
+        x_HII = (n_H - nabs[0]) / n_H
+        n_He = args[3]
+        n_e = args[4]
+        
+        Gamma = np.zeros([3, self.rs.Ns])
+        gamma = np.zeros([3, 3, self.rs.Ns])
+        alpha = np.zeros(3)
+        Beta = np.zeros(3)
[…]

### 2012-12-18 rt1d/mods/ControlSimulation.py

@@ -1,0 +1,247 @@
+"""
+
+ControlSimulation.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on: Wed Jan 18 13:39:42 2012
+
+Description: Routines used to control simulation (adaptive timestep, 
+load-balancing, etc.)
+
+"""
+
+import copy
+import numpy as np
+from .Cosmology import Cosmology
+
+try:
+    from mpi4py import MPI
+    rank = MPI.COMM_WORLD.rank
+    size = MPI.COMM_WORLD.size
+except ImportError:
+    rank = 0
+    size = 1    
+
+class ControlSimulation:
+    def __init__(self, pf):
+        self.pf = pf
+        self.cosm = Cosmology(pf)
+        
+        self.GridDimensions = int(pf['GridDimensions'])
+        self.grid = np.arange(self.GridDimensions)
+        self.R0 = pf['LengthUnits'] * pf['StartRadius']
+        self.z0 = pf['InitialRedshift']
+                
+        self.mask = np.array([[1, 0, 0], [0, 1, 0], [0, 0, 1]])
+                
+    def ComputeInitialPhotonTimestep(self, data, r):
+        """
+        Compute photon timestep based on maximum allowed fractional change
+        in hydrogen and helium neutral fractions (Shapiro et al. 2004).
+        """          
+        
+        # Read in some data, mostly for first cell
+        n_H = data['HIDensity'][0] + data['HIIDensity'][0]
+        n_He = data['HeIDensity'][0] + data['HeIIDensity'][0] + data['HeIIIDensity'][0]
+        n_e = data['HIIDensity'][0] + data['HeIIDensity'][0] + 2. * data['HeIIIDensity'][0]
+        nabs = np.array([data['HIDensity'][0], data['HeIDensity'][0], data['HeIIDensity'][0]])
+        nion = np.array([data['HIIDensity'][0], data['HeIIDensity'][0], data['HeIIIDensity'][0]])
+        T = data['Temperature'][0]
+        
+        ncol = 1. * np.ones(3)
+        ncell = r.dx[0] * nabs    
+        nout = np.log10(ncell)
+        Vsh = r.coeff.ShellVolume(self.R0, r.dx[0])    
+        logxHII = np.log10(nion[0] / n_H)
+        
+        Lbol = []
+        indices_in = []
+        for source in r.rs.all_sources:
+            Lbol.append(source.BolometricLuminosity(0)) 
+            if source.TableAvailable:
[…]

### 2012-12-18 rt1d/mods/Interpolate.py

@@ -1,0 +1,214 @@
+"""
+Interpolate.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on 2010-10-30.
+
+Description: Various interpolation routines
+     
+"""
+
+import numpy as np
+from .SecondaryElectrons import SecondaryElectrons
+from scipy.interpolate import LinearNDInterpolator
+
+class Interpolate:
+    def __init__(self, pf, iits):
+        self.pf = pf
+        self.esec = SecondaryElectrons(pf)
+        self.iits = iits
+                
+        self.locs = iits.locs
+        self.dims = iits.dims        
+        self.columns = iits.columns
+        self.dcolumns = iits.dcolumns
+        self.colmin = np.array(map(np.min, self.columns))
+        self.colmax = np.array(map(np.max, self.columns))
+        self.offsets = self.colmin / self.dcolumns
+                
+        # This is a dictionary with all the lookup tables
+        self.itabs = iits.itabs
+                
+        # What kind of interpolator do we need? Where to look for values?
+        if iits.Nd == 1:
+            self.interp = self.InterpolateLinear
+            self.GetIndices = self.GetIndices1D
+        elif iits.Nd == 2:
+            self.interp = self.InterpolateBiLinear
+            self.GetIndices = self.GetIndices2D
+        elif iits.Nd == 3:
+            self.interp = self.InterpolateTriLinear
+            self.GetIndices = self.GetIndices3D
+        elif iits.Nd == 4:
+            self.interp = self.InterpolateSphericalLinear
+            self.GetIndices = self.GetIndices4D
+        else:
+            self.interp = self.InterpolateQuintLinear
+            self.GetIndices = self.GetIndices5D    
+                    
+    def OpticalDepth(self, value, species):
+        return 10**np.interp(value, self.columns[species], self.itabs['logOpticalDepth%i' % species])
+                
+    def InterpolateLinear(self, indices, integral, value):
+        """
+        Use this technique for hydrogen-only calculations.  For consistency with MultiSpecies > 0 
+        methods, 'value' should still be a 3-element list.  
+        """    
+
+        return 10**np.interp(value[0], self.columns[0], self.itabs[integral])
+        
+    def InterpolateBiLinear(self, indices, integral, value):
[…]

### 2012-12-18 rt1d/mods/ProblemTypes.py

@@ -1,0 +1,340 @@
+"""
+
+ProblemTypes.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on: Wed Mar  7 15:53:15 2012
+
+Description: Non-default parameter sets for certain test problems.
+
+Note: Integer problem types imply use of a continuous SED, while their
+non-integer counterparts imply a discrete SED, except for ProblemType = 1 = 1.1.
+I generally (for clarity) order parameters from top to bottom in the following
+way:
+    Units, time/data dump interval
+    Integral tabulation
+    Initial conditions
+    Source parameters
+    Physics parameters
+    
+More notes:
+10 + ProblemType = same as ProblemType but MultiSpecies = 1    
+
+"""
+
+import numpy as np
+from .Constants import cm_per_kpc, cm_per_mpc, s_per_myr, m_H
+
+def ProblemType(ptype):
+    """
+    Storage bin for predefined problem types, like those used in the 
+    radiative transfer comparison project ('RT06').
+    """
+    
+    # RT06-0.3, Single zone ionization/heating, then source switches off.
+    if ptype == 0:
+        pf = {"ProblemType": 0, 
+              "TabulateIntegrals": 1, 
+              "LengthUnits": 1e-4 * cm_per_kpc, # 100 milliparsecs 
+              "GridDimensions": 1, 
+              "StartRadius": 0.99, # cell = 1 milliparsec across
+              "StopTime": 10, 
+              "LogarithmicDataDump": 1,
+              "InitialLogDataDump": 1e-12,
+              "ElectronRestrictedTimestep": 1,
+              "NlogDataDumps": 100,
+              "dtDataDump": 0.1, 
+              "ODEMinStep": 1e-15,
+              "MinimumSpeciesFraction": 1e-9, 
+              "ColumnDensityBinsHI": 100,
+              "DensityUnits": m_H,
+              "DensityProfile": 0,
+              "InitialDensity": 1., 
+              "TemperatureProfile": 0, 
+              "InitialTemperature": 1e2,
+              "IonizationProfile": 0, 
+              "InitialHIIFraction": 1e-6, 
+              "SourceType": 1, 
+              "SpectrumType": 1,
+              "SourceLifetime": 0.5,
+              "OpticalDepthDefiningIFront": [0, 0, 0],
[…]

### 2012-12-18 rt1d/mods/RadiationSourceFromFile.py

@@ -1,0 +1,159 @@
+"""
+
+RadiationSourceFromFile.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on: Sun Jul 22 16:27:37 2012
+
+Description: RadiationSource class used to handle user-defined SEDs supplied
+via ASCII or HDF5 file.
+
+"""
+
+import h5py
+import numpy as np
+
+from .InitializeIntegralTables import *
+
+class RadiationSourceFromFile:
+    def __init__(self, pf):
+        self.pf = pf
+        
+        # Create SpectrumPars attribute    
+        self.SpectrumPars = listify(pf)
+        self.N = len(self.SpectrumPars['Type'])
+        
+        # Cast types to int to avoid indexing complaints
+        self.SpectrumPars['Type'] = map(int, self.SpectrumPars['Type'])   
+        
+        self._name = 'RadiationSourceFromFile'
+        
+    def initialize(self):
+        """
+        Create attributes we need, normalize, etc.
+        """
+            
+        self.fn = self.pf['SpectrumFile']    
+        self.tau = self.pf['SourceLifetime'] * self.pf['TimeUnits']
+                
+        # Read spectrum - expect hdf5 with (at least) E, L_E, and time_yr datasets.
+        f = h5py.File(self.fn)
+        self.E = f['E'].value
+        self.t = self.Age = f['time_yr'].value * s_per_yr
+        self.maxAge = np.max(self.Age)
+        self.Nt = len(self.t)
+        self.L_E = f['L_E'].value
+        
+        self.Emin = np.min(self.E)
+        self.Emax = np.max(self.E)
+        
+        # Threshold indices
+        self.i_Eth = np.zeros(3)
+        for i, energy in enumerate(E_th):
+            loc = np.argmin(np.abs(energy - self.E))
+            
+            if self.E[loc] < energy:
+                loc += 1
+            
+            self.i_Eth[i] = loc
+        
+        self.Lbol = self.BolometricLuminosity(0)
+        
+    def SourceOn(self, t):
+        if t <= self.maxAge:
+            return True
+        else:
+            print 'WARNING: Current time lies outside bounds of spectrum table. Source now OFF.'
+            return False
+        
+    def Spectrum(self, E = None, t = 0.0):
+        """
+        Return specific luminosity.
+        """
+        
+        i = self.get_time_index(t)        
+        return self.L_E[i] / self.BolometricLuminosity(t = t)
[…]

### 2012-12-18 rt1d/util/fields.py

@@ -1,0 +1,19 @@
+"""
+
+fields.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on: Thu Sep 20 13:29:23 2012
+
+Description: 
+
+"""
+
+all_fields = \
+{
+
+ "eta": 'RadiativeRecombinationRate'
+
+}
+

### 2012-12-18 scripts/AnalyzeResolutionSuite.py

@@ -1,0 +1,98 @@
+"""
+
+AnalyzeResolutionSuite.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on: Wed Sep  7 10:12:23 2011
+
+Description: Compare results of resolution suites for RT1 and RT2.
+
+Conventions: 
+-Black solid line is reserved for the analytic I-front solution.
+-Blue = highest grid resolution 
+-Green = second highest
+-Red = poorest
+-Solid -> highest time resolution (adaptive)
+-Dashed -> dt = 1e-2
+-Dotted -> dt = 1e-1
+
+*A dash-dotted yellow line will be assigned to something not fitting the above criteria
+
+Notes: Run from directory containing all parameter files.
+
+"""
+
+import os, re, sys
+import pylab as pl
+import numpy as np
+from jmpy.stats import *
+import rt1d.analysis as rta
+
+pl.rcParams['figure.subplot.left'] = 0.12
+pl.rcParams['figure.subplot.right'] = 0.88
+pl.rcParams['legend.fontsize'] = 10
+
+path = os.getcwd()
+pfs = ['dx25', 'dx50', 'dx100', 'dx200', 'dx400', 'dx800']#, 'dx1600', 'dx3200']
+
+resolution = []
+meanerror = []
+maxerror = []
+minerror = []
+
+# Highest resolution run
+ds = rta.Analyze("{0}/{1}.dat".format(path, 'dx800'))
+ds.ComputeIonizationFrontEvolution()
+ptype = ds.pf['ProblemType']
+
+# Loop over simulations
+mp = rta.multiplot(dims = (2, 1), panel_size = (1, 1), useAxesGrid = False)
+for pf in pfs:
+    print "Loading {0}/{1}".format(path, pf)
+    
+    try: ds = rta.Analyze("{0}/{1}.dat".format(path, pf))
+    except OSError: continue
+    
+    if not os.path.exists("{0}/{1}".format(path, ds.pf["OutputDirectory"])): 
+        continue
+
+    if ds.pf['GridDimensions'] == 25: color = 'cyan'
+    elif ds.pf['GridDimensions'] == 50: color = 'magenta'
+    elif ds.pf['GridDimensions'] == 100: color = 'red'
[…]

### 2012-12-18 scripts/AnalyzeTest1.py

@@ -1,0 +1,74 @@
+"""
+AnalyzeTest1.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on 2010-11-09.
+
+Description: Locate the position of the I-front (defined as x_H = 0.5) for each data dump in our 1D calculation, and
+plot this position as a function of time (distance in kpc, time in Myr).  Also, plot neutral and ionized fractions
+vs. radius and time.  Eventually compare with Enzo.
+
+Notes: Supply parameter file as commmand line argument.     
+     
+"""
+
+import os, re, h5py, sys
+import numpy as np
+import pylab as pl
+import rt1d.analysis as rta
+
+cm_per_kpc = 3.08568 * 10**21
+s_per_myr = 365.25 * 24 * 3600 * 10**6
+
+# Read in data
+ds = rta.Analyze(sys.argv[1])
+ds.ComputeIonizationFrontEvolution()
+
+# Plot I-front evolution (2-panels)
+mp = rta.multiplot(dims = (2, 1), panel_size = (1, 1), useAxesGrid = False)
+
+mp.grid[0].plot(ds.t / ds.trec, ds.rIF, color = 'k', ls = '--')
+mp.grid[0].plot(ds.t / ds.trec, ds.ranl, linestyle = '-', color = 'k')
+mp.grid[0].set_xlim(0, max(ds.t / ds.trec))
+mp.grid[0].set_ylim(0, 1.1 * max(max(ds.rIF), max(ds.ranl)))
+mp.grid[0].set_ylabel(r'$r \ (\mathrm{kpc})$')  
+
+mp.grid[1].plot(ds.t / ds.trec, ds.rIF / ds.ranl, ls = '-', color = 'k')
+mp.grid[1].set_xlim(0, max(ds.t / ds.trec))
+mp.grid[1].set_ylim(0.9, 1.1)
+mp.grid[1].set_xlabel(r'$t / t_{\mathrm{rec}}$')
+mp.grid[1].set_ylabel(r'$r/r_{\mathrm{anl}}$') 
+
+mp.grid[0].xaxis.set_ticks(np.linspace(0, 4, 5))
+mp.grid[1].xaxis.set_ticks(np.linspace(0, 4, 5))
+
+mp.fix_ticks()
[…]

### 2012-12-18 scripts/RunAllTests.py

@@ -1,0 +1,57 @@
+"""
+
+RunAllTests.py
+
+Author: Jordan Mirocha
+Affiliation: University of Colorado at Boulder
+Created on: Wed Aug 29 11:07:49 2012
+
+Description: Run all test problems, and analyze results.
+
+Note: This script will copy everything it needs from $RT1D/doc/tests, and
+store the output in your CWD.
+
+"""
+
+import os, re
+
+RT1D = os.environ.get('RT1D')
+os.system('cp -r %s/doc/tests/* .' % RT1D)
+
+testdir = os.getcwd()
+
+if not os.path.exists('frames'):
+    os.mkdir('frames')
+
+for fn in os.listdir(os.getcwd()):
+    if not os.path.isdir(fn):
+        continue
+        
+    os.chdir(fn)
+    os.system('cp %s/bin/RT1D.py .' % RT1D)
+
+    # Run all simulations in this dir (given by .dat parameter files)
+    for pf in os.listdir(os.getcwd()):
+        if not re.search('.dat', pf):
+            continue
+        
+        os.system('python RT1D.py %s' % pf)
+        
+    # Run analysis
+    for py in os.listdir(os.getcwd()):
+        if not re.search('.py', py):
+            continue
+        if py == 'RT1D.py':
+            continue    
+                        
+        os.system('python %s' % py)  
+        os.system('cp *.png ../frames')      
+    
+    os.chdir(testdir)
+    
+        
+
+
+
+
+

### 2012-12-18 setup.py

@@ -1,0 +1,19 @@
+#!/usr/bin/env python
+
+import os
+from distutils.core import setup
+
+setup(name='rt1d',
+      version='1.0',
+      description='1D Radiative Transfer',
+      author='Jordan Mirocha',
+      author_email='mirochaj@gmail.com',
+      url='https://bitbucket.org/mirochaj/rt1d',
+      packages=['rt1d', 'rt1d.mods', 'rt1d.analysis'],
+     )
+     
+if not os.path.exists('bin'): 
+    os.mkdir('bin')     
+os.system('cp rt1d/mods/RT1D.py bin')
+
+     
